---
title: "Climate Negotiation Update"
author: "UNFCCC secretariat"
date: "`r format(Sys.time(), '%H:%M %z  %d %b, %Y')`"
output: html_document
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../docs", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r getgoogledata, echo=FALSE, results='hide', message=FALSE}
# climate negotiation simulation

## Get googlesheet from survey
#Install the required package
#install.packages('googlesheets4')
#Load the required library 
library(googlesheets4)
library(dplyr)
library(lubridate)
library(xtable)
library(kableExtra)
#gs4_auth(options(gargle_verbosity = "debug"))

gs4_deauth()
import <- read_sheet("https://docs.google.com/spreadsheets/d/1pazIs994LwzfG5BSb_2wC8aGsM3GwDg_uc5ecO9Od0Y/edit?usp=sharing")
str(import)

df <- import
```

```{r cleanup, echo=FALSE, results='hide', warning=FALSE}
names(df)=c("Timestamp", "Country", "Peak.Yr", "Reduction.Yr", "Reduction.Rate",
            "Deforestation", "Afforestation", "Mitigation", "Billion")
df$Block = NA

### Select Most Recent Submission
df <- df %>%
group_by(Country) %>%
filter(Timestamp == max(Timestamp)) %>%
ungroup()


### Define Negotiations Blocks
df$Block[df$Country == "USA"] <- "USA"
df$Block[df$Country == "European Union"] <- "EU"
        
df$Block[df$Country == "Japan" | df$Country == "Russia"] <- "Other Developed"

df$Block[df$Country == "China"] = "China"
df$Block[df$Country == "India"] = "India"

df$Block[df$Country == "Ghana" | df$Country == "Philippines" |
           df$Country == "Rwanda" | df$Country == "Venezuela" |
           df$Country == "Nigeria"] <- "Developing"


df$Billion[df$Mitigation=="Withdrawing"] <- df$Billion[df$Mitigation=="Withdrawing"] *-1


a <- max(df$Timestamp)
a.country <- df$Country[df$Timestamp==a]

#"2017-01-01 03:01:02 GMT"

```
  
# Climate in the News

## AKP Negotiation News

KYOTO, Japan -- EU makes a late breaking commitment to delay reductions, but makes not contribution to the mitigation funds. 

Climate Activists surprised by the abject failure of the developed countries to address the North-South inequality in climate change emissions sources and environmental costs. 

# Negotiation Updates

## Recent Climate Commitment

This report is based on the most recent entry submitted by the `r a.country` (`r a`).

```{r processing, echo=FALSE, results='hide'}

## Process Data (averages?)

### Select Most Recent Timestamp
names(df)

Block.means = df %>% 
  group_by(Block) %>% 
  summarise_at(c("Peak.Yr", "Reduction.Yr", "Reduction.Rate", "Deforestation", "Afforestation"),
               ~round(mean(., na.rm=T),0)) %>% as.data.frame()


names(df)
Block.mitigation = df %>% 
  group_by(Block) %>% 
  summarise_at(c("Billion"), ~round(sum(., na.rm=T), 0)) %>% as.data.frame()


Block.Results <- merge(Block.means, Block.mitigation)
Block.Results

## Normalize based on number of developing countries and population %
df$Country[df$Block=="Developing"]
df$Population <- NA
df$Population[df$Block=="Developing"] <- c(3.2, 210, 111, 13.2, 28)
# df$NormalizingPop <- 67000/df$Population 
df$BillionNormalized <- df$Billion
df$BillionNormalized[df$Block=="Developing"] <- 152/5 * df$Billion[df$Block=="Developing"]

mitigation <- df %>% 
  group_by(Mitigation) %>% 
  summarise_at(c("BillionNormalized"), sum, na.rm=T)

total = data.frame(Mitigation="Total", BillionNormalized=sum(mitigation$BillionNormalized))
if(total$BillionNormalized <0) total$Mitigation="Deficit"

mitigation = rbind(mitigation, total)

names(mitigation) = c("Mitigation Transfers", "Funds")

## Link to Climate Negotiation?


## Create Table
### C-Road Order
index = c(6, 3, 5, 1, 4, 2)

Block.Results$Index = index
Block.Results

str(Block.Results)
```

## Negotiation Results

Based on the submissions, the commitments are as as follows:

```{r results, echo=FALSE, results='asis'}

knitr::kable(Block.Results[index,][,-8], align = "lccrrrc", row.names = FALSE) %>%
  kable_styling(full_width = F)
#print(xtable(Block.Results[index,][,-8]), include.rownames=FALSE, type = "html")
```

These have values have been used to update the C-Roads Projections. 

## Projected Temperature Change

Based on our current commitments, we are looking at a 2.3 degree temperature change by the year 2100. The potential impacts for this remain extremely worrisome and states need to do more. 

[2.3 degree Projection -- EU Revised Commitment ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2024&pc612=5&pc613=12&pc620=2026&pc621=2032&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=9&pc643=40&pc644=10&pc650=2032&pc651=2039&pc652=4&pc653=15&pc654=3&g0=515-6r&g1=86)

### Previous Results

[2.3 degree Projection -- Russia  & India Commitment Update 2x ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2026&pc621=2032&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=9&pc643=40&pc644=10&pc650=2032&pc651=2039&pc652=4&pc653=15&pc654=3&g0=515-6r&g1=140&g1-type=2&g1-lat=23.3544&g1-lon=91.0301&g1-zm=6&g1-yr=2100&g1-ss=4)

[2.4 degree Projection -- India Commitment Update ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=4&pc643=40&pc644=10&pc650=2032&pc651=2039&pc652=4&pc653=15&pc654=3&g0=502-6r&g1=140&g1-type=2&g1-lat=23.3544&g1-lon=91.0301&g1-zm=6&g1-yr=2100&g1-ss=4)

[2.4 degree Projection -- Nigeria Commitment Update ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2032&pc651=2039&pc652=4&pc653=15&pc654=3&g0=515-6r&g1=86)




[2.4 degree Projection -- Ghana Commitment Update ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2037&pc651=2043&pc652=5&pc653=19&pc654=7&g0=515-6r&g1=86)


[2.5 degree Projection -- Philippines Commitment Update ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2039&pc651=2049&pc652=3&pc653=21&pc654=9&g0=515-6r&g1=86)


[2.6 degree Projection USA Commitment Update](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2051&pc651=2062&pc652=3&pc653=6&pc654=6&g0=501-6r&g1=86)

### Previous Results

[Link to Scenario based on 11/20/22 Commitments](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2051&pc651=2062&pc652=3&pc653=6&pc654=6&g0=501-6r&g1=86)

## Mitigation Funds -- Totals

The UNFCCC has estimated a need of 200 billion USD to fund the mitigation fund, and probably requiring continuous updating and the need grows. 

In addition to this, a new fund, "Losses and Damages", is needing funding. At this point, we are combining the two funds, which are normalized (we have 5 developing countries and their demand is being mulitiplied by 152/5 (152 developing countries/5 represented)).

The Philippines is asking for 1 Trillion USD! While it is true they are facing massive impacts, the Developed Countries should get a lot for their money!

The developed countries need to do more to address these massive economic impacts being faced in developing countries. 

But developing countries should give up more of their reliance on fossil fuels. India and China should be setting as role models.


```{r mitigation, echo=FALSE, results='asis'}
knitr::kable(mitigation, align = "lr", row.names = FALSE) %>%
  kable_styling(full_width = F)
#print(xtable(mitigation), include.rownames=FALSE, type = "html")
```
