---
title: "Climate Negotiation Update"
author: "UNFCCC secretariat"
date: "`r format(Sys.time(), '%H:%M %z  %d %b, %Y')`"
output: html_document
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../docs", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r getgoogledata, echo=FALSE, results='hide', message=FALSE}
# climate negotiation simulation

## Get googlesheet from survey
#Install the required package
#install.packages('googlesheets4')
#Load the required library 
library(googlesheets4)
library(dplyr)
library(lubridate)
library(xtable)
library(kableExtra)
#gs4_auth(options(gargle_verbosity = "debug"))

gs4_deauth()
import <- read_sheet("https://docs.google.com/spreadsheets/d/1pazIs994LwzfG5BSb_2wC8aGsM3GwDg_uc5ecO9Od0Y/edit?usp=sharing")
str(import)

df <- import
```

```{r cleanup, echo=FALSE, results='hide', warning=FALSE}
names(df)=c("Timestamp", "Country", "Peak.Yr", "Reduction.Yr", "Reduction.Rate",
            "Deforestation", "Afforestation", "Mitigation", "Billion")
df$Block = NA

### Select Most Recent Submission
df <- df %>%
group_by(Country) %>%
filter(Timestamp == max(Timestamp)) %>%
ungroup()


### Define Negotiations Blocks
df$Block[df$Country == "USA"] <- "USA"
df$Block[df$Country == "European Union"] <- "EU"
        
df$Block[df$Country == "Japan" | df$Country == "Russia"] <- "Other Developed"

df$Block[df$Country == "China"] = "China"
df$Block[df$Country == "India"] = "India"

df$Block[df$Country == "Ghana" | df$Country == "Philippines" |
           df$Country == "Rwanda" | df$Country == "Venezuela" |
           df$Country == "Nigeria"] <- "Developing"


df$Billion[df$Mitigation=="Withdrawing"] <- df$Billion[df$Mitigation=="Withdrawing"] *-1


a <- max(df$Timestamp)
a.country <- df$Country[df$Timestamp==a]

#"2017-01-01 03:01:02 GMT"

```
  
# Climate in the News

## AKP 

KYOTO, Japan -- Negotiators agree on commitments that suggest the world will be 2.4 degrees warmer in 2100, where developing countries are making strong statements. Ghana and the Philippines have made substantial commitments (and demands). The question for the negotiations focuses more and more on India, China, and Japan. According to Professor Los Huertos, even modest reductions in fossil fuels by China and India could have substantial impacts.

"We must see more concrete action to reduce our emissions and the developed world has to demonstrate some leadership there," says Los Huertos.

## Nigeria

Dickson, 29, from Abuja, Nigeria, is the founder of the Eco Clean Active Initiative, a youth-led nonprofit working on improving grassroots climate mitigation, and an organiser for Fridays for Future in Africa. He was unable to attend Cop26 due to Britain’s vaccine rules.

“I am privileged to have knowledge about the climate crisis, so came here to seek justice for the communities in my country who are mostly impacted, especially the young people who make up 16.8% of our population. Getting to know other activists whom I’d only known on social media has been special.

Sign up to Down to Earth

Free weekly newsletter
The planet's most important stories. Get all the week's environment news - the good, the bad and the essential


“I was very, very happy to meet Vanessa Nakate because she has been an inspiration and we have been working together for the past three years, so it was a joy to finally talk in person. Also Licypriya Kangujam, the 11-year-old activist from India. I’ve also had the opportunity of meeting political leaders like our environment minister, and I called his attention to how young people are suffering in our country, and that we grassroots leaders need support to develop our communities.

“Getting access to such personalities in Nigeria is quite difficult. I was also invited to a meeting with other youth leaders with the European delegation, and I am hopeful that something will come from sharing about my community.

“My highlight has been meeting Al Gore, this was a great privilege for me because his organisation trained me in 2020 as a climate leader. Meeting him and expressing myself about our challenges is something I will never forget. He shook my hand and said he has been following the floods in Nigeria which have killed 600 people and affected 2 million, and that as climate leaders we all have to work together to address the crisis.

“The most disappointing thing is that so many African activists were denied access because they couldn’t fund it, but also hearing the US president committing $150m to help Africa adapt to the climate crisis. This won’t do anything, we need trillions. It’s my first Cop, so it has not been easy following the negotiations, and as an observer with a yellow badge I can’t access the meeting rooms. It’s also very tiring. We’re here all day without eating because the food is so expensive. But I have gained a lot of experience and knowledge so hopefully at the next Cop in Dubai I can be even more active.”

## Rwanda

The Government of Rwanda will strengthen its support to the private sector, particularly in the implementation of climate-friendly economy projects. On 7 November 2022, Rwandan President Paul Kagame launched a $104 million fund at the Rwandan pavilion at the 27th United Nations Climate Change Conference (COP27) in Sharm-el-Sheikh, Egypt. The green investment facility called “Ireme Invest” will provide financing for the implementation of green and resilient projects in Rwanda. The aim is to accelerate the transition to a low-carbon economy in the East African country.

The fund will target projects in key sectors, including renewable energy deployment (solar and hydro), improving energy efficiency in industrial processes, introducing vehicle emission standards, deploying electric vehicles and promoting the use of biogas on farms. “Climate adaptation requires big changes in the way the world produces and consumes energy, in the way people move, in what they eat and in the way it is grown,” said President Paul Kagame at the launch of the Ireme Invest fund.

The Development Bank of Rwanda (BRD) has invested seed capital in the fund. BRD, which builds on a strong track record of the Rwanda Green Fund (FONERWA). According to Paul Kagame, the governments of France, Sweden and the United Kingdom, as well as the European Investment Bank (EIB) and the Green Climate Partnership Fund have also invested in the Green Investment Facility.

The new green investment fund is expected to enable the East African country to move forward with the implementation of its climate strategy unveiled in 2020. Rwanda’s plan includes 10 projects worth $2.4 billion planned by the Rwandan government to meet the Paris Agreement’s 2030 targets. One of these targets is the reduction of greenhouse gas (GHG) emissions. In 2015, these emissions were estimated at 5.3 million tonnes in Rwanda. Experts estimate that they could double by 2030.

## Ghana

The girl who scolded delegates
“Are there any other delegations that wish to make a statement?” President Shoukry asked the so-called stocktaking plenary, where an update of progress has been held daily.

The delegation of Ghana requested the floor, and passed the microphone to 10-year-old Nakeeyat Dramani Sam.

The young activist then began to scold delegates for their seeming failure to take the climate catastrophe seriously: they would act faster to end global warming if they were her age, she declared.

“If all of you were young people like me, wouldn’t you have already agreed to do what is needed to save our planet? Should we let the youth take over? Maybe only the youth delegation should be at the next COP”, she said, igniting a standing ovation from all the delegates present.

Ms. Dramani Sam urged adults to “have a heart” and “do the math”, referring to the science that indicates the gravity of climate change in the future.

“It is my real hope that COP27 will act for us. I'm sure nobody wants to betray us,” she said.

The child activist also asked nations to dig into their pockets and provide funds for those suffering the most.

“Some of the communities in my country are paying heavy prices since our planet was lit on fire by some people. This puts a simple question on the table…When can you pay us back? Because payment is overdue,” she stated.


## India

It gets almost three-quarters of its electricity from coal, and has 39 new coal-fired power plants under construction. It digs up and burns more of the stuff than any other country except China. And it is coal’s loudest advocate internationally: at last year’s climate conference in Glasgow, it was the skunk at the garden party, blocking efforts to phase out the fuel most responsible for global warming.

This soot-smeared intransigence, however, distracts from a dramatic countervailing trend. While his underlings defended coal, Narendra Modi, India’s prime minister, made a series of pledges in Glasgow that, if kept, will make his country a green-energy powerhouse. The most eye-catching was the promise that India would achieve “net-zero” emissions of greenhouse gases (ghgs) by 2070—meaning that any emissions that had not been eliminated by then would be offset in some way.

## Japan

Japan has launched an initiative for international knowledge sharing and capacity building on Article 6 at the sidelines of Cop 27, while disagreements remain over the article's draft text.

Article 6 governs the use of the voluntary carbon market and international carbon credit trading in meeting countries' Nationally Determined Contributions (NDCs).

Japan's environment ministry launched the Paris Agreement Article 6 Implementation Partnership on 16 November, which attracted pledges of participation from 40 countries and 23 institutions at its release.

The partnership aims to complement Article 6 by building a platform for countries to share knowledge and experiences on carbon markets and their links with NDCS. It also plans to provide training and support the development of carbon credit methodologies.

Although the Article 6 "rulebook" was finalised at Cop 26 last year, there are still unresolved issues around mechanisms to implement it that have struggled to make progress at this year's summit.

Parties remain at odds over the deadline for transferring emissions reduction projects listed under the Kyoto protocol's Clean Development Mechanism to the new registry, the transaction procedure for moving credits between countries and conditions for mandatory cancellations of credits, according to the latest draft Article 6.4 text from 17 November seen by Argus.

## USA-China

The crucial issue of loss and damage payments, as well as its energy transition funding burden, will weigh on China's green goals

China and the US have decided to resume climate talks, seen as crucial to managing climate change, after they collapsed following tensions over Taiwan in August. But China's climate challenges are growing, with the US and global leaders increasing pressure on it to cough up its fair share of climate reparations, on top of the already high costs associated with meeting its own climate goals.

UN chief Antonio Guterres called for a "climate solidarity pact" between developed and emerging economies at the Cop 27 UN climate conference in Sharm el-Sheikh, Egypt, with increasing demands for China to play its part. The inclusion of loss and damage in this year's summit was a milestone. The concept advocates rich countries, which have failed to deliver on $100bn promised annually for climate mitigation and adaptation, pay for the effects of climate change that cannot be avoided by mitigation, adaptation or other measures.

But countries are at odds on the issue, with differing views from the world's two biggest emitters — China and the US — threatening to derail their broader climate goals. The US is softening its resistance to paying developing countries. It is unclear if China will pay into any loss and damage fund, given its struggling economy, even though the US thinks it should.

China may struggle to meet its own climate goals. US climate envoy John Kerry has accused China of not doing enough to curb CO2 and other greenhouse gas emissions, even though it has deployed renewables and electric vehicles at a fast pace. To reach net zero emissions by 2060, China needs a staggering $14 trillion-17 trillion in additional investments for green infrastructure and technology in the power and transport sectors alone, with $2.1 trillion needed in the next decade to meet its nationally determined contribution (NDC) targets, according to a report by the World Bank last month.

"Given the immense price tag, public investments won't be sufficient to meet these needs, so China needs policy and regulatory reforms to spur the private sector and fully tap the potential for investment and innovation," the report said. China's renewable targets will also need to be 1,700GW by 2030 compared with 1,200GW in its current NDC, it said.

Despite emitting more than a third of global CO2 emissions last year, China is designated as a developing country by the World Trade Organisation. China emits over 10bn t/yr of CO2 but only has carbon capture, utilisation and storage (CCUS) to cut 3mn t/yr, according to figures from the environment ministry. Its CO2 emissions were 11.9bn t last year, the IEA says. To meet its climate pledges, China needs 20mn-408mn t of CCUS capacity by 2030, 600mn-1.45bn t by 2050 and 1bn-1.82bn t by 2060, according to a report by the ministry last year. The IEA estimates global CCUS costs to be from $15/t to as much as $120/t, depending on the CO2 source.


# Negotiation Updates

## Recent Climate Commitment

This report is based on the most recent entry submitted by the `r a.country` (`r a`).

```{r processing, echo=FALSE, results='hide'}

## Process Data (averages?)

### Select Most Recent Timestamp
names(df)

Block.means = df %>% 
  group_by(Block) %>% 
  summarise_at(c("Peak.Yr", "Reduction.Yr", "Reduction.Rate", "Deforestation", "Afforestation"),
               ~round(mean(., na.rm=T),0)) %>% as.data.frame()


names(df)
Block.mitigation = df %>% 
  group_by(Block) %>% 
  summarise_at(c("Billion"), ~round(sum(., na.rm=T), 0)) %>% as.data.frame()


Block.Results <- merge(Block.means, Block.mitigation)
Block.Results

## Normalize based on number of developing countries and population %
df$Country[df$Block=="Developing"]
df$Population <- NA
df$Population[df$Block=="Developing"] <- c(3.2, 210, 111, 13.2, 28)
# df$NormalizingPop <- 67000/df$Population 
df$BillionNormalized <- df$Billion
df$BillionNormalized[df$Block=="Developing"] <- 152/5 * df$Billion[df$Block=="Developing"]

mitigation <- df %>% 
  group_by(Mitigation) %>% 
  summarise_at(c("BillionNormalized"), sum, na.rm=T)

total = data.frame(Mitigation="Total", BillionNormalized=sum(mitigation$BillionNormalized))
if(total$BillionNormalized <0) total$Mitigation="Deficit"

mitigation = rbind(mitigation, total)

names(mitigation) = c("Mitigation Transfers", "Funds")

## Link to Climate Negotiation?


## Create Table
### C-Road Order
index = c(6, 3, 5, 1, 4, 2)

Block.Results$Index = index
Block.Results

str(Block.Results)
```

## Negotiation Results

Based on the submissions, the commitments are as as follows:

```{r results, echo=FALSE, results='asis'}

knitr::kable(Block.Results[index,][,-8], align = "lcclrrc", row.names = FALSE) %>%
  kable_styling(full_width = F)
#print(xtable(Block.Results[index,][,-8]), include.rownames=FALSE, type = "html")
```

These have values have been used to update the C-Roads Projections. 

## Projected Temperature Change

Based on our current commitments, we are looking at a 2.5 degree temperature change by the year 2100. The potential impacts for this remain extremely worrisome and states need to do more. 

[2.4 degree Projection -- Ghana Commitment Update ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2037&pc651=2043&pc652=5&pc653=19&pc654=7&g0=515-6r&g1=86)

### Previous Results

[2.5 degree Projection -- Philippines Commitment Update ](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2039&pc651=2049&pc652=3&pc653=21&pc654=9&g0=515-6r&g1=86)


[2.6 degree Projection USA Commitment Update](https://c-roads.climateinteractive.org/scenario.html?v=22.11.0&pc600=2022&pc601=2022&pc602=5&pc603=13&pc604=1&pc610=2022&pc611=2022&pc612=7&pc620=2030&pc621=2035&pc622=2&pc630=2035&pc631=2040&pc632=2&pc633=10&pc634=10&pc640=2040&pc641=2050&pc642=3&pc643=10&pc644=10&pc650=2051&pc651=2062&pc652=3&pc653=6&pc654=6&g0=501-6r&g1=86)

## Mitigation Funds -- Totals

The UNFCCC has estimated a need of 200 billion USD to fund the mitigation fund, and probably requiring continuous updating and the need grows. 

In addition to this, a new fund, "Losses and Damages", is needing funding. At this point, we are combining the two funds, which are normalized (we have 5 developing countries and their demand is being multiplied by 152/5 (152 developing countries/5 represented)).

The Philippines is asking for 1 Trillion USD! While it is true they are facing massive impacts, the Developed Countries should get a lot for their money!

The developed countries need to do more to address these massive economic impacts being faced in developing countries. 

But developing countries should give up more of their reliance on fossil fuels. India and China should be setting as role models.


```{r mitigation, echo=FALSE, results='asis'}
knitr::kable(mitigation, align = "lr", row.names = FALSE) %>%
  kable_styling(full_width = F)
#print(xtable(mitigation), include.rownames=FALSE, type = "html")
```
