---
title: "Draft Blog"
author: "Claire Manning"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This blog contains a r data collected from a weather station in McCall, Idaho. This data goes back to March, 1906. The data includes daily minimum and maximum temperatures at the station location on the given date.

```{r readdata, echo=FALSE}
#Reading Climate Data
filepath= "/home/CAMPUS/cemc2018/Climate_Change_Narratives/Student_Folders/Manning/manning_mccall_data.csv"
climate_data=read.csv(filepath)
#Change dataframe to readable dates
str(climate_data)
min(climate_data$TMAX)
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
