---
title: "Claire's Blog"
author: "Claire Manning"
date: "9/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This plot contains data collected from a weather station in McCall, Idaho. This data goes back to March, 1906. The data includes daily minimum and maximum temperatures at the station location on the given date.


```{r readdata, echo=FALSE}
# path to Data/FA20/ using marc's username...
#filepath = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Data/FA20/manning_mccall_data.csv"

filepath= "/home/CAMPUS/cemc2018/Climate_Change_Narratives/Student_Folders/Manning/manning_mccall_data.csv"

climate_data=read.csv(filepath)

str(climate_data)
min(climate_data$TMAX)
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")

```


## Plotting Data

This is the data plotted in graphs. These graphs just show the changes in daily maximum temperature which varies alot from the hot summers to the cold winters.

```{r plotingdailydata, }
plot(TMAX~NewDate, climate_data, ty='l') # This works!

TMAX.lm = lm(TMAX ~ NewDate, data= climate_data)
slopeintercept = coef(TMAX.lm)

plot(TMAX ~ NewDate, data= climate_data, las=1)
abline(slopeintercept, col="red", lwd=3)
       
       (TMAX.lm.sum = summary(lm(TMAX ~ NewDate, data=climate_data)))
```
