---
title: "Draft Regional Analysis "
author: "Claire Manning"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This blog contains a regional analysis of data collected from a weather station in McCall, Idaho. This data goes back to March, 1906. The data includes daily minimum and maximum temperatures at the station location on the given date.


```{r readdata, echo=FALSE}
# path to Data/FA20/ using marc's username...
#filepath = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Data/FA20/manning_mccall_data.csv"
filepath= "/home/CAMPUS/cemc2018/Climate_Change_Narratives/Student_Folders/Manning/manning_mccall_data.csv"
climate_data=read.csv(filepath)
#Change dataframe to readable dates
str(climate_data)
min(climate_data$TMAX)
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
```

## Plotting Data

This is the data plotted in graphs. These graphs just show the changes in daily maximum temperature which varies alot from the hot summers to the cold winters in Mccall. I changed the data from daily highs and lows to monthy means.

```{r plotingdailydata, echo=FALSE}
#Plot data
plot(TMAX~NewDate, climate_data, ty='l') 
lm(TMAX ~ NewDate, data=climate_data)
TMAX.lm = lm(TMAX ~ NewDate, data= climate_data)
slopeintercept = coef(TMAX.lm)
plot(TMAX ~ NewDate, data= climate_data, las=1)
abline(slopeintercept, col="red", lwd=3)
(TMAX.lm.sum = summary(lm(TMAX ~ NewDate, data=climate_data)))
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
#Claculating Monthly Mean
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX, ty='l')
#plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month=="05"], ty='l')
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],
ty='l', xlim=c(1950, 2020))
May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
summary(May.lm)
abline(coef(May.lm), col="red")
```

The P-value for the month of may is .53 which means that the results of the analysis are not statistically significant.

```{r plotingmin, echo=FALSE}
#Analyzing minimum daily temperatures
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
# Fixing the Format of Month and Year as numeric
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
# First I create a vector of months
Months = c("January", "February", "March", "April",
"May", "June", "July", "August", "September", "October",
"November", "December")
# Create a panel so I can see all the figures at
# once.
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMAXresult <- NA
for (i in 1:12) {
# plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month==i],
# ty='l')
plot(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH ==
i, ], ty = "l", las = 1, xlim = c(1940, 2020),
main = Months[i])
Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH ==
i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMAXresult <- rbind(TMAXresult, cbind(Months[i],
round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2,
4], 4), round(summary(Month.lm)$r.squared,
3)))
}
#Monthly average
plot(MonthlyTMINMean$TMIN, ty='l')
```
```{r}

```

