---
title: "<center>Portugal Forest Fire Susceptibility</center>"
author: "Brandon Tolentino"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---
```{r load-data, knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center", cache=TRUE)
library(readr)
#Rank <- read_csv("~/Wildfires_img/Rank.csv")
# Rank <- read.csv("/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/docs/F2018_Blogs/data/Forest Fire Blog Data.csv")
library(knitr)
library(rmdformats)
library(dplyr)
library(ggplot2)
data(simple)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```
<center>

[A massive wildfire is reflected in a stream in central Portugal on Sunday, June 18, 2017.](figure/Wildfires_img/fire.jpg)

</center> 

***
#*Background*

<center>

[Figure 1. The country of Portugal is located on the western-most coast of Europe ](/home/CAMPUS/btpm2015/Wildfires_img/whereportugal.PNG)

</center> 
##**Where In The World Are We?** 
###Bordered by the Atlantic Ocean on the west and Spain on the east, Portugal boasts a diverse climate and landscape composed of 92,090 km^2 of total land area, approximately the size of the U.S. state of Maine. A glimpse of the diversity of Portugal can be found in both figures 2 and 3. The former shows a fumarole, which is an opening near a volcano that expels gases, where the latter shows a highly flammable coastal ecosystem similar to those of California and Florida.
<center>

[Figure 2. Fumaroles along a waterway created by springs in central part of Furnas village on Sao Miguel Island, Azores, Portugal](/home/CAMPUS/btpm2015/Wildfires_img/Portugal1.PNG)

</center>

<center>

[Figure 3. Reminiscent of Southern California's Coastal Sage Scrub, a scenic Atlantic trail named Costa Verde attracts visitors year-round.](/home/CAMPUS/btpm2015/Wildfires_img/Portugal2.PNG)

</center>
##**Why Does It Matter?**
###Over 11 million people call this diverse landscape home. For this reason alone, studying natural disasters in this part of the world is especially pertinent. Furthermore, property value damages often surpass millions of dollars as well as reducing available funds for wildlife management and forest fire prevention in the future. 

##**What's Going On?** 
###In Portugal, wildfires larger than 100 ha in size make up less than 1% of fires but account for about 75% of the total area burned (DGRF, 2006)(Fig.4). 
<center>

[Figure 4. A typical Football field or the inside area of a Track field are approximately one hectare in size ](/home/CAMPUS/btpm2015/Wildfires_img/hectare.PNG)

</center>

##**Who Is Responsible?**
###Studies have shown that WUIs (Wildlife Urban Interfaces) are prone to fires because of more risks associated with human activities. Wildfires are not always started by humans though; many wildfires are started by lightning striking tall trees within the canopy. These types of fires are often less damaging than those caused by humans. Policymakers, firefighting personnel, government officials, and constituents themselves are responsible for the mitigation of forest fires in any region. 
***
#*Goal*
##Throughout this blog, I will be focusing on Portugal's varying forest ecosystems, and determining how they differ in terms of fire hazard and risks. After analyzing the data associated with the different forest types, recommendations can be made for how each forest type should be managed to best avoid future forest fires.

***
#*Climate* 

###Portugal can be characterized by the different ecosystems across the country. These diverse ecosystems differ in susceptibility to forest fires because they have different structures. That is, forests differ in height of overall canopy as well as the physical spacing between each individual plant (Fig. 5). These particular attributes are of importance because fires readily depend on the availability of biomass to spread. These factors, alongside the uncertainties of Climate Change, are making fires more common and much more dangerous.
<center>

[Figure 5. Forests can be better undestood in terms of their structure. Forests can have heavily concentrated plants or dispersed individuals as well as tall canopies versus short stocked plants.](/home/CAMPUS/btpm2015/Wildfires_img/rainforest.jpg)

</center> 
###Rainy winters are dispersed unevenly across the Mediterranean and hot, dry summers dominate for half of the year. Much like in Australia and California, wildfires thrive on the accessability of biomass following rainy seasons when the ecosystems are hot and dry once more (Fig. 6). The raging fires in the hills of Monchique show a striking juxtaposition between the quiet streets of the town and the dangers of wild fires. 

<center>

[Figure 6. Wildfires swept across Monchique in the south of Portugal today in the wake of a heatwave](/home/CAMPUS/btpm2015/Wildfires_img/Portugal3.PNG)

</center>
***
##*Data*
###Forest ecosystems in mainland Portugal were identified and described by Godinho-Ferreira et al. in 2005. This data will be used to identify plants and structure of forests in Portugal. The following section will elaborate on this data as well as its significance. 
***
#*Forest Types Across Portugal* 
###As previously mentioned, Godinho-Ferreira et al. (2005) actively recorded and analyzed the different forest types across Portugal. His team contributed to the Portuguese National Forest Inventory (NFI), and his team's forest data are used to roughly assess relative fire hazards across the country. Using this data alongside the observations and analysis performed by Fernandes (2009), a better understanding of the risks of fire across the Portuguese landscape is possible (Fig.7).
<center>

[Figure 7. Paulo Fernandes (2009) combines forest structure data and fuel modelling to better describe fire hazards across Portugal](/home/CAMPUS/btpm2015/Wildfires_img/article.PNG)

</center>
###Often, fuel load (i.e. available biomass) is the only factor that can be address when preventing wildfires(Agee and Skinner, 2005). Forest structures and fuel models can be used in unison to best identify which forests are most susceptible to forest fires. Thus, giving us an idea of what areas require the most attention and resources. 

###The fuel model used here, designed by Rothermel (1972), considers four aspects when determining surface fire propagation. The model looks at general availability of plant litter, herbs, shrubs, and the slashing of plants. This model has been edited and improved over time to obtain a more realistic estimate of fire characteristics for specific fuel types (Andrews et al. 2005)

###Forests across the globe can be physically characterized by their overall plant structures, species population densities, and concentration of individual species across the landscape. More generally, they each have a specific "forest structure". Forest structure is *vital* because it has an effect on wind speed and dead fuel moisture as well as immediately available biomass.

***
##Forest Structures
  * ###Open and Tall 
<center>

[Figure 8. Open and Tall forest composed of *Eucalyptus* trees](/home/CAMPUS/btpm2015/Wildfires_img/ot.jpg)

</center>
  * ###Open and Low
<center>

[Figure 9. Open and Low Forest composed of *Quercus suber*](/home/CAMPUS/btpm2015/Wildfires_img/ol.PNG)

</center>
  * ###Closed and Tall
<center>

[Figure 10. Closed and tall Forest made up predominantly by *Pinus Pinaster*](/home/CAMPUS/btpm2015/Wildfires_img/ct.PNG)

</center>
  * ###Closed and Low
<center>

[Figure 11. Closed and Low Forest composed of *P. pinaster* ](/home/CAMPUS/btpm2015/Wildfires_img/cl.PNG)

</center>
***
##Each of the forest types can be evaluated using three fire behavior descriptors: rate of spread, fireline intensity, and crown fire potential. 

##* Rate of Spread: How fast fire spreads in a specific area

##* Fireline Intensity: Rate of spread estimate x fuel load x heat content

##* Crown Fire Potential: Probability of forest fires spreading from treetop to treetop 

***
#*Fire Hazard by Forest Type* 
### By looking at the four different types of forest structure in relation to Fire Spread, Intensity, and Crowning, we get a better idea of which forest structures have the highest risk of seriously dangerous fires. 

###* Generally, both open and closed low structures have the highest risk of crowning, fire spread, as well as the highest fire intensity 

###* Of the tall systems, the closed structures have higher risks of crowning, fire spread, and higher fire intensity, but these values differ only slightly. 
```{r include=TRUE}
# qplot(x = Type, y = Spread, data = Rank)
# qplot(x = Type, y = Intensity, data = Rank)
# qplot(x = Type, y = Crowning, data = Rank)
```

###In conclusion, these figures suggest that flammability is more closely related to stand structure, rather than of cover type. Open structures have a lower understory layer and less fuel loading where closed structures tend to be more tightly packed and more capable of spreading fires quickly.

###Furthermore, fuel load is more important than rate of spread in regards to fireline intensity. Crown base height is more important than fireline intensity in determining variations in crowning potential for the different forest types. 
***
***
#*Future Directions* 
###* Stand thinning and pruning increases fire spread via changes in within-stand wind movement. 

###* Surface fuel treatments can favor fire suppression and decrease fire severity.

###* Vertical discontinuity is vital in avoiding crown fire. 

##With these main ideas in mind, the following can be done to prevent forest fires much more effectively:
###* Open and tall forest types can remain as is given that they have low to moderate surface fuel accumulaton as well as low fireline intensity. 

###* Closed and low forest types can have heterogenous fuel conditions, high crowning potential and high fireline potential due to well-developed shrub layers and high vertical continuities within the system.

###* Open and low forest types may be our best bet in mitigating forest fire effects. These types of forests have high spread and crowning due to heavy wind exposure and a large vertical continuity. These forest types exacerbate the effects of wild fires by hacing fuel loads in very flammable arrangements. 
***
##Future studies should continue to look at differences in forest structures across the globe to piece together the pieces that are missing in Portugal. Given that Portugal employs a similar climate to places like California, and Australia, it is a valuable case-study when looking into the future of forest fires around the world. This area of study is of special importance given that Climate Change is increasing the likelihood and severity of wild fires everywhere. 
***

#*Bibliography*
* Agee J.K. and Skinner C., 2005. Basic principles of forest     fuel reduction treatments. For. Ecol. Manage. 211: 83–96.
* Andrews P., Bevins C., and Seli R., 2005. BehavePlus Fire      Modeling System, version 3.0: user’s guide, USDA Forest      Service, Ogden, 134 p.
* DGRF, 2006. Incêndios orestais, relatório de 2005, Divisão     de   Defesa da Floresta Contra Incêndios, DGRF, Lisboa.
* Fernandes, Paulo M. "Combining Forest Structure Data and       Fuel Modelling to Classify Fire Hazard in Portugal."         Annals of Forest Science 66, no. 4 (2009): 415.
* Godinho-Ferreira P., Azevedo A., and Rego F., 2005. Carta      da   tipologiaorestal de Portugal Continental. Silva         Lusitana     13: 1–34. 
* Godinho-Ferreira P., Azevedo A., Vaz P., and Rego F., 2006.     Composition, conguration and vertical structure of           Portuguese forests: implications in wildre probability.      In:Viegas D.X. (Ed.), Proceedings V International            Conference on Forest Fire Research, Elsevier B.V.,           Amsterdam, CD-ROM.
* Rothermel R.C., 1972. A mathematical model for predicting     respread in wildland fuels, USDA Forest Service, Ogden, 40    p.