---
title: "Should Alabamians be Concerned About Climate Change?"
author: "Cheyenne Vance"
date: "3/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Should Alabamians be Concerned About Climate Change?

  According to a survey done by National Surveys on Energy and Environment in 2018, 60% of Americans believe that global warming is happening, and that “humans are at least partially responsible for the rising temperatures.” It was also found that only 35% of Republicans believe that climate change is partially a result of human actions, while 50% of Republicans do not believe climate change at all. (http://closup.umich.edu/files/ieep-nsee-2018-spring-climate-belief.pdf) I chose to include this information because my area of interest,  Alabama, is a red state, meaning most people in the state align with Republican ideas. 
Several politicians who are also climate skeptics hail from the south which is quite ironic considering the south is predicted to bear the brunt of climate change impacts. According to the Brookings Institute, Alabama will suffer the fifth highest economic loss, behind only Florida, Mississippi, Louisiana and Arkansas, all of which are red states. Aside from economic loss, the south will most likely see an increase in mortality, loss of coastal lands and homes, and decreased agricultural production. (Map 1)
In this blog I will be testing the null hypothesis: The climate is not changing in Alabama. If I am able to reject this hypothesis, that means that the climate is changing, and Alabamians should in fact be concerned about climate change. I will first analyze climate and precipitation trends in cities in North and South Alabama, and then provide some detailed information on the impacts of these trends, including sea level rise, economic damage, and flooding. 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
