---
title: "Upstate New York Weather Changes"
author: "Olivia Howie"
date: "12/12/16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing CSV Files

```{r importingCSV}
#Importing Scripts

file = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Olivia/856083.csv"

import = read.csv(file)
```

## Checking the Data

```{r checking}
plot(TMAX~DATE, import)
```

What's going on?  What is the deal with the -9999? These are used for missing data. We need to remove them!

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```

Okay, now we'll check again, for dates less then Dec. 31, 1913:
```{r}
plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```

Yikes! What's wrong? As it turns it the problem is that with how the dates are specified. In particular, the Dec 31 to Jan 1 transition: 

19132131 -> 19140101, if you use these as numbers it should be 19131232, 19131232, etc. You see we are missing lots numbers!


```{r}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

## Fixing the Dates

First, we convert the date a string or character values. Next, we'll convert the strings to a data format. 

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```
## Subset Sites

Olivia has two sites in here data, so, we need to subset it.

```{r}
unique(import$STATION_NAME)
```

Let's choose the POUGHKEEPSIE NY US because the record is longer than the airport.

```{r}
poughkeepsie <- subset(import, STATION_NAME=="POUGHKEEPSIE NY US", select=c(STATION, STATION_NAME, LONGITUDE, LATITUDE, ELEVATION, DATE, NewDate, TMIN, TMAX))


## A new plot
```{r}
plot(TMAX~NewDate, poughkeepsie, ty='l')
```

## Evaluating the trend

Creating a linear model and then evaluate the rate (slope). 
```{r}
# Linear Model
poughkeepsie.lm <- lm(TMAX~NewDate, data=poughkeepsie)
summary(poughkeepsie.lm)
plot(TMAX~NewDate, poughkeepsie, ty='l')
abline(coef(poughkeepsie.lm), col='red')
```

## Creating Month and Year Variables

So, let's figure out how to see how changes happen for individual months.
```{r}
# Get months
poughkeepsie$Month = months(poughkeepsie$NewDate)
poughkeepsie$Year = format(poughkeepsie$NewDate, format="%Y")
```

### Creating Monthly Means
```{r creatingmonthlymeans}
MonthlyMean = aggregate(TMAX ~ Month + Year, poughkeepsie, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
head(MonthlyMean)
```


```{r plottingMonthyMeans}
plot(MonthlyMean$TMAX, ty='l')
```

### Selecting for Each Month 

```{r January}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="January"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="January",], ty='l') 
January.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="January",])
summary(January.lm)

abline(coef(January.lm), col="red")

```

Now, the change is `r coef(January.lm)[2]` degress/year or `r round(coef(January.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(January.lm)$coefficients[2,4],4)`.

```{r February}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="February"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="February",], ty='l') 
February.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="February",])
summary(February.lm)

abline(coef(February.lm), col="red")

```

Now, the change is `r coef(February.lm)[2]` degress/year or `r round(coef(February.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(February.lm)$coefficients[2,4],4)`.

```{r March}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="March"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="March",], ty='l') 
March.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="March",])
summary(March.lm)

abline(coef(March.lm), col="red")

```

Now, the change is `r coef(March.lm)[2]` degress/year or `r round(coef(March.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(March.lm)$coefficients[2,4],4)`.

```{r April}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="April"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="April",], ty='l') 
April.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="April",])
summary(April.lm)

abline(coef(April.lm), col="red")

```

Now, the change is `r coef(January.lm)[2]` degress/year or `r round(coef(April.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(April.lm)$coefficients[2,4],4)`.

```{r May}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="May"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="May",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="May",])
summary(May.lm)

abline(coef(May.lm), col="red")

```{r June}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="June"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="June",], ty='l') 
June.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="June",])
summary(January.lm)

abline(coef(June.lm), col="red")

```

Now, the change is `r coef(June.lm)[2]` degress/year or `r round(coef(June.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(June.lm)$coefficients[2,4],4)`.

```{r July}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="July"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="July",], ty='l') 
July.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="July",])
summary(July.lm)

abline(coef(July.lm), col="red")

```

Now, the change is `r coef(July.lm)[2]` degress/year or `r round(coef(July.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(July.lm)$coefficients[2,4],4)`.

```{r August}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="August"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="August",], ty='l') 
August.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="August",])
summary(August.lm)

abline(coef(August.lm), col="red")

```

Now, the change is `r coef(August.lm)[2]` degress/year or `r round(coef(August.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(August.lm)$coefficients[2,4],4)`.

```{r September}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="September"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="September",], ty='l') 
September.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="September",])
summary(September.lm)

abline(coef(September.lm), col="red")

```

Now, the change is `r coef(September.lm)[2]` degress/year or `r round(coef(September.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(September.lm)$coefficients[2,4],4)`.

```{r October}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="October"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="October",], ty='l') 
October.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="October",])
summary(October.lm)

abline(coef(October.lm), col="red")

```

Now, the change is `r coef(October.lm)[2]` degress/year or `r round(coef(October.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(October.lm)$coefficients[2,4],4)`.

```{r November}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="November"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="November",], ty='l') 
November.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="November",])
summary(November.lm)

abline(coef(November.lm), col="red")

```

Now, the change is `r coef(November.lm)[2]` degress/year or `r round(coef(November.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(November.lm)$coefficients[2,4],4)`.

```{r December}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="December"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="December",], ty='l') 
December.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="December",])
summary(December.lm)

abline(coef(December.lm), col="red")

```

Now, the change is `r coef(December.lm)[2]` degress/year or `r round(coef(December.lm)[2]*100, 3)` degress/100 years with a probability of `r round(summary(December.lm)$coefficients[2,4],4)`.

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, poughkeepsie, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

2. Now we plot the mins, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="January"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="January",], ty='l') 
Jan.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="January",])
summary(Jan.lm)

abline(coef(Jan.lm), col="red")

```

We get a slope, `r coef(Jan.lm)[2]` degress/year and a probability of `r round(summary(Jan.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Jan.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="February"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="February",], ty='l') 
Feb.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="February",])
summary(Feb.lm)

abline(coef(Feb.lm), col="red")

```

We get a slope, `r coef(Feb.lm)[2]` degress/year and a probability of `r round(summary(Feb.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Feb.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="March"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="March",], ty='l') 
Mar.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="March",])
summary(Mar.lm)

abline(coef(Mar.lm), col="red")

```

We get a slope, `r coef(Mar.lm)[2]` degress/year and a probability of `r round(summary(Mar.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Mar.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="April"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="April",], ty='l') 
Apr.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="April",])
summary(Apr.lm)

abline(coef(Apr.lm), col="red")

```

We get a slope, `r coef(Apr.lm)[2]` degress/year and a probability of `r round(summary(Apr.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Apr.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="May"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="May",], ty='l') 
May.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="May",])
summary(May.lm)

abline(coef(May.lm), col="red")

```

We get a slope, `r coef(May.lm)[2]` degress/year and a probability of `r round(summary(May.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(May.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="June"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="June",], ty='l') 
Jun.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="June",])
summary(Jun.lm)

abline(coef(Jun.lm), col="red")

```

We get a slope, `r coef(Jun.lm)[2]` degress/year and a probability of `r round(summary(Jun.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Jun.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="July"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="July",], ty='l') 
Jul.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="July",])
summary(Jul.lm)

abline(coef(Jul.lm), col="red")

```

We get a slope, `r coef(Jul.lm)[2]` degress/year and a probability of `r round(summary(Jul.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Jul.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="August"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="August",], ty='l') 
Aug.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="August",])
summary(Aug.lm)

abline(coef(Aug.lm), col="red")

```

We get a slope, `r coef(Aug.lm)[2]` degress/year and a probability of `r round(summary(Aug.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Aug.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="September"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="September",], ty='l') 
Sep.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="September",])
summary(Sep.lm)

abline(coef(Sep.lm), col="red")

```

We get a slope, `r coef(Sep.lm)[2]` degress/year and a probability of `r round(summary(Sep.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Sep.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="October"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="October",], ty='l') 
Oct.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="October",])
summary(Oct.lm)

abline(coef(Oct.lm), col="red")

```

We get a slope, `r coef(Oct.lm)[2]` degress/year and a probability of `r round(summary(Oct.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Oct.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="November"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="November",], ty='l') 
Nov.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="November",])
summary(Nov.lm)

abline(coef(Nov.lm), col="red")

```

We get a slope, `r coef(Nov.lm)[2]` degress/year and a probability of `r round(summary(Nov.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Nov.lm)$r.squared, 3)`

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="December"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="December",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="December",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degress/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temerpature, that year is one, is quite problematic.

4. What we have not determined is the cause. So, be careful when you describe the results, cause and effect cannot be analyzed using this method.
