---
title: "Analyzing Increase in Floods in Sugar Land, Texas"
author: "Kirara Tsutsui"
date: "10/2/2017"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = FALSE)
climate_data <-read.csv("/home/CAMPUS/ktaa2016/Climate_Change_Narratives/Student_Folders/Kirara_Tsutsui/kirara_houston_data.csv")
#head(climate_data)


strDates <-as.character(climate_data$DATE) 
#head(strDates)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
#str(climate_data)

```

# Monthly TMIN Averages
``````{r, echo=FALSE}
climate.lm<- lm(formula = TMIN ~ NewDate, data = climate_data)

climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m") 
climate_data$Year = format(climate_data$NewDate, format="%Y")

MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year) 
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month) 
#str(MonthlyTMINMean)

#January not significant 0.878
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="01",], ty='l', xlim=c(1850, 2020), main="January") 
January.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="01",])
#summary(January.lm)
abline(coef(January.lm), col="purple")

#February not significant 0.0604
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="02",], ty='l', xlim=c(1850, 2020), main="February") 
February.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="02",])
#summary(February.lm)
abline(coef(February.lm), col="purple")

#March not significant 0.212
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="03",], ty='l', xlim=c(1850, 2020),main="March") 
March.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="03",])
#summary(March.lm)
abline(coef(March.lm), col="purple")

#April not significant 
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="04",], ty='l', xlim=c(1850, 2020), main="April") 
April.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="04",])
#summary(April.lm)
abline(coef(April.lm), col="purple")

#May significant
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="05",], ty='l', xlim=c(1850, 2020), main="May") 
May.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="05",])
#summary(May.lm)
abline(coef(May.lm), col="purple")

#June significant
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="06",], ty='l', xlim=c(1850, 2020), main="June") 
June.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="06",])
#summary(June.lm)
abline(coef(June.lm), col="purple")

#July significant 
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="07",], ty='l', xlim=c(1850, 2020), main="July") 
July.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="07",])
#summary(July.lm)
abline(coef(July.lm), col="purple")

```
#write about the significance of the min average temperature change of July

``````{r, echo=FALSE}

#August significant
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="08",], ty='l', xlim=c(1850, 2020), main="August") 
August.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="08",])
#summary(August.lm)
abline(coef(August.lm), col="purple")

#September significant 0.0017
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="09",], ty='l', xlim=c(1850, 2020), main="September") 
September.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="09",])
#summary(September.lm)
abline(coef(May.lm), col="purple")

#October slightly significant 0.0495
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="10",], ty='l', xlim=c(1850, 2020), main="October") 
October.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="10",])
#summary(October.lm)
abline(coef(October.lm), col="purple")

#November significant 0.00398
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="11",], ty='l', xlim=c(1850, 2020), main="November") 
November.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="11",])
#summary(November.lm)
abline(coef(November.lm), col="purple")

#December not significant 0.462
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="12",], ty='l', xlim=c(1850, 2020), main="December") 
December.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="12",])
#summary(December.lm)
abline(coef(December.lm), col="purple")


`````
#Use Min temp values for the summers to compare how increase in min temperatures have increased either the number or magnitutde (volatility) of hurricanes in Sugarland


#Precipitation

```{r, echo=TRUE}
#Yearly Average

TotalPPT<-aggregate(climate_data$PRCP,list(climate_data$Year), sum, na.rm=T)

#Create two variable groups

names(TotalPPT)=c("Year", "PRCP") 


TotalPPT$Date<-as.numeric(TotalPPT$Year)
plot(PRCP~Date, data=TotalPPT, ty="p", pch=20)

TotalPPT.lm <- lm(PRCP~Date, data=TotalPPT)
summary(TotalPPT.lm)
abline(coef(TotalPPT.lm), col="blue")

names(TotalPPT)=c("Year", "PRCP", "Month")

plot(PRCP~Date, data=TotalPPT[TotalPPT$Month=="07",], ty='p', pch=20, main = "July")
JulyPRCP.lm<-lm(PRCP~Month, data=TotalPPT[TotalPPT$Month=="07",])
summary(July.lm)
abline(coef(JulyPRCP.lm), col="blue")


``````````````
