---
title: "Meilinger_Sarah_Test Blog"
author: "Sarah Meilinger"
date: "9/12/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome To My Blog

Hello. I am going to be writing about climate data of Orlando, Florida.



This is an image of the NOAA station in Orlando, taken March 23rd, 1999. The objective of the photos was to document possible obtructions of the weather it is to measure.
![orlando](https://www.ncdc.noaa.gov/EdadsV2/api/ASOSStationPhotos/image/byhash/DE057C49AB3E1D5A61E2376C5ED947AA)


read.csv(/home/CAMPUS/smac2019/Climate_Change_Narratives/Student_Folders/Meilinger/Orlando stuff.R)



When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
