---
title: "AAaaa"
author: "Sarah Meilinger"
date: "10/18/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Hello! I looked at PM from a station in Oakland, CA. Here are my results.

```{r,  echo=FALSE, results='hide'}
SCZ2020.csv <- "/home/CAMPUS/smac2019/Climate_Change_Narratives/Student_Folders/Meilinger/Daily 2020 Oak.csv"
SCZ2019.csv <- "/home/CAMPUS/smac2019/Climate_Change_Narratives/Student_Folders/Meilinger/Daily 2019 Oak.csv"
SCZ2018.csv <- "/home/CAMPUS/smac2019/Climate_Change_Narratives/Student_Folders/Meilinger/Daily 2018 Oak.csv"
SCZ2017.csv <- "/home/CAMPUS/smac2019/Climate_Change_Narratives/Student_Folders/Meilinger/Daily 2017 Oak.csv"
SCZ2016.csv <- "/home/CAMPUS/smac2019/Climate_Change_Narratives/Student_Folders/Meilinger/Daily 2016 Oak.csv"

SCZ2019 = read.csv(SCZ2019.csv)
SCZ2020 = read.csv(SCZ2020.csv)
SCZ2018 = read.csv(SCZ2018.csv)
SCZ2017 = read.csv(SCZ2017.csv)
SCZ2016 = read.csv(SCZ2016.csv)

SCZ=rbind(SCZ2018, SCZ2019, SCZ2020, SCZ2017, SCZ2016)

Date.char = as.character(SCZ$Date)

Date.char = as.character(SCZ$Date)
# testing to make sure this works...
# as.Date(Date.char, format="%Y-%m-%d")
SCZ$Date = as.Date(Date.char, format="%m/%d/%Y")
#str(SCZ)
names(SCZ)

# [1] "Date"                           "Source"                        
# [3] "Site.ID"                        "POC"                           
# [5] "Daily.Mean.PM2.5.Concentration" "UNITS"                         
# [7] "DAILY_AQI_VALUE"                "Site.Name"                     
# [9] "DAILY_OBS_COUNT"                "PERCENT_COMPLETE"              
# [11] "AQS_PARAMETER_CODE"             "AQS_PARAMETER_DESC"            
# [13] "CBSA_CODE"                      "CBSA_NAME"                     
# [15] "STATE_CODE"                     "STATE"                         
# [17] "COUNTY_CODE"                    "COUNTY"                        
# [19] "SITE_LATITUDE"                  "SITE_LONGITUDE"  

#plotting
par(mfrow= c(1,2), cex.main=0.75, cex.lab=0.75, cex.axis=0.75, pin= c(1.5, 1))
min(SCZ$Daily.Mean.PM2.5.Concentration, na.rm=T)
max(SCZ$Daily.Mean.PM2.5.Concentration, na.rm=T)
plot(SCZ$Daily.Mean.PM2.5.Concentration ~ SCZ$Date, data= SCZ, ylim= c(1,40), xlab = "Year", ylab = "Daily PM 2.5", main = "Daily PM 2.5 Readings in Oakland Califonia, Zoomed In", mar= c(.5,.5,.5,.5))
abline(h=12, col="red")
plot(SCZ$Daily.Mean.PM2.5.Concentration ~ SCZ$Date, data= SCZ, xlab = "Year", ylab = "Daily PM 2.5", main = "Daily PM 2.5 Readings in Oakland Califonia")
abline(h=12, col="red")

# library(dplyr)
# library(lubridate)
# #str(SCZ)
# tmp1 <- mutate(SCZ, Date = ymd(Date),Year = year(Date), Month = month(Date))
# tmp2 <- group_by(tmp1, Month, Year)
# Monthly <- summarise(tmp2, result = mean(Daily.Mean.PM2.5.Concentration) )
# Monthly <- SCZ %>%
#   mutate(Date = ymd(Date),Year = year(Date), Month = month(Date)) %>%
#   group_by(Month) %>%
#   summarise(mean = mean(Daily.Mean.PM2.5.Concentration),
#             sd = sd(Daily.Mean.PM2.5.Concentration),
#             N = length(Daily.Mean.PM2.5.Concentration) )
# 
# qnorm(.975)
# ## [1] 1.959964
# Monthly$UCL95 = Monthly$mean + qnorm(.975)*Monthly$sd/sqrt(Monthly$N)
# Monthly$LCL95 = Monthly$mean - qnorm(.975)*Monthly$sd/sqrt(Monthly$N)
# 
# library(xtable)
# Monthly$Months
# 4
# ## Warning: Unknown or uninitialised column: `Months`.
# NULL
# print(xtable(Monthly), include.rownames=FALSE)

```

## A table of some things about the numbers

This is a very crude table of 
Month & mean & standard deviation & N & UCL95 & LCL95 

1.00 & 10.20 & 6.39 & 155 & 11.21 & 9.20 

2.00 & 8.25 & 4.95 & 142 & 9.07 & 7.44 

3.00 & 6.88 & 3.38 & 153 & 7.41 & 6.34  

4.00 & 7.95 & 3.94 & 150 & 8.58 & 7.32  

5.00 & 7.89 & 4.00 & 155 & 8.52 & 7.26  

6.00 & 8.86 & 4.14 & 150 & 9.52 & 8.20  

7.00 & 8.70 & 3.65 & 154 & 9.28 & 8.13  

8.00 & 11.52 & 7.47 & 155 & 12.70 & 10.35 

9.00 & 16.26 & 21.51 & 145 & 19.76 & 12.76  

10.00 & 11.29 & 9.35 & 131 & 12.89 & 9.69  

11.00 & 18.78 & 28.03 & 120 & 23.80 & 13.77  

12.00 & 10.60 & 7.88 & 124 & 11.99 & 9.22 

Some things of not are how both September and Novembers standard deviations are relatively higher than the other months, and how most charts have upper confidence limits of less than 15, excepting the month of September, and all lower confidence limits are below 15, and most below 10, excepting August, September, and November, at 10.35, 12.76, and 13.77 respectively.


## Health Good? No, Health Bad

<<<<<<< HEAD
This graph shows the air frequently has particle matter of smaller than 2.5 Âµm over the recommended rate of 12 Âµg/m3. As Wu et al says in *Exposure to air pollution and COVID-19 mortality in the United States: A nationwide cross-sectional study*, having "an  increase  of  1 ug/m3in  long-term  PM 2.5 exposure is associated with an 8% increase in the COVID-19 mortality rate" (Wu et al). Thus, any amount of additional PM2.5 above the recommended amount can have very negative effects on health.
=======
This graph shows the air frequently has particle matter of smaller than 2.5 Âµm over the recommended rate of 12 Âµg/m3. As Wu et al says in *Exposure to air pollution and COVID-19 mortality in the United States: A nationwide cross-sectional study*, having "an  increase  of  1 ðœ‡g/m3in  long-term  PM 2.5 exposure is associated with an 8% increase in the COVID-19 mortality rate" (Wu et al). Thus, any amount of additional PM2.5 above the recommended amount can have very negative effects on health.
>>>>>>> 555fd2d... ye



## Thank You
Sarah Meilinger