---
title: "The Effects of Anthropogenic Climate Change on Farming in Michigan"
author: "Lila Rubin"
date: "October 2020"
output: html_document
---

*The state of Michigan sits on land stolen from multiple Indigenous communities: the Ojibwa, in the Upper Peninsula and northern Lower Peninsula; the Potawatomi, in the southern, eastern, and central parts of the state; and the Odawa, in the western Lower Peninsula. Known as the Council of the Three Fires, the three nations were primarily farming communities, raising corn, beans, squash, and more. These nations’ populations have been decimated due to the genocide of the Three Fires at the hands of European settlers. While small Ojibwa, Potawatomi, and Odawa communities continue to rebuild and thrive in the state of Michigan, it would be unacceptable to ignore the destruction and genocide of Native nations carried out by the creators of this state. As we discuss modern agriculture, those of us who are settlers remember that this land is not ours to sow.*

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Coasts are ravaged by hurricanes, flooding, and wildfires. The central United States is plagued by drought, the South by heat waves. It is almost unanimously agreed in the scientific community that these drastic and deadly changes in climate are due to human-caused global warming (Cook, John, et al). In the agricultural cradle of the U.S. known as the Midwest, however, citizens are seeing the effects of climate change not only in the need for more deodorant as a result of hotter summers. Production of staple crops like corn, wheat, beans, and squash, along with the export and uses of water from the Great Lakes (tourism, fishing, and transport, among others), are all projected to falter or are already in the process of declining under the increasingly warm and unpredictable world. This blog examines and analyzes the impact of climate statistics and projections on the Michigan agricultural industry. We can conclude that the “Great Lakes State,” with its farmers, university students, fishermen, and urban dwellers, may be overwhelmed by the impacts of climate change on its agricultural, ecological, and economical success.


```{r climate_data, echo=FALSE, eval=FALSE}
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
plot(TMAX~NewDate, climate_data, ty='l', main="Daily Temperatures", ylab= "Max Temperature (C)", xlab="Year")
bestfitline.lm <- lm(TMAX~NewDate, climate_data)
abline(coef(bestfitline.lm), col="red")
```
###### Data Retrieval and Analysis

Statistical support in the form of graphs for this hypothesis are based on data obtained from the National Oceanic and Atmospheric Administration (NOAA), which was collected at the University of Michigan in Ann Arbor. After retrieving such data from the NOAA, it was processed through the coding interface RStudio to create summaries and graphs of the data points. Close analysis of trends and patterns in this information, along with the review of other research done in the region, informed the conclusions which will be presented in this blog. It should be noted that we will be focusing on Ann Arbor data to paint a general picture of weather anomalies in the whole state of Michigan.

```{r TMAX}
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

# January
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",],
ty='l', xlim=c(1900, 2020), main="January", ylab="Max Temperature (C)", xlab="Year")

January.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",])
abline(coef(January.lm), col="red")

# February
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="02",],
ty='l', xlim=c(1900, 2020), main="February", ylab="Max Temperature (C)", xlab="Year")

February.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="02",])
abline(coef(February.lm), col="red")

# March
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="03",],
ty='l', xlim=c(1900, 2020), main="March", ylab="Max Temperature (C)", xlab="Year")

March.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="03",])
abline(coef(March.lm), col="red")

# April
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",],
ty='l', xlim=c(1900, 2020), main="April", ylab="Max Temperature (C)", xlab="Year")

April.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",])
abline(coef(April.lm), col="red")

# May
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],
ty='l', xlim=c(1900, 2020), main="May", ylab="Max Temperature (C)", xlab="Year")

May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
abline(coef(May.lm), col="red")

# June
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="06",],
ty='l', xlim=c(1900, 2020), main="June", ylab="Max Temperature (C)", xlab="Year")

June.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="06",])
abline(coef(June.lm), col="red")

# July
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="07",],
ty='l', xlim=c(1900, 2020), main="July", ylab="Max Temperature (C)", xlab="Year")

July.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="07",])
abline(coef(July.lm), col="red")

# August
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",],
ty='l', xlim=c(1900, 2020), main="August", ylab="Max Temperature (C)", xlab="Year")

August.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",])
abline(coef(August.lm), col="red")

# September
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",],
ty='l', xlim=c(1900, 2020), main="September", ylab="Max Temperature (C)", xlab="Year")

September.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",])
abline(coef(September.lm), col="red")

# October
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",],
ty='l', xlim=c(1900, 2020), main="October", ylab="Max Temperature (C)", xlab="Year")

October.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",])
abline(coef(October.lm), col="red")

# November
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",],
ty='l', xlim=c(1900, 2020), main="November", ylab="Max Temperature (C)", xlab="Year")

November.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",])
abline(coef(November.lm), col="red")

# December
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",],
ty='l', xlim=c(1900, 2020), main="December", ylab="Max Temperature (C)", xlab="Year")

December.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",])
abline(coef(December.lm), col="red")
```

In the few months where the maximum temperature has not experienced a significant increase in average temperatures, the climate is still changing. The average *minimum* temperature of winters is rising quickly.


```{r TMIN}

# changing date formatting
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

# temp business
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)

# January
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="01",], ty="l", xlim=c(1891, 2020), main="January", ylab="Min Temperature (C)", xlab="Year")

JanuaryTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="01",])
abline(coef(JanuaryTMIN.lm))

# February
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="02",], ty="l", xlim=c(1891, 2020), main="February", ylab="Min Temperature (C)", xlab="Year")

FebruaryTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="02",])
abline(coef(FebruaryTMIN.lm))

# March
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="03",], ty="l", xlim=c(1891, 2020), main="March", ylab="Min Temperature (C)", xlab="Year")

MarchTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="03",])
abline(coef(MarchTMIN.lm))

# April
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="04",], ty="l", xlim=c(1891, 2020), main="April", ylab="Min Temperature (C)", xlab="Year")

AprilTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="04",])
abline(coef(AprilTMIN.lm))

# May
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="05",], ty="l", xlim=c(1891, 2020), main="May", ylab="Min Temperature (C)", xlab="Year")

MayTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="05",])
abline(coef(MayTMIN.lm))

# June
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="06",], ty="l", xlim=c(1891, 2020), main="June", ylab="Min Temperature (C)", xlab="Year")

JuneTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="06",])
abline(coef(JuneTMIN.lm))

# July
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="07",], ty="l", xlim=c(1891, 2020), main="July", ylab="Min Temperature (C)", xlab="Year")

JulyTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="07",])
abline(coef(JulyTMIN.lm))

# August
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="08",], ty="l", xlim=c(1891, 2020), main="August", ylab="Min Temperature (C)", xlab="Year")

AugustTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="08",])
abline(coef(AugustTMIN.lm))

# September
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="09",], ty="l", xlim=c(1891, 2020), main="September", ylab="Min Temperature (C)", xlab="Year")

SeptemberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="09",])
abline(coef(SeptemberTMIN.lm))

# October
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="10",], ty="l", xlim=c(1891, 2020), main="October", ylab="Min Temperature (C)", xlab="Year")

OctoberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="10",])
abline(coef(OctoberTMIN.lm))

# November
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="11",], ty="l", xlim=c(1891, 2020), main="November", ylab="Min Temperature (C)", xlab="Year")

NovemberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="11",])
abline(coef(NovemberTMIN.lm))

# December
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="12",], ty="l", xlim=c(1891, 2020), main="December", ylab="Min Temperature (C)", xlab="Year")

DecemberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="12",])
abline(coef(DecemberTMIN.lm))
```
```{r PRCP_total}
# changing date formatting
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

# PRCP business
MonthlyPRCPSum = aggregate(PRCP ~ Month + Year, climate_data, sum)
MonthlyPRCPSum$YEAR = as.numeric(MonthlyPRCPSum$Year)
MonthlyPRCPSum$MONTH = as.numeric(MonthlyPRCPSum$Month)

# January
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="01",], ty="l", xlim=c(1891, 2020), main="January", ylab="Total Monthly Precipitation", xlab="Year")

JanuaryPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="01",])
abline(coef(JanuaryPRCP.lm))

# February
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="02",], ty="l", xlim=c(1891, 2020), main="February", ylab="Total Monthly Precipitation", xlab="Year")

FebPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="02",])
abline(coef(FebPRCP.lm))

# March
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="03",], ty="l", xlim=c(1891, 2020), main="March", ylab="Total Monthly Precipitation", xlab="Year")

MarPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="03",])
abline(coef(MarPRCP.lm))

# April
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="04",], ty="l", xlim=c(1891, 2020), main="April", ylab="Total Monthly Precipitation", xlab="Year")

AprPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="04",])
abline(coef(AprPRCP.lm))

# May
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="05",], ty="l", xlim=c(1891, 2020), main="May", ylab="Total Monthly Precipitation", xlab="Year")

MayPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="04",])
abline(coef(MayPRCP.lm))

# June
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="06",], ty="l", xlim=c(1891, 2020), main="June", ylab="Total Monthly Precipitation", xlab="Year")

JunePRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="06",])
abline(coef(JunePRCP.lm))

# July
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="07",], ty="l", xlim=c(1891, 2020), main="July", ylab="Total Monthly Precipitation", xlab="Year")

JulyPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="07",])
abline(coef(JulyPRCP.lm))

# August
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="08",], ty="l", xlim=c(1891, 2020), main="August", ylab="Total Monthly Precipitation", xlab="Year")

AugustPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="08",])
abline(coef(AugustPRCP.lm))

# September
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="09",], ty="l", xlim=c(1891, 2020), main="September", ylab="Total Monthly Precipitation", xlab="Year")

SeptPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="09",])
abline(coef(SeptPRCP.lm))

# October
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="10",], ty="l", xlim=c(1891, 2020), main="October", ylab="Total Monthly Precipitation", xlab="Year")

OctPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="10",])
abline(coef(OctPRCP.lm))

# November
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="11",], ty="l", xlim=c(1891, 2020), main="November", ylab="Total Monthly Precipitation", xlab="Year")

NovPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="11",])
abline(coef(NovPRCP.lm))

# December
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="12",], ty="l", xlim=c(1891, 2020), main="December", ylab="Total Monthly Precipitation", xlab="Year")

DecPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="12",])
abline(coef(DecPRCP.lm))
```

```{r SNOW}
# changing date formatting
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

# SNOW business
MonthlySNOWSum = aggregate(SNOW ~ Month + Year, climate_data, sum)
MonthlySNOWSum$YEAR = as.numeric(MonthlySNOWSum$Year)
MonthlySNOWSum$MONTH = as.numeric(MonthlySNOWSum$Month)

# January
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="01",], ty="l", xlim=c(1891, 2020), main="January", ylab="Total Monthly Snowfall", xlab="Year")

JanuarySNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="01",])
abline(coef(JanuarySNOW.lm))
summary(JanuarySNOW.lm)

# February
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="02",], ty="l", xlim=c(1891, 2020), main="February", ylab="Total Monthly Snowfall", xlab="Year")

FebSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="02",])
abline(coef(FebSNOW.lm))
summary(FebSNOW.lm)

# March
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="03",], ty="l", xlim=c(1891, 2020), main="March", ylab="Total Monthly Snowfall", xlab="Year")

MarSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="03",])
abline(coef(MarSNOW.lm))
summary(MarSNOW.lm)

# April
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="04",], ty="l", xlim=c(1891, 2020), main="April", ylab="Total Monthly Snowfall", xlab="Year")

AprSnw.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="04",])
abline(coef(AprSnw.lm))
summary(AprSnw.lm)

# May
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="05",], ty="l", xlim=c(1891, 2020), main="May", ylab="Total Monthly Snowfall", xlab="Year")

MaySNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="05",])
abline(coef(MaySNOW.lm))
summary(MaySNOW.lm)

# June
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="06",], ty="l", xlim=c(1891, 2020), main="June", ylab="Total Monthly Snowfall", xlab="Year")

JuneSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="06",])
abline(coef(JuneSNOW.lm))
summary(JuneSNOW.lm)

# July
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="07",], ty="l", xlim=c(1891, 2020), main="July", ylab="Total Monthly Snowfall", xlab="Year")

JulSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="07",])
abline(coef(JulSNOW.lm))
summary(JulSNOW.lm)

# August
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="08",], ty="l", xlim=c(1891, 2020), main="August", ylab="Total Monthly Snowfall", xlab="Year")

AugSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="08",])
abline(coef(AugSNOW.lm))
summary(AugSNOW.lm)

# September
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="09",], ty="l", xlim=c(1891, 2020), main="September", ylab="Total Monthly Snowfall", xlab="Year")

SeptSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="09",])
abline(coef(SeptSNOW.lm))
summary(SeptSNOW.lm)

# October
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="10",], ty="l", xlim=c(1891, 2020), main="October", ylab="Total Monthly Snowfall", xlab="Year")

OctSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="10",])
abline(coef(OctSNOW.lm))
summary(OctSNOW.lm)

# November
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="11",], ty="l", xlim=c(1891, 2020), main="November", ylab="Total Monthly Snowfall", xlab="Year")

NovSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="11",])
abline(coef(NovSNOW.lm))
summary(NovSNOW.lm)

# December
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="12",], ty="l", xlim=c(1891, 2020), main="December", ylab="Total Monthly Snowfall", xlab="Year")

DecSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="12",])
abline(coef(DecSNOW.lm))
```

