---
title: "Temperature Trends in Ann Arbor, MI"
author: "Lila Rubin"
date: "9/5/2020"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This blog will examine data collected regarding temperature and precipitation in Ann Arbor, Michigan from 1891 to the present day. After displaying and analyzing the data, we will discuss what conclusions may be drawn from it regarding the area's climate trends over time.


```{r climate_data, echo=FALSE, eval=FALSE}
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
plot(TMAX~NewDate, climate_data, ty='l', main="Daily Temperatures", ylab= "Max Temperature (C)", xlab="Year")
bestfitline.lm <- lm(TMAX~NewDate, climate_data)
abline(coef(bestfitline.lm), col="red")
```
## Monthly Maximum Temperatures over Time

*Research effect of temperature on crops*
*Graph precipitation levels*

```{r monthly}
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

# January
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",],
ty='l', xlim=c(1900, 2020), main="January", ylab="Max Temperature (C)", xlab="Year")

January.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",])
abline(coef(January.lm), col="red")

# February
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="02",],
ty='l', xlim=c(1900, 2020), main="February", ylab="Max Temperature (C)", xlab="Year")

February.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="02",])
abline(coef(February.lm), col="red")

# March
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="03",],
ty='l', xlim=c(1900, 2020), main="March", ylab="Max Temperature (C)", xlab="Year")

March.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="03",])
abline(coef(March.lm), col="red")

# April
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",],
ty='l', xlim=c(1900, 2020), main="April", ylab="Max Temperature (C)", xlab="Year")

April.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",])
abline(coef(April.lm), col="red")

# May
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],
ty='l', xlim=c(1900, 2020), main="May", ylab="Max Temperature (C)", xlab="Year")

May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
abline(coef(May.lm), col="red")

# June
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="06",],
ty='l', xlim=c(1900, 2020), main="June", ylab="Max Temperature (C)", xlab="Year")

June.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="06",])
abline(coef(June.lm), col="red")

# July
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="07",],
ty='l', xlim=c(1900, 2020), main="July", ylab="Max Temperature (C)", xlab="Year")

July.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="07",])
abline(coef(July.lm), col="red")

# August
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",],
ty='l', xlim=c(1900, 2020), main="August", ylab="Max Temperature (C)", xlab="Year")

August.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",])
abline(coef(August.lm), col="red")

# September
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",],
ty='l', xlim=c(1900, 2020), main="September", ylab="Max Temperature (C)", xlab="Year")

September.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",])
abline(coef(September.lm), col="red")

# October
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",],
ty='l', xlim=c(1900, 2020), main="October", ylab="Max Temperature (C)", xlab="Year")

October.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",])
abline(coef(October.lm), col="red")

# November
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",],
ty='l', xlim=c(1900, 2020), main="November", ylab="Max Temperature (C)", xlab="Year")

November.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",])
abline(coef(November.lm), col="red")

# December
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",],
ty='l', xlim=c(1900, 2020), main="December", ylab="Max Temperature (C)", xlab="Year")

December.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",])
abline(coef(December.lm), col="red")
```

