---
title: "The Effects of Temperature Rise on Farming in Michigan"
author: "Lila Rubin"
date: "9/5/2020"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This blog will examine data collected regarding maximum and mininimum temperatures in Ann Arbor, Michigan from 1891 to the present day. After displaying and analyzing the data, we will discuss what effect our conclusiosn may have on farming in the surrounding area.


```{r climate_data, echo=FALSE, eval=FALSE}
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
plot(TMAX~NewDate, climate_data, ty='l', main="Daily Temperatures", ylab= "Max Temperature (C)", xlab="Year")
bestfitline.lm <- lm(TMAX~NewDate, climate_data)
abline(coef(bestfitline.lm), col="red")
```
## Monthly Maximum Temperatures over Time

*Research effect of temperature on crops*

```{r TMAX}
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

# January
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",],
ty='l', xlim=c(1900, 2020), main="January", ylab="Max Temperature (C)", xlab="Year")

January.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",])
abline(coef(January.lm), col="red")

# February
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="02",],
ty='l', xlim=c(1900, 2020), main="February", ylab="Max Temperature (C)", xlab="Year")

February.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="02",])
abline(coef(February.lm), col="red")

# March
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="03",],
ty='l', xlim=c(1900, 2020), main="March", ylab="Max Temperature (C)", xlab="Year")

March.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="03",])
abline(coef(March.lm), col="red")

# April
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",],
ty='l', xlim=c(1900, 2020), main="April", ylab="Max Temperature (C)", xlab="Year")

April.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",])
abline(coef(April.lm), col="red")

# May
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],
ty='l', xlim=c(1900, 2020), main="May", ylab="Max Temperature (C)", xlab="Year")

May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
abline(coef(May.lm), col="red")

# June
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="06",],
ty='l', xlim=c(1900, 2020), main="June", ylab="Max Temperature (C)", xlab="Year")

June.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="06",])
abline(coef(June.lm), col="red")

# July
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="07",],
ty='l', xlim=c(1900, 2020), main="July", ylab="Max Temperature (C)", xlab="Year")

July.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="07",])
abline(coef(July.lm), col="red")

# August
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",],
ty='l', xlim=c(1900, 2020), main="August", ylab="Max Temperature (C)", xlab="Year")

August.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",])
abline(coef(August.lm), col="red")

# September
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",],
ty='l', xlim=c(1900, 2020), main="September", ylab="Max Temperature (C)", xlab="Year")

September.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",])
abline(coef(September.lm), col="red")

# October
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",],
ty='l', xlim=c(1900, 2020), main="October", ylab="Max Temperature (C)", xlab="Year")

October.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",])
abline(coef(October.lm), col="red")

# November
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",],
ty='l', xlim=c(1900, 2020), main="November", ylab="Max Temperature (C)", xlab="Year")

November.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",])
abline(coef(November.lm), col="red")

# December
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",],
ty='l', xlim=c(1900, 2020), main="December", ylab="Max Temperature (C)", xlab="Year")

December.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",])
abline(coef(December.lm), col="red")
```

In the few months where the maximum temperature has not experienced a significant increase in average temperatures, the climate is still changing. The average *minimum* temperature of winters is rising quickly.


```{r TMIN}

# changing date formatting
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

# temp business
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)

# January
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="01",], ty="l", xlim=c(1891, 2020), main="January", ylab="Min Temperature (C)", xlab="Year")

JanuaryTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="01",])
abline(coef(JanuaryTMIN.lm))
summary(JanuaryTMIN.lm)

# February
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="02",], ty="l", xlim=c(1891, 2020), main="February", ylab="Min Temperature (C)", xlab="Year")

FebruaryTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="02",])
abline(coef(FebruaryTMIN.lm))
summary(FebruaryTMIN.lm)

# March
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="03",], ty="l", xlim=c(1891, 2020), main="March", ylab="Min Temperature (C)", xlab="Year")

MarchTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="03",])
abline(coef(MarchTMIN.lm))
summary(MarchTMIN.lm)

# April
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="04",], ty="l", xlim=c(1891, 2020), main="April", ylab="Min Temperature (C)", xlab="Year")

AprilTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="04",])
abline(coef(AprilTMIN.lm))
summary(AprilTMIN.lm)

# May
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="05",], ty="l", xlim=c(1891, 2020), main="May", ylab="Min Temperature (C)", xlab="Year")

MayTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="05",])
abline(coef(MayTMIN.lm))
summary(MayTMIN.lm)

# June
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="06",], ty="l", xlim=c(1891, 2020), main="June", ylab="Min Temperature (C)", xlab="Year")

JuneTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="06",])
abline(coef(JuneTMIN.lm))
summary(JuneTMIN.lm)

# July
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="07",], ty="l", xlim=c(1891, 2020), main="July", ylab="Min Temperature (C)", xlab="Year")

JulyTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="07",])
abline(coef(JulyTMIN.lm))
summary(JulyTMIN.lm)

# August
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="08",], ty="l", xlim=c(1891, 2020), main="August", ylab="Min Temperature (C)", xlab="Year")

AugustTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="08",])
abline(coef(AugustTMIN.lm))
summary(AugustTMIN.lm)

# September
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="09",], ty="l", xlim=c(1891, 2020), main="September", ylab="Min Temperature (C)", xlab="Year")

SeptemberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="09",])
abline(coef(SeptemberTMIN.lm))
summary(SeptemberTMIN.lm)

# October
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="10",], ty="l", xlim=c(1891, 2020), main="October", ylab="Min Temperature (C)", xlab="Year")

OctoberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="10",])
abline(coef(OctoberTMIN.lm))
summary(OctoberTMIN.lm)

# November
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="11",], ty="l", xlim=c(1891, 2020), main="November", ylab="Min Temperature (C)", xlab="Year")

NovemberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="11",])
abline(coef(NovemberTMIN.lm))
summary(NovemberTMIN.lm)

# December
plot(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="12",], ty="l", xlim=c(1891, 2020), main="December", ylab="Min Temperature (C)", xlab="Year")

DecemberTMIN.lm <- lm(TMIN~YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month=="12",])
abline(coef(DecemberTMIN.lm))
summary(DecemberTMIN.lm)
```
```{r PRCP total}
# changing date formatting
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

# PRCP business
MonthlyPRCPSum = aggregate(PRCP ~ Month + Year, climate_data, sum)
MonthlyPRCPSum$YEAR = as.numeric(MonthlyPRCPSum$Year)
MonthlyPRCPSum$MONTH = as.numeric(MonthlyPRCPSum$Month)

# January
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="01",], ty="l", xlim=c(1891, 2020), main="January", ylab="Total Monthly Precipitation", xlab="Year")

JanuaryPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="01",])
abline(coef(JanuaryPRCP.lm))
summary(JanuaryPRCP.lm)

# February
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="02",], ty="l", xlim=c(1891, 2020), main="February", ylab="Total Monthly Precipitation", xlab="Year")

FebPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="02",])
abline(coef(FebPRCP.lm))
summary(FebPRCP.lm)

# March
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="03",], ty="l", xlim=c(1891, 2020), main="March", ylab="Total Monthly Precipitation", xlab="Year")

MarPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="03",])
abline(coef(MarPRCP.lm))
summary(MarPRCP.lm)

# April
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="04",], ty="l", xlim=c(1891, 2020), main="April", ylab="Total Monthly Precipitation", xlab="Year")

AprPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="04",])
abline(coef(AprPRCP.lm))
summary(AprPRCP.lm)

# May
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="05",], ty="l", xlim=c(1891, 2020), main="May", ylab="Total Monthly Precipitation", xlab="Year")

MayPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="04",])
abline(coef(MayPRCP.lm))
summary(MayPRCP.lm)

# June
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="06",], ty="l", xlim=c(1891, 2020), main="June", ylab="Total Monthly Precipitation", xlab="Year")

JunePRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="06",])
abline(coef(JunePRCP.lm))
summary(JunePRCP.lm)

# July
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="07",], ty="l", xlim=c(1891, 2020), main="July", ylab="Total Monthly Precipitation", xlab="Year")

JulyPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="07",])
abline(coef(JulyPRCP.lm))
summary(JulyPRCP.lm)

# August
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="08",], ty="l", xlim=c(1891, 2020), main="August", ylab="Total Monthly Precipitation", xlab="Year")

AugustPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="08",])
abline(coef(AugustPRCP.lm))
summary(AugustPRCP.lm)

# September
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="09",], ty="l", xlim=c(1891, 2020), main="September", ylab="Total Monthly Precipitation", xlab="Year")

SeptPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="09",])
abline(coef(SeptPRCP.lm))
summary(SeptPRCP.lm)

# October
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="10",], ty="l", xlim=c(1891, 2020), main="October", ylab="Total Monthly Precipitation", xlab="Year")

OctPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="10",])
abline(coef(OctPRCP.lm))
summary(OctPRCP.lm)

# November
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="11",], ty="l", xlim=c(1891, 2020), main="November", ylab="Total Monthly Precipitation", xlab="Year")

NovPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="11",])
abline(coef(NovPRCP.lm))
summary(NovPRCP.lm)

# December
plot(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="12",], ty="l", xlim=c(1891, 2020), main="December", ylab="Total Monthly Precipitation", xlab="Year")

DecPRCP.lm <- lm(PRCP~YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$Month=="12",])
abline(coef(DecPRCP.lm))
summary(DecPRCP.lm)
```

```{r SNOW}
# changing date formatting
climate_data <-read.csv("/home/CAMPUS/lraa2020/Climate_Change_Narratives/Student_Folders/Rubin/lila_annarbor_temp_and_precip.csv")
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

# PRCP business
MonthlySNOWSum = aggregate(SNOW ~ Month + Year, climate_data, sum)
MonthlySNOWSum$YEAR = as.numeric(MonthlySNOWSum$Year)
MonthlySNOWSum$MONTH = as.numeric(MonthlySNOWSum$Month)

# January
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="01",], ty="l", xlim=c(1891, 2020), main="January", ylab="Total Monthly Snowfall", xlab="Year")

JanuarySNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="01",])
abline(coef(JanuarySNOW.lm))
summary(JanuarySNOW.lm)

# February
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="02",], ty="l", xlim=c(1891, 2020), main="February", ylab="Total Monthly Snowfall", xlab="Year")

FebSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="02",])
abline(coef(FebSNOW.lm))
summary(FebSNOW.lm)

# March
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="03",], ty="l", xlim=c(1891, 2020), main="March", ylab="Total Monthly Snowfall", xlab="Year")

MarSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="03",])
abline(coef(MarSNOW.lm))
summary(MarSNOW.lm)

# April
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="04",], ty="l", xlim=c(1891, 2020), main="April", ylab="Total Monthly Snowfall", xlab="Year")

AprSnw.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="04",])
abline(coef(AprSnw.lm))
summary(AprSnw.lm)

# May
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="05",], ty="l", xlim=c(1891, 2020), main="May", ylab="Total Monthly Snowfall", xlab="Year")

MaySNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="05",])
abline(coef(MaySNOW.lm))
summary(MaySNOW.lm)

# June
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="06",], ty="l", xlim=c(1891, 2020), main="June", ylab="Total Monthly Snowfall", xlab="Year")

JuneSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="06",])
abline(coef(JuneSNOW.lm))
summary(JuneSNOW.lm)

# July
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="07",], ty="l", xlim=c(1891, 2020), main="July", ylab="Total Monthly Snowfall", xlab="Year")

JulSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="07",])
abline(coef(JulSNOW.lm))
summary(JulSNOW.lm)

# August
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="08",], ty="l", xlim=c(1891, 2020), main="August", ylab="Total Monthly Snowfall", xlab="Year")

AugSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="08",])
abline(coef(AugSNOW.lm))
summary(AugSNOW.lm)

# September
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="09",], ty="l", xlim=c(1891, 2020), main="September", ylab="Total Monthly Snowfall", xlab="Year")

SeptSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="09",])
abline(coef(SeptSNOW.lm))
summary(SeptSNOW.lm)

# October
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="10",], ty="l", xlim=c(1891, 2020), main="October", ylab="Total Monthly Snowfall", xlab="Year")

OctSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="10",])
abline(coef(OctSNOW.lm))
summary(OctSNOW.lm)

# November
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="11",], ty="l", xlim=c(1891, 2020), main="November", ylab="Total Monthly Snowfall", xlab="Year")

NovSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="11",])
abline(coef(NovSNOW.lm))
summary(NovSNOW.lm)

# December
plot(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="12",], ty="l", xlim=c(1891, 2020), main="December", ylab="Total Monthly Snowfall", xlab="Year")

DecSNOW.lm <- lm(SNOW~YEAR, data = MonthlySNOWSum[MonthlySNOWSum$Month=="12",])
abline(coef(DecSNOW.lm))
summary(DecSNOW.lm)
```

