---
title: "San Diego_45393"
author: "45393"
date: "9/17/2020"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



```{r, fig.show='hide',results= 'hide'}

## Data Sources
file.path="/home/CAMPUS/vipa2017/Github/Climate_Change_Narratives/Student_Folders/Plancarte/Plancarte_SanDiego_data.csv"
#file.path="/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Student_Folders/Plancarte/Plancarte_SanDiego_data.csv"
climate_data=read.csv(file.path)
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
plot(TMAX~NewDate, climate_data,ty='l')
bestfitline.lm=lm(TMAX~NewDate,data=climate_data)
coef(bestfitline.lm)
abline(coef(bestfitline.lm),col="red", lwd=3)
(TMAX.lm.sum=summary(lm(TMAX~NewDate,data=climate_data)))
#Although the slope is decreasing at a rate of -2.516x10^-5. We cannot reject the null hypothesis which states that there is no change in temperature, since p-value is 0.3267 and greater than 0.05. Therefore we cannot conclude that temperatures are decreasing at a rate of 0.0025 degrees celcius per 100 years

#R-squared estimate was <3x10^-5 (why is r-squared so low &what does it mean)
```




```{r, fig.show='hide',results= 'hide'}
## Monthly Means
climate_data$Month=format(as.Date(climate_data$NewDate),format="%m")
climate_data$Year=format(climate_data$NewDate,format="%Y")
MonthlyTMAXMean=aggregate(TMAX~Month+Year,climate_data,mean)
 MonthlyTMAXMean$Year=as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$Month=as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX,ty='l')
#Selecting 1 Month (May)
plot(TMAX~Year,data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="5",],ty='l', xlim=c(1950,2020))
May.lm<-lm(TMAX~Year,data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="5",])
summary(May.lm)
abline(coef(May.lm),col="red")
```


```{r, fig.show='hide',results= 'hide'}
#Now the change is -0.005537 degrees C/year or 0.5 degrees Celcius per 100 years with a probability of 0.3528. Since p>.05, we cannot reject the null hypothesis that there is no change (and thus conclude there's a significant change in temp.)
MonthlyTMINMean=aggregate(TMIN~Month+Year,climate_data,mean)
MonthlyTMINMean$Year  =as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Year =as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Month=as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
str(MonthlyTMINMean)
```


```{r, fig.show='hide',results= 'hide'}
## Annual Maximum Temperatures
Months=c("January", "February", "March","April","May","June","July","August","September","October","November","December")
par(mfrow=c(4,3),mar=c(2,2,2,2)+0.1)
TMAXresult<-NA
# loop 
for(i in 1:12){
  plot(TMAX~Year,data=MonthlyTMAXMean[MonthlyTMAXMean$Month==i,],ty="l",las=1,xlim=c(1940,2020), main=Months[i])
  Month.lm<-lm(TMAX~Year,data=MonthlyTMAXMean[MonthlyTMAXMean$Month==i,])
  summary(Month.lm)
  abline(coef(Month.lm),col="red")
  TMAXresult<-rbind(TMAXresult,cbind(Months[i],round(coef(Month.lm)[2],4),round(summary(Month.lm)$coefficients[2,4],4),round(summary(Month.lm)$r.squared,3)))}

```



```{r, fig.show='hide',results= 'hide', message=FALSE}
## Annual Minimum Temperatures
Months=c("January", "February", "March","April","May","June","July","August","September","October","November","December")
par(mfrow=c(4,3),mar=c(2,2,2,2)+0.1)
TMINresult<-NA
for(i in 1:12){
  plot(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month==i,],ty="l",las=1,xlim=c(1940,2020), main=Months[i])
  Month.lm<-lm(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month==i,])
  summary(Month.lm)
  abline(coef(Month.lm),col="red")
  TMINresult<-rbind(TMINresult,cbind(Months[i],round(coef(Month.lm)[2],4),round(summary(Month.lm)$coefficients[2,4],4),round(summary(Month.lm)$r.squared,3)))}
```


```{r,fig.show='hide', message=FALSE}
plot(MonthlyTMINMean$TMIN,ty='l')
```


```{r, fig.show='hide',results= 'hide', message=FALSE}
## Precipitation
climate_data$PRCP[climate_data$PRCP==-999]<-NA
Missing<-aggregate(is.na(climate_data$PRCP),
                   list(climate_data$Month,climate_data$Year),sum)
Missing$Date=as.numeric(Missing$Group.1)+as.numeric(Missing$Group.2)/12
plot(x~Date,data=Missing)
#check to see what you created
TotalPPT<-aggregate(climate_data$PRCP,
                    list(climate_data$Month,climate_data$Year),sum,na.rm=T)
names(TotalPPT)=c("Group.1","Group.2","ppt")
NonMissing<-Missing[Missing$x<5,c(1:3)]
library(dplyr)
##Attaching package:'dplyr'
PPT<-merge(TotalPPT,NonMissing,all.y=TRUE)
PPT$Date<-as.numeric(PPT$Group.1)+as.numeric(PPT$Group.2)/12
head(PPT)
PRCP_mean=mean(PPT$ppt)
plot(ppt~Date,data=PPT)
abline(h=PRCP_mean,col="blue")

```

```{r, fig.show='hide',results= 'hide', message=FALSE}
climate_data$Month=format(as.Date(climate_data$NewDate),format="%m")
climate_data$Year=format(climate_data$NewDate,format="%Y")
MonthlyTMINMean=aggregate(TMIN~Month+Year,climate_data,mean)
 MonthlyTMINMean$Year=as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Month=as.numeric(MonthlyTMINMean$Month)
str(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN,ty='l')
#Selecting 1 Month (August)
plot(TMAX~Year,data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="8",],ty='l', xlim=c(1950,2020))
August.lm<-lm(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month=="8",])
summary(August.lm)
abline(coef(August.lm),col="red")


```






Climate data from the Global Historial Climate Newtwork (GHCN-Daily) includes daily land surface observations. Land surface observations are collected via satellites, which measure wavelengths to make inferences of heat. There are some limitations to this data as some entries were missing, and reliable data collection only began comparatively recently in 1934; however the data from GHCN-Daily is rather robust as it allows for updates to occur 7 days a week providing the most up-to-date information. Additionally, station values for each daily update undergo quality checks. 



## Minimum Temperatures in San Diego in August
```{r echo=FALSE}
par(cex.main=1, cex.lab=1, cex.axis=1)
plot(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month=="8",],
ty="l", xlim=c(1939,2020), main="Minumum Temperatures in San Diego in August",
xlab="Year", ylab="Minimum Temperatures (°)", col="grey")
bestfitAugust<-lm(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month=="8",])
abline(bestfitAugust, col="red")
```

Prolonged drought conditions are predicted to increase the likelihood of large wildfires in the San Diego region in the upcoming decades, especially during summer months such as August when temperatures are at their peak. This year, San Diego saw one fire in August, the Volcano Fire, lasting two days and burning 45 acres west of Temecula City. 


We can reject the null hypothesis that there is no change in average minimum temperatures in San Diego in August for the period between 1939 and 2020. The slope is increasing at a rate of 0.0283 with a p-value of 1.321e-06 (p<.001); thus, we can be fairly certain that temperatures are generally increasing at a rate of 0.0283 degrees Celcius per year.The R-squared value is 0.255, thus for the month of August, about 25.5% of the variance in the model is explained by time.




## Minimum Temperatures in September in San Diego
```{r echo=FALSE}
par(cex.main=1, cex.lab=1, cex.axis=1)
plot(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month=="9",],
ty="l", xlim=c(1939,2020), main="Minumum Temperatures in September in San Diego",
xlab="Year", ylab="Minimum Temperatures (°)", col="grey")
bestfitSeptember<-lm(TMIN~Year,data=MonthlyTMINMean[MonthlyTMINMean$Month=="9",])
abline(bestfitSeptember, col="red")
```



In six climate and emissions simulations produced by Messner et al., (2011), climate change expected to cause atmospheric warming in the San Diego region over the next five decades. Although historical data has shown that extreme warm temperatures mostly occurred from July to August in the region, as climate change continues, extreme warm temperatures are likely to be seen sooner in the year (June) and last longer into September (Messner et al., 2011).
Temperature data for the month of September gathered from NOAA climate records reveal that average minimum temperatures in San Diego have been generally increasing since the first reliable measurements were gathered 1939. It's these rise in temperatures that have likely contributed to the Valley Fire, an ongoing fire that was sparked on September 5th, 2020.



We can reject the null hypothesis that there is no change in average minimum temperatures in San Diego in the month of September for the period between 1939 and 2020. The slope is increasing at a rate of 0.0337 with a p-value of (p<.001); thus, we can be fairly sure that temperatures are generally increasing at a rate of 0.03 degrees Celcius per year. Additionally, the R-squared value is 0.267, meaning that for the month of September, about 26.7% of the variance in the model is explained by time.

## Minimum Temperatures in San Diego Data Table
```{r, results='asis'}
library(xtable)
Results<-data.frame(Month=TMINresult[c(2:13),1],TMINSlope=TMINresult[c(2:13),2], TMIN_P=as.numeric(TMINresult[c(2:13),3]),
                    TMINRsq=TMINresult[c(2:13),4],TMAXSlope=TMAXresult[c(2:13),2],TMAX_P=as.numeric(TMAXresult[c(2:13),3]),
                    TMAXRsq=TMAXresult[c(2:13),4])
Results$starTMIN="NS"
Results$starTMIN[Results$TMIN_P<=.05]="*"
Results$starTMIN[Results$TMIN_P<0.01]="**"             
Results$starTMIN[Results$TMIN_P<0.001]="***"
Results$starTMAX="NS"
Results$starTMAX[Results$TMAX_P<0.05]="*"
Results$starTMAX[Results$TMAX_P<0.01]="**"
Results$starTMAX[Results$TMAX_P<0.001]="***"
Results$TMINslope=paste(Results$TMINSlope,Results$starTMIN)
Results$TMAXslope=paste(Results$TMAXSlope,Results$starTMAX)
colnames(Results)<-c("Month","2","3","R^2","5","6","R^2","8","9", "SlopeTMIN","SlopeTMAX")
print(xtable(Results[,c(1,10, 4,11,7)]), type = "html")
```
