---
title: "Aberdeen Blog"
author: "Minah Choi"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#file="/home/CAMPUS/mmc02014/Climate_Change_Narratives/Data/SP18/choi_aberdeen station_data.csv"
file="/home/CAMPUS/mmc02014/Climate_Change_Narratives/Data/SP18/choi_aberdeen station_data.csv"
climate_data <- read.csv(file)

strDates <- as.character(climate_data$DATE)
head(strDates)
climate_data$NewDate <- as.Date(strDates, "%m/%d/Y%")
str(climate_data)
```

## Markdown

```{r cars}
summary(cars)

#file="/home/CAMPUS/mmc02014/Climate_Change_Narratives/Data/SP18/choi_aberdeen station_data.csv"
file="/home/CAMPUS/mmc02014/Climate_Change_Narratives/Data/SP18/choi_aberdeen station_data.csv"
climate_data=read.csv(file)

head(climate_data)
str(climate_data)
names(climate_data)
min(climate_data$TMAX)
min(climate_data$TMAX, na.rm = T)
plot(TMAX~DATE, climate_data)
climate_data$TMAX[climate_data$TMAX==-9999] = NA
climate_data$TMIN[climate_data$TMIN==-9999] = NA
plot(TMAX~DATE, climate_data[1:1891,], ty='1')
str(climate_data)
strDates <- as.character(climate_data$DATE)
head(strDates)
climate_data$NewDate <- as.Date(strDates, "%m/%d/Y%")
str(climate_data)
```

my dates are all goofy--


```{r plot, echo=FALSE}
summary(plot)

plot(TMAX )

lm(TMAX ~ NewDate, data = climate_data)

summary(lm(TMAX ~ NewDate, data = climate_data))

climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)

MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

str(MonthlyTMAXMean)

plot(MonthlyTMAXMean$TMAX, ty='1')

```

i'm not sure what exactly happened here, but i went ahead and started moving through the coding on SOP86_
it won't recognize my objects for some reason? I read through some of the coding on previous blogs (hope that was ok) to try to see what was going on.

```{r 1 month}
summary(month)

plot(TMAX~YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",], ty='1', xlim=c(1950, 2020))
May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
summary(May.lm)



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
