---
title: "Climate Change and Logical Fallacies"
subtitle: "Critical Thinking and Climate Change Implications"
author: "EA30, Pomona College"
date: "Nov. 2017"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: ../skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```
## EPA Adminstrator and Logical Fallacies

According the the Washington Post, Scott Pruitt, Trump's EPA adminstrator told an audience at the Heritage Foundation, "What's most important at the agency is to have scientific advisers that are objective, independent-minded, providing transparent recommendations" [@dennis2017]. By having EPA science advisors who also receive funding from the agency causes Pruitt to question "the independence and the veracity and the transparency" of the recommendations made by these individuals. 

Apparently, Pruitt's concern objectivity in science. While this is a concern all scientists share, his claim that scientists that recieve EPA funding impedes objectivity. 
Let's unpack his assumptions here and determine if there is a logical fallacy...

First, what EPA panels is Pruitt refering to? I was able to find 13 advsisory panels: 

- EPA Board of Scientific Counselors
- Federal Insecticide, Fungicide, and Rodenticide Act Scientific Advisory Panel
- Science Advsisor Board -- Chemical Assessment Advisory Committee
- Science Advsisor Board -- Drinking Water Committee
- Science Advsisor Board -- Ecological Processes and Effects Committee
- Science Advsisor Board -- Environmental Economics Advisory Committee
- Science Advsisor Board -- Environmental Engineering Committee
- Science Advsisor Board -- Radiation Advisory Committee
- Economy-wide Modeling of the Benefits and Costs of Environmental Regulation
- IRIS Assessment for Ethyl Tertiary Butyl Ether
- IRIS Assessment for tert-Butyl Alcohol
- Scoping and Approach for Revising Guidelines for Devising Numerical Water Quality Criteria to Protect Aquatic Life
- Screening Methodologies to Support Risk and Technology Reviews: A Case Study Analysis


### Independence

Independence is often articulated as value when thinking about the scientific process. There are two ways we can think about scientific indendence. First, that the results and conclusions are not influenced by forces outside the experimental or procedural methods. Second, that funding of the sciences does not influece the outcomes of the science.

Pruitt's argument might stem from a criticism that scientist are lemmings and willing to do research to support specific finacial incentives (funding sources with particular goals). That scientists are willing to sacrafice their scientific integrity to pursue these funding opportunities (so they can be rewarded, e.g. money to the universities, and publications for promotion...). 

One question my students asked it's more of a question of who is ultimately tapped -- you could replace with more or less indendent, then we need to have to measure independence. 

First, university faculty are paid by the university, although research grants might augment their salaries, they remain univerisyt employess. It's pretty independent.


### Veracity

The question of varacity questions panels ability to conform to the facts. Of course, the irony that Pruitt's boss struggles with this more than most, we'll try to focus on the validity of the the  concern. 

### Transparency

On of the key aspects in science is transparency. Transpareny in how a study was designed, how data was collected and analyzed. These form the central tennets of high quality scientific work. 


## Carbon Dioxide as a Major Greenhouse Gas

Although a host of gases can trap heat, carbon dioxide or CO~2~ is particularly important. CO~2~ concentrations are relatively large compared to most greenhouse gases and have increased due land use changes and the combustion of fossil fuels. We even know how much of the CO~2~ in the atmosphere has come from burning of fossil fuels.

```{r maunaloa, echo=FALSE}
address <- "ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt"
download.file(address, "maunaloa", quiet = F, mode = "w", cacheOK = T)

maunaloa <- read.table("maunaloa", skip=70)
names(maunaloa) <- c("year", "month", "decimal.date", "average", "interpolated", "trend", "days")
maunaloa$average[maunaloa$average==-99.99] <- NA
maunaloa <- data.frame(year=maunaloa$year, month=maunaloa$month, decimal.date=maunaloa$decimal.date, average=maunaloa$average)
#head(maunaloa)
#str(maunaloa)
#write.csv(maunaloa, file = "H:\\My Documents\\My Webs\\Mauna_Loa.csv", eol = "\n", row.names = F)
```
Atmospheric CO<sub>2</sub> concentrations are measured in a continuous fashion from several locations around the world. The longest record is measured on the volcano Mauna Loa, Hawaii. Using these data, we can estimate how CO~2~ concentrations have changed.


```{r fig-margin, fig.margin = TRUE, fig.cap = "Observed CO~2~ concentrations (black) have been steadily increasing since we have begun measuring them in the 1950s, with a best fit line (red). Note that the slope or the rate of change is also increasing.", fig.width=3.5, fig.height=3.5, cache=TRUE, echo=FALSE}
par(las=1)
plot(maunaloa$decimal.date, maunaloa$average, type="l", ylab=expression(CO[2]*~ "(ppm)"), xlab="Year", 
main="Carbon Dioxide Concentration \n Mauna Loa, HI" )
abline(coef(lm(average~year, data=maunaloa)), col="red", lwd=2)
```

```

Although political debate about the causes of climate change exists as EPA Pruitt correctly points out, the scientific debate is limited to subtle areas of uncertainty -- for example, at what ocean temperature might methane hydrates from the ocean be released into the atmosphere? Or how quickly will high latitude tundra become a major CO~2~ source? Or when will melting ice shelves significantly contribute to sea level rise?

In contrast to the red herring nature of political debates about the role of greenhouse gases on climate change, we know how they influence the planet's climate from multiple sources of evidence [@pachauri2015ipcc]. Furthermore, as a major greenhouse gas, CO<sub>2</sub> concentrations are increasing as a result of our use of fossil fuels and already influences the planet's climate.

We have been quite intentional addressing logical falacies...

## Daily Temperature Data

As a class we used daily data to evaluate temperature changes at single stations. Collected in some cases for over 130 years represents some of the longest records of temperature. Originally collected to support local economic activity, especially farming, the uses of these data has more diverse in recent decades. 

In our case, to analyze long-term temperature changes. According to XX, daily data are not ideal for long-term trend analyses. This is because... 

## Climate Change Scientists and Claiming Implications

Without a doubt, one of the most probablematic areas for scientists is to figure out how to place their research in "broader impacts", a category used by the National Science Foundation for evaluating research grants, and remain within the solid footing of scientifically based conclusions....

As a class, we have been cognizant of extrapolating beyond the science and making sure we are explicit when making normative statements that might be beyond normal scientific conclusions...



## Fall 2017 Blogs and Op Ed

[Torrance, CA by Allison Joseph](Joseph_blog_1700g.html)

[Santa Monica, Bebe Phornprapha](phornprapha-blog-171009.html)

[Sugar Land, TX by Kirara Tsutsui](Tsutsi_Blog_17_10_9.html)

[Claremont California, CA by Valentina Emanuele](Emanuele_blog_171009.html)

[Cape May, by Caroline Koh Smith](draft_blog.html)

[San Francisco, CA by Meily Wu Fung](2017-10-09_SanFranciscoBlog.html)

