---
title: "Denver Temperature Changes"
author: "Thea Piccone"
date: "12/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Last year, Denver, the capital of Colorado, was named one of the best places to live in the US, according to a U.S. News and World Report. The rich and diverse culture, a strong economy, and an arid and sunny climate that lends itself to enjoying the multitude of possible outdoor activities combine to create a very attractive city, especially to millennials. Denver has developed greatly since its inception as a mining hub during the gold rush of the mid 1800’s.  Today, a diverse group of mostly young, progressive-minded nature lovers have replaced the miners who originally founded the city; but, although there is no gold rush, a surging economy is still a fundamental attribute of the city. A combination of a strong footing in the tech industry, a booming marijuana industry, and an abundance of highly profitable natural resources have resulted in a very strong economy with low unemployment rates, a high average annual salary, and a thriving job market. 

Characterized by a typically sunny and dry climate, Denver is located within the semi-arid, continental zone at the foot of the Rocky mountains. Denver has four distinct seasons that are all subject to sudden weather changes due to Denver’s geographical location in the inland High Plains and at the edge of the Rockies. Although weather patterns may be infrequent, Denver locals are able to enjoy roughly 300 days of sunshine throughout the year. Both the summer and the winter seasons in Denver are relatively mild with average temperatures of 31.1 degrees fahrenheit in the winter and of 71.4 degrees fahrenheit in the summer. These comfortable temperature trends are in part due to Denver’s high altitude of 5,250 feet above sea level; an altitude that lends itself to Denver’s nickname, the mile-high city. Such a climate helps to foster locals’ fervent love for the outdoors, and surrounded by forests, mountains, and lakes, the opportunities to act on such an admiration can feel endless. Thus, with a strong economy, a rich history, an agreeable climate, and beautiful natural landscape, it becomes clear as to why Denver was ranked as one of the best places to live in the US. However, as temperatures continue to rise due to climate change the vitality of this young and lively city will be severely threatened. 

According to the following climate data from the National Centers for Environmental Information (NOAA) the average maximum temperatures for the months of March and July in Denver have been steadily increasing since 1948. In other words, these two months have been steadily warming for the past seven decades. This data reflects the average rise of temperature globally reported by the Intergovernmental Panel on Climate Change (IPCC). 
```{r importingCSV, include=FALSE}
# Importing Scripts

file = "/home/CAMPUS/tgpv2015/github/Climate Change Narratives/Data/denver1-TP.csv"

import = read.csv(file)
```
```{r checking, include=FALSE}
#plot(TMAX~DATE, import)
```
```{r missings, include=FALSE}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```
```{r, include=FALSE}
#plot(TMAX~DATE, import[import$DATE<19501231,], ty='l')
```
```{r, include=FALSE}
#plot(TMAX~DATE, import[import$DATE<19500102,], ty='l')
```
```{r, include=FALSE}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```
```{r, include=FALSE}
#plot(TMAX~NewDate, import[import$DATE<19500102,], ty='l')
```
```{r, include=FALSE}
unique(import$STATION_NAME)
```
```{r, include=FALSE}
LosAngeles <- subset(import, STATION_NAME=="DENVER STAPLETON CO US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
```
```{r, include=FALSE}
#plot(TMAX~NewDate, LosAngeles, ty='l')
```
```{r, include=FALSE}
# Linear Model
LosAngeles.lm <- lm(TMAX~NewDate, data=LosAngeles)
summary(LosAngeles.lm)
plot(TMAX~NewDate, LosAngeles, ty='l')
abline(coef(LosAngeles.lm), col='red')
```
```{r, include=FALSE}
# Get months
LosAngeles$Month = months(LosAngeles$NewDate) # Creates problems.
LosAngeles$Month = format(as.Date(LosAngeles$NewDate), format = "%m")
LosAngeles$Year = format(LosAngeles$NewDate, format="%Y")
```
```{r creatingmonthlymeans, include=FALSE}
MonthlyMean = aggregate(TMAX ~ Month + Year, LosAngeles, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(TMAX ~ Month + Year, LosAngeles, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```
```{r plottingMonthyMeans, include=FALSE}
plot(MonthlyMean$TMAX, ty='l')

```
```{r plottingMonthySD, include=FALSE}
plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(TMAX~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```
`
`

The following graphs model the linear relationship between time (in years) and average maximum temperature (in degrees fahrenheit) for the month of March and July:

## March
```{r Mar, echo=FALSE, results='hide'}
#plot(MonthlyMean$TMAX[MonthlyMean$Month=="03"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="03",], ty='l', las=1, ylab="Maximum Temperature (F)", xlab="Year") 
Mar.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="03",])
summary(Mar.lm)

abline(coef(Mar.lm), col="purple")
```

## July 
```{r Jul, echo=FALSE, results='hide'}
#plot(MonthlyMean$TMAX[MonthlyMean$Month=="07"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="07",], ty='l', las=1, ylab="Maximum Temperature (F)", xlab="Year") 
Jul.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="07",])
summary(Jul.lm)

abline(coef(Jul.lm), col="purple")
```
`

Both of these graphs represent significant data trends that reject the null hypothesis, or the position that there is no relationship between time and temperature rise for March and July in Denver since 1948. We can confidently state that there is such a relationship because the data produces a p-value of less than 0.05. In other words, the probability that the data does in fact support the null hypothesis and that our calculations are reflctive of an error is less than five percent. This tells us that we have strong evidence in support of the alternative hypothesis that there is a significant relationship between time and temperature rise for these two months.      


These trends representing a rise in temperature, when paired with global climate data and research on climate change reveal many important changes that we will see in Denver. The rise in temperature for the months of March and July are projected to reflect some of the devastating consequences that accompany such temperature increase. Denver’s crucial connection to the natural resources of the Rocky mountains will be especially threatened as temperatures shift.  The most threatened component of this relationship is Denver’s water supply. Denver receives most of its water from rivers and streams fed by snowmelt from the Rocky mountains.  As the spring season grows longer and warmer, as is reflected in the NOAA data, the effective and crucial transferring of water from the relatively wet winter season to the typically dry summers is hindered. These warmer spring temperatures are therefore increasing early snowmelt from the Rocky Mountains and we have already seen over a 20 percent decline in spring snow cover throughout the Rockies since 1980. This early loss of snowpack results in initial flooding and erosion followed by extended periods of drought in the Summer. These periods of drought result in issues of water scarcity that will lead to unavoidable and costly consequences for the city of Denver and its residents. 

The loss of snowpack will further impair Denver’s economy as Colorado’s largest economic asset is severely threatened by rising temperatures. Although Denver, being a city, does not contain any ski or snowboard mountains, it is surrounded by mountains filled with these snow sport activities. Additionally, the city of Denver owns and operates one of these mountain resorts. As mentioned earlier, Denver tends to attract both residents and tourists that are very enthusiastic about outdoor activities, and given Denver’s proximity to a multitude of mountain ranges this enthusiasm especially pertains to snow sport activities such as skiing and snowboarding. As the spring season grows longer and warmer, an estimate of roughly 30 days will be cut from the snow season. The Denver Post reported that ski resorts have lost up to 2 million dollars in revenue due to scarcer annual snowfall. Even an annual decrease in only 1 percent in the amount of tourists visiting Colorado in order to ski or snowboard would result in an expected loss of over $375 million and over 4,500 jobs lost in the span of under 10 years. This hit to the state’s economy will undoubtedly hinder Denver’s economy as well as the morale of Denver’s citizens.

Given the current tumultuous state of the United States government and its proposed responses to climate change, it is important that we are nonetheless able to identify and present significant climate data trends. In Denver, such a trend will undeniably affect the economy and landscape of Denver and the well-being of its citizens.  
`
`

```{r, include=FALSE}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="03"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="03",], ty='l') 
Mar.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="03",])
summary(Mar.lm)

abline(coef(Mar.lm), col="red")

```



```{r, include=FALSE}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, LosAngeles, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```


```{r, include=FALSE}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="03"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="03",], ty='l') 
Mar.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="03",])
summary(Mar.lm)

abline(coef(Mar.lm), col="red")

```


```{r, include=FALSE}
PRCP_mean = mean(LosAngeles$PRCP, na.rm=T)

plot(PRCP~NewDate, data=LosAngeles)
abline(h=PRCP_mean, col="blue")

```


```{r, include=FALSE}
LosAngeles$PRCP[LosAngeles$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, LosAngeles, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```


```{r, include=FALSE}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="green")

n <- 5
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
lines(YearlySum$YEAR, y_lag, col="red")
```


```{r, include=FALSE}
summary(YearlySum.lm)
```


```{r, include=FALSE}
YearlySum2 = subset(YearlySum, select=c(PRCP))
YearlyMean = mean(YearlySum2$PRCP)

#filter(YearlySum2)



#Doesn't work...
# ma <- function(x,n=5){filter(x,rep(1/n,n), sides=2)}
# ma(YearlySum$PRCP)  


yearlyts =ts(YearlySum2) 
str(yearlyts)




```