%\subsection{Op-Ed 1}

\subsection{Rationale}

Climate change may be the most controversial environmental issue in history. However, compared to other issues, this history is relatively short. Fueled by opposing political parties and industry goals, the conclusions of scientists is a fundamental source of conflict -- thus, science itself has become extremely politicized. 

Nevertheless, how and where science and scientists became embroiled became a battle ground negotiating the appropriate level of regulation (regulatory reach), economic and industrial \textit{Laissez-faire}, and environmental risks. Environmental issues are almost always controversial and in the case of climate change, few dominate the political agenda like climate change. 

Nevertheless, in a pandemic and election year, the political states may be higher than normal. The role of activism in the US (and world) has changed in recent years with a higher reliance on social media. Of course, with the current requirements for Covid-19 social distance, social media might be the primary source of information for many. 

In this context, we need to determine the role of and changes in activism and it's role in climate science. Moreover, we need to evaluate the how activist are using scientific information, especially climate records as they promote their agendas.

\subsection{Assignment}

Write an Op-Ed piece that describe the role of activism in climate change, their use of climate data during a pandemic and election year. %outlines why residents in a specific US region should care about temperature changes. 
Spend sometime deciding what is currently in the news that you consider a compelling issue to your audience.

\subsection{Submission Format and Naming Convention}

Submit your Op-Ed as a pdf via \texttt{Sakai}, using the following naming convention:

\begin{center}
\textbf{Op-Ed\_1\_XXXXX.pdf},
\end{center}

\noindent using one of your 5 digit random numbers for the Xs. See \url{https://github.com/marclos/Climate_Change_Narratives/raw/master/Admin/RandomNumbers.pdf} to get the list of assigned random numbers. 

\subsection{Grading}

The Op-Ed will be graded using several criteria. First, the topic must be compelling -- connecting current affairs to the historical issues of climate. Second, the Op-Ed should rely on several sources of evidence and citations, while creating fluid prose that compel the reader to continue reading. If the reader gets stuck in statistics or technical jargon, it can be like wading in mud -- but without some ``numbers'' the argument may become glittering generalities without a sense of a gritty reality. Again, your job is to find a compelling balance. Finally, you want the read to jump out of their seat and ``do something''. Thus, the Op-Ed should compel the reader into action, see assignment handout for more information.

%\subsection{Readings}




