---
title: "Los Angeles Analysis"
author: "19890"
date: "9/18/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r climate}
climate_data <- read.csv("/home/CAMPUS/cdma2019/ChristinaMarsh_LosAngeles_data.csv")
head(climate_data)
str(climate_data)
names(climate_data)
plot(TMAX~DATE, climate_data) 
min(climate_data$TMAX)
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
plot(TMAX~NewDate, climate_data[1:1835,], ty='l')
TMAX.lm = lm(TMAX ~ NewDate, data=climate_data)
plot(TMAX ~ NewDate, data= climate_data, las=1)
```


```{r TMAX monthly Data}
plot(TMAX~NewDate, climate_data[1:1835,], ty='l')
lm(TMAX ~ NewDate, data=climate_data)
plot(TMAX ~ NewDate, data= climate_data, las=1)
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean) 
str(MonthlyTMAXMean)
MonthlyTMAXMean$Year.num = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$Month.num = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX, ty='l')
```
```{r August TMAX Data}
plot(TMAX~NewDate, climate_data[1:1835,], ty='l')
lm(TMAX ~ NewDate, data=climate_data)
plot(TMAX ~ NewDate, data= climate_data, las=1)
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean) 
str(MonthlyTMAXMean)
MonthlyTMAXMean$Year.num = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$Month.num = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX, ty='l')
plot(TMAX~Year.num, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",], ty='l', xlim=c(1906, 2014))
August.lm <- lm(TMAX~Year.num, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="08",])
summary(August.lm)
abline(coef(August.lm), col="red")
```
```{r September TMAX DATA}
plot(TMAX~NewDate, climate_data[1:1835,], ty='l')
lm(TMAX ~ NewDate, data=climate_data)
plot(TMAX ~ NewDate, data= climate_data, las=1)
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean) 
str(MonthlyTMAXMean)
MonthlyTMAXMean$Year.num = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$Month.num = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX, ty='l')
plot(TMAX~Year.num, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",], ty='l', xlim=c(1906, 2014))
September.lm <- lm(TMAX~Year.num, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="09",])
summary(September.lm)
abline(coef(September.lm), col="red")
```

```{r Monthly TMIN Data}
plot(TMIN~NewDate, climate_data[1:1835,], ty='l')
lm(TMIN ~ NewDate, data=climate_data)
plot(TMIN ~ NewDate, data= climate_data, las=1)
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)    
MonthlyTMINMean$Year.num = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Month.num = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN, ty='l') 
```

```{r September TMIN Data}
plot(TMIN~NewDate, climate_data[1:1835,], ty='l')
lm(TMIN ~ NewDate, data=climate_data)
plot(TMIN ~ NewDate, data= climate_data, las=1)
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean) 
str(MonthlyTMINMean)
MonthlyTMINMean$Year.num = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Month.num = as.numeric(MonthlyTMINMean$Month)
str(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN, ty='l')
plot(TMIN~Year.num, data=MonthlyTMINMean[MonthlyTMINMean$Month=="09",], ty='l', xlim=c(1906, 2014))
September.lm <- lm(TMIN~Year.num, data=MonthlyTMINMean[MonthlyTMINMean$Month=="09",])
summary(September.lm)
abline(coef(September.lm), col="red")
```
```{r August TMIN DATA}
plot(TMIN~NewDate, climate_data[1:1835,], ty='l')
lm(TMIN ~ NewDate, data=climate_data)
plot(TMIN ~ NewDate, data= climate_data, las=1)
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean) 
str(MonthlyTMINMean)
MonthlyTMINMean$Year.num = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Month.num = as.numeric(MonthlyTMINMean$Month)
str(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN, ty='l')
plot(TMIN~Year.num, data=MonthlyTMINMean[MonthlyTMINMean$Month=="08",], ty='l', xlim=c(1906, 2014))
August.lm <- lm(TMIN~Year.num, data=MonthlyTMINMean[MonthlyTMINMean$Month=="08",])
summary(August.lm)
abline(coef(August.lm), col="red")
```

```{r TMAX monthly}
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$Year.num = as.numeric(MonthlyTMINMean$Year) 
MonthlyTMINMean$Month.num = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN, ty='l')
Months = c("January", "February", "March", "April","May", "June", "July", "August", "September", "October","November", "December")
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMAXresult <- NA
par(mar=c(1,1,3,2))
for (i in 1:12) {
plot(TMAX ~ Year.num, data = MonthlyTMAXMean[MonthlyTMAXMean$Month.num == i, ], ty = "l", las = 1, xlim = c(1906, 2014), main = Months[i], ylim = c(5, 35))
Month.lm <- lm(TMAX ~ Year.num, data = MonthlyTMAXMean[MonthlyTMAXMean$Month.num == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Month.lm)[2], 4),round(summary(Month.lm)$coefficients[2,4], 4), round(summary(Month.lm)$r.squared,3))) }
```

```{r TMIn Monthly}
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$Year.num = as.numeric(MonthlyTMINMean$Year) 
MonthlyTMINMean$Month.num = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN, ty='l')
Months = c("January", "February", "March", "April","May", "June", "July", "August", "September", "October","November", "December")
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMINresult <- NA
par(mar=c(1,1,3,2))
for (i in 1:12) {
plot(TMIN ~ Year.num, data = MonthlyTMINMean[MonthlyTMINMean$Month.num == i, ], ty = "l", las = 1, xlim = c(1906, 2014), main = Months[i]) 
Month.lm <- lm(TMIN ~ Year.num, data = MonthlyTMINMean[MonthlyTMINMean$Month.num == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMINresult <- rbind(TMINresult, cbind(Months[i], round(coef(Month.lm)[2], 4),round(summary(Month.lm)$coefficients[2,4], 4), round(summary(Month.lm)$r.squared,3))) }
```
Based on my anaylsis there is a clear upward trend in the average monthly minuimum temperatures excluding November and December.  

```{r precipitation data}
climate_data$PRCP[climate_data$PRCP==-9999] <- NA 
Missing <- aggregate(is.na(climate_data$PRCP),
list(climate_data$Month, climate_data$Year), sum)
Missing$Date = as.numeric(Missing$Group.1) + as.numeric(Missing$Group.2)/12
TotalPPT <- aggregate(climate_data$PRCP,
list(climate_data$Month, climate_data$Year), sum, na.rm=T)
names(TotalPPT) = c("Group.1", "Group.2", "ppt")
NonMissing <- Missing[Missing$x < 5, c(1:3)]
library(dplyr)
PPT <- merge(TotalPPT, NonMissing, all.y=TRUE)
PPT$Date <- as.numeric(PPT$Group.1) + as.numeric(PPT$Group.2)/12
head(PPT)
PRCP_mean = mean(PPT$ppt)
plot(ppt~Date, data=PPT)
abline(h=PRCP_mean, col="blue")
``` 
Based on my analysis, my data shows little fluctuation in average annual percipatation. 

```{r results}
Results <- data.frame(Month = TMINresult[c(2:13),1],
TMINSlope = TMINresult[c(2:13),2],
TMIN_P = as.numeric(TMINresult[c(2:13),3]),
TMINRsq = TMINresult[c(2:13),4],
TMAXSlope = TMAXresult[c(2:13),2],
TMAX_P = as.numeric(TMAXresult[c(2:13),3]),
TMAXRsq = TMAXresult[c(2:13),4])
Results$starTMIN = "NS"
Results$starTMIN[Results$TMIN_P <= .05] = "*"
Results$starTMIN[Results$TMIN_P < 0.01] = "**"
Results$starTMIN[Results$TMIN_P < 0.001] = "***"
Results$starTMAX = "NS"
Results$starTMAX[Results$TMAX_P < 0.05] = "*"
Results$starTMAX[Results$TMAX_P < 0.01] = "**"
Results$starTMAX[Results$TMAX_P < 0.001] = "***"
Results$TMINslope=paste(Results$TMINSlope, Results$starTMIN)
Results$TMAXslope=paste(Results$TMAXSlope, Results$starTMAX)
colnames(Results) <- c("Month", "2", "3", "R^2", "5", "6",
"R^2", "8", "9", "Slope TMIN", "Slope TMAX") 
library(xtable)
print(xtable(Results[,c(1, 10, 4, 11, 7)])) 
```

Based on my analysis there is significant variation in the data collected in all months excluding November and December which showed relatvivley stable temperature trends.

