---
title: "20190224_boulder_graphs"
author: "bljj2015"
date: "2/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read NWT and NOAA data, include=FALSE}
library(readr)
C1_daily_precip <- read_csv("C1_daily_precip.csv")
C1_daily_temp <- read_csv("C1_daily_temp.csv")
D1_daily_temp <- read_csv("D1_daily_temp.csv")
D1_daily_precip <- read_csv("D1_daily_precip.csv")
Saddle_daily_temp <- read_csv("Saddle_daily_temp.csv")
NOAA <- read_csv("NOAA_boulder_area_temp_precip.csv")
```

```{r combining different date formats from Clair Generous}
strDates <-as.character(climate_data$DATE)
CYear1 <-as.Date(strDates[1:3805], "%Y-%m-%d")

CYear2 = as.Date(sub("/([0-9]+)$", "/19\\1", strDates[3806:40320]), "%m/%d/%Y") # Defined for this dataset! 
CYear3 = c(CYear1, CYear2, as.Date(sub("/([0-9]+)$", "/20\\1", strDates[40321:47286]), "%m/%d/%Y"))
CYear3[sample(1:47286, 5)]

climate_data$NewDate <- as.Date(CYear3, "%m/%d/%Y")

head(strDates)
head(climate_data)

plot(TMAX ~ NewDate, data=climate_data, ty='l')
```

```{r my attempt at changing dates}
strDates <-as.character(NOAA$DATE)
CYear1 <-as.Date(strDates[1:60719], "%m/%d/%Y") #1/1/1900 onward
CYear2 = as.Date(strDates[60720:62190], "%Y-%m-%d") #1893-10-01 to 1899-12-31
CYear3 = c(CYear1, CYear2) 
CYear3[sample(1:62190, 10)] #sample 10 dates to see if they're all the same format: YYYY-MM-DD
NOAA$NewDate <- CYear3
```

```{r separating NOAA data into three separate data frames for each of the 3 stations, echo = FALSE}
Boulder <- NOAA[ which(NOAA$STATION=='USC00050848'),]
Boulder_14_W <- NOAA[ which(NOAA$STATION=='USW00094075'),]
Coal_Creek_Canyon <- NOAA[ which(NOAA$STATION=='USC00051681'),]
```


####Daily Max Temperature graphs

```{r C-1 daily TMAX graph & linear regression model, echo=FALSE}
plot(airtemp_max~date, C1_daily_temp, main="Daily Max Temp at C-1 Station", ty='l') & abline(lm(airtemp_max ~ date, data=C1_daily_temp), col="red")
summary(lm(airtemp_max~date, data=C1_daily_temp))
```
slope = +9.226e-05  
p-value < 2e-16, STATISTICALLY SIGNIFICANT  

```{r D-1 daily TMAX graph & linear regression model, echo=FALSE}
plot(airtemp_max~date, D1_daily_temp, main="Daily Max Temp at D-1 Station", ty='l') & abline(lm(airtemp_max ~ date, data=D1_daily_temp), col="red")
summary(lm(airtemp_max~date, data=D1_daily_temp))
```
slope = -3.691e-05  
p-value = 7.08e-05, STATISTICALLY SIGNIFICANT  

```{r Saddle daily TMAX graph & linear regression model, echo=FALSE}
plot(airtemp_max~date, Saddle_daily_temp, main="Daily Max Temp at Saddle Station", ty='l') & abline(lm(airtemp_max ~ date, data=Saddle_daily_temp), col="red")
summary(lm(airtemp_max~date, data=Saddle_daily_temp))
```
slope = +1.254e-04  
p-value = 1.15e-08 -> STATISTICALLY SIGNIFICANT  

```{r Boulder daily TMAX graph & linear regression model, echo=FALSE}
plot(TMAX~NewDate, Boulder, main="Daily Max Temp at USC00050848 (Boulder) Station", ty='l') & abline(lm(TMAX ~ NewDate, data=Boulder), col="red")
summary(lm(TMAX~NewDate, data=Boulder))
```
slope = +4.741e-05  
p-value < 2e-16 -> STATISTICALLY SIGNIFICANT  

```{r Coal Creek Canyon daily TMAX graph & linear regression model, echo=FALSE}
plot(TMAX~NewDate, Coal_Creek_Canyon, main="Daily Max Temp at USC00051681 (Coal Creek Canyon) Station", ty='l') & abline(lm(TMAX ~ NewDate, data=Coal_Creek_Canyon), col="red")
summary(lm(TMAX~NewDate, data=Coal_Creek_Canyon))
```
slope = +3.645e-05  
p-value = 0.124, NOT STATISTICALLY SIGNIFICANT  

```{r Boulder 14 W daily TMAX graph & linear regression model, echo=FALSE}
plot(TMAX~NewDate, Boulder_14_W, main="Daily Max Temp at USW00094075 (Boulder 14 W) Station", ty='l') & abline(lm(TMAX ~ NewDate, data=Boulder_14_W), col="red")
summary(lm(TMAX~NewDate, data=Boulder_14_W))
```
slope = +2.385e-04  
p-value = 0.00247, STATISTICALLY SIGNIFICANT 
  
  
####Daily Precipitation graphs

```{r testing C1 daily precipitation, echo=FALSE}
plot(ppt_tot~date, main="Daily Precipitation at C-1 Station", C1_daily_precip, ty='l') & abline(lm(ppt_tot ~ date, data=C1_daily_precip), col="red")
summary(lm(ppt_tot~date, data=C1_daily_precip))
```
slope = -8.744e-05
p-value < 2e-16 -> STATISTICALLY SIGNIFICANT

```{r testing D1 daily precipitation, echo=FALSE}
plot(pppt_tot~date, main="Daily Precipitation at D-1 Station", D1_daily_precip, ty='l') & abline(lm(pppt_tot ~ date, data=D1_daily_precip), col="red")
summary(lm(pppt_tot~date, data=D1_daily_precip))
```
slope = +2.243e-05  
p-value = 0.0164 -> NOT STATISTICALLY SIGNIFICANT

```{r testing Saddle daily precipitation, echo=FALSE}
plot(ppt_tot~date, main="Daily Precipitation at Saddle Station", Saddle_daily_precip, ty='l') & abline(lm(ppt_tot ~ date, data=Saddle_daily_precip), col="red")
summary(lm(ppt_tot~date, data=Saddle_daily_precip))
```
slope = +2.571e-04  
p-value < 2e-16, STATISTICALLY SIGNIFICANT

```{r Boulder daily PRCP graph & linear regression model, echo=FALSE}
plot(PRCP~NewDate, main="Daily Precipitation at USC00050848 (Boulder) Station", Boulder, ty='l') & abline(lm(PRCP ~ NewDate, data=Boulder), col="red")
summary(lm(PRCP~NewDate, data=Boulder))
```
slope = +1.621e-06  
p-value = 0.379, NOT STATISTICALLY SIGNIFICANT  

```{r creating NOAA station graph and linear regression model for Coal Creek Canyon, echo=FALSE}
plot(PRCP~NewDate, main="Daily Precipitation at USC00051681 (Coal Creek Canyon) Station", Coal_Creek_Canyon, ty='l') & abline(lm(PRCP ~ NewDate, data=Coal_Creek_Canyon), col="red")
summary(lm(PRCP~NewDate, data=Coal_Creek_Canyon))
```
slope = +5.154e-06  
p-value = 0.688 -> NOT STATISTICALLY SIGNIFICANT 

```{r creating NOAA station graph and linear regression model for Boulder 14 W, echo=FALSE}
plot(PRCP~NewDate, main="Daily Precipitation at USW00094075 (Boulder 14 W) Station", Boulder_14_W, ty='l') & abline(lm(PRCP ~ NewDate, data=Boulder_14_W), col="red")
summary(lm(PRCP~NewDate, data=Boulder_14_W))
```
slope = +8.296e-05  
p-value = 0.0531 -> NOT STATISTICALLY SIGNIFICANT (but close!)  




##converting the "NaN" characters to NA values??? never figured it out
```{r}
as.numeric(C1_daily_precip$ppt_tot)
is.nan(as.numeric(C1_daily_precip$ppt_tot)) #shows whether value is NaN

```

##creating monthly comparisons for Boulder
```{r}
Boulder$Month = format(as.Date(Boulder$NewDate), format = "%m")
Boulder$Year = format(as.Date(Boulder$NewDate), format="%Y")
```

```{r}
Boulder_MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, Boulder, mean)
Boulder_MonthlyTMAXMean$YEAR = as.numeric(Boulder_MonthlyTMAXMean$Year)
Boulder_MonthlyTMAXMean$MONTH = as.numeric(Boulder_MonthlyTMAXMean$Month)
str(Boulder_MonthlyTMAXMean)
plot(Boulder_MonthlyTMAXMean$TMAX, ty='l')
```

```{r Boulder TMAX averages for May}
#plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month=="05"], ty='l')
plot(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="05",],
ty='l', xlim=c(1890, 2020))
Boulder_May_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="05",])
summary(Boulder_May_TMAX.lm)
##
## Call:
## lm(formula = TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$Month ==
## "05", ])
##
7
## Residuals:
## Min 1Q Median 3Q Max
## -2.85376 -0.93210 -0.04633 0.81231 3.15347
##
## Coefficients:
## Estimate Std. Error t value Pr(>|t|)
## (Intercept) 17.238621 13.753642 1.253 0.214
## YEAR 0.008756 0.006946 1.261 0.211
##
## Residual standard error: 1.302 on 73 degrees of freedom
## Multiple R-squared: 0.0213,Adjusted R-squared: 0.007897
## F-statistic: 1.589 on 1 and 73 DF, p-value: 0.2115
abline(coef(Boulder_May_TMAX.lm), col="red")
```

```{r Boulder TMAX averages for June}
#plot(Boulder_MonthlyTMAXMean$TMAX[Boulder_MonthlyTMAXMean$Month=="06"], ty='l')
plot(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="06",],
ty='l', xlim=c(1890, 2020))
Boulder_Jun_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="06",])
summary(Boulder_Jun_TMAX.lm)
abline(coef(Boulder_Jun_TMAX.lm), col="red")
```

```{r Boulder TMAX averages for July}
#plot(Boulder_MonthlyTMAXMean$TMAX[Boulder_MonthlyTMAXMean$Month=="07"], ty='l')
plot(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="07",],
ty='l', xlim=c(1890, 2020))
Boulder_Jul_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="07",])
summary(Boulder_Jul_TMAX.lm)
abline(coef(Boulder_Jul_TMAX.lm), col="red")
```

```{r Boulder TMAX averages for August}
#plot(Boulder_MonthlyTMAXMean$TMAX[Boulder_MonthlyTMAXMean$Month=="08"], ty='l')
plot(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="08",],
ty='l', xlim=c(1890, 2020))
Boulder_Aug_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="08",])
summary(Boulder_Aug_TMAX.lm)
abline(coef(Boulder_Aug_TMAX.lm), col="red")
```

```{r Boulder TMAX averages for September}
#plot(Boulder_MonthlyTMAXMean$TMAX[Boulder_MonthlyTMAXMean$Month=="09"], ty='l')
plot(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="09",],
ty='l', xlim=c(1890, 2020))
Boulder_Sep_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="09",])
summary(Boulder_Sep_TMAX.lm)
abline(coef(Boulder_Sep_TMAX.lm), col="red")
```

```{r Boulder TMAX averages for October}
#plot(Boulder_MonthlyTMAXMean$TMAX[Boulder_MonthlyTMAXMean$Month=="10"], ty='l')
plot(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="10",],
ty='l', xlim=c(1890, 2020))
Boulder_Oct_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="10",])
summary(Boulder_Oct_TMAX.lm)
abline(coef(Boulder_Oct_TMAX.lm), col="red")
```


