---
title: "Importing and Processing CSV Files"
author: "Marc Los Huertos"
date: "12/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing CSV Files

First, we will import the CSV file. In most cases, we don't need to preprocess the csv files. Use file.choose() to get your data...
```{r importingCSV}
# Importing Scripts

file = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/David Wagner/Hagerstown Maryland NOAA data.csv"

import = read.csv(file)
```

## Checking the Data

```{r checking}
plot(TMAX~DATE, import)
```

What's going on?  What is the deal with the -9999? These are used for missing data. We need to remove them!

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```

Okay, now we'll check again, for dates less then Dec. 31, 1913:
```{r}
plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```

Yikes! What's wrong? As it turns it the problem is that with how the dates are specified. In particular, the Dec 31 to Jan 1 transition: 

19131231 -> 19140101, if you use these as numbers it should be 19131232, 19131232, etc. You see we are missing lots numbers!


```{r}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

## Fixing the Dates

First, we convert the date a string or character values. Next, we'll convert the strings to a data format. 

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19931031,], ty='l')
```
## Subset Sites

Olivia has two sites in here data, so, we need to subset it.

```{r}
unique(import$STATION_NAME)
```

Let's choose the FAIRPLEX NY US because the record is longer than the airport.

```{r}
HAGERSTOWN <- subset(import, STATION_NAME=="HAGERSTOWN MD US", select=c(STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
```


## A new plot
```{r}
plot(TMAX~NewDate, HAGERSTOWN, ty='l')
```

## Evaluating the trend

Creating a linear model and then evaluate the rate (slope). 
```{r}
# Linear Model
HAGERSTOWN.lm <- lm(TMAX~NewDate, data=HAGERSTOWN)
summary(HAGERSTOWN.lm)
plot(TMAX~NewDate, HAGERSTOWN, ty='l')
abline(coef(HAGERSTOWN.lm), col='red')
```

## Creating Month and Year Variables

So, let's figure out how to see how changes happen for individual months.
```{r}
# Get months
HAGERSTOWN$Month = months(HAGERSTOWN$NewDate) # Creates problems.
HAGERSTOWN$Month = format(as.Date(HAGERSTOWN$NewDate), format = "%m")
HAGERSTOWN$Year = format(HAGERSTOWN$NewDate, format="%Y")
```

### Creating Monthly Means
```{r creatingmonthlymeans}
MonthlyMean = aggregate(TMAX ~ Month + Year, HAGERSTOWN, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(TMAX ~ Month + Year, HAGERSTOWN, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```


```{r plottingMonthyMeans}
plot(MonthlyMean$TMAX, ty='l')

```
below is sd
```{r plottingMonthySD}
plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(TMAX~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```

### Selecting for 1 Month -- May

Perhaps, we can get a better handle on this stuff if we analyze for just one month at a time -- certainly easier to visualize!

```{r September}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="09"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="09",], ty='l') 
September.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="09",])
summary(September.lm)

abline(coef(September.lm), col="red")

```

Now, the change is `r coef(May.lm)[2]` degress/year or `r round(coef(May.lm)[2]*100, 3)` degress/100 years (YIKES!) with a probability of `r round(summary(May.lm)$coefficients[2,4],4)`. Although we can't reject the null hypothesis, it is worrying!

## Next Steps

I think you should evaluate every month and see what happens. You might also consider looking at the TMIN as well. Could be important!

## Lets try Dec

```{r}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, HAGERSTOWN, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

2. Now we plot the mins, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="12"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degress/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temerpature, that year is one, is quite problematic.

4. What we have not determined is the cause. So, be careful when you describe the results, cause and effect cannot be analyzed using this method.


# Departure from Mean

import$PRCP[import$PRCP==-9999] = NA
```{r}

PRCP_mean = mean(HAGERSTOWN$PRCP, na.rm=T)

plot(PRCP~NewDate, data=HAGERSTOWN)
abline(h=PRCP_mean, col="blue")

```
import$PRCP[import$PRCP==-9999] = NA
PRCP_mean = mean(HAGERSTOWN$PRCP, na.rm=T)

plot(PRCP~NewDate, data=HAGERSTOWN)
abline(h=PRCP_mean, col="blue")
```{r}

YearlySum = aggregate(PRCP ~ Year, HAGERSTOWN, sum)

YearlyMean = mean(YearlySum$PRCP)
                  
YearlySum$Departure = YearlyMean - YearlySum$PRCP
YearlySum$YEAR = as.numeric(YearlySum$Year)

plot(PRCP~YEAR, data=YearlySum, las=1, ty="l")
abline(h=YearlyMean, col="blue")

lines(Departure ~ YEAR, data=YearlySum, col="red")

```