---
title: "Fire Blog"
author: "Marc Los Huertos"
date: "9/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ants and Fire

These data were collected to determine the impact of fire on ant in forests and a field (grassland). 

```{r ants}
#Data input
forest <- c(9,6,4,6,7,10)
field <- c(12,9,12,10)

forest
field
```

Now let's explore the data with a histogram...

```{r hist}
hist(forest)
hist(field)
```

Note: Does this look like a normal distribution


What's the difference in the means?  How do we know if these are signficant?

```{r meandif}

mean(forest) - mean(field)

```

Now, let's create a data frame with the data -- this form of data is useful for more complex analyses. 

```{r}
df <- data.frame(id=seq(1, 10, 1), 
              Habitat=c(rep("Forest", 6),rep("Field", 4)), 
              Colonies=c(forest, field)
              )
                 
df


boxplot(Colonies~Habitat, data=df, ylab="No. of Colonies")
```



## Getting the means and standard deviation

```{r}
mean(forest);sd(forest)
mean(field);sd(field)

mean(field)-mean(forest)
```

## Monte Carlo

And here we go!

```{r}
colonies <- c(forest,field)
colonies


(mean(forest)-mean(field))

```

Here's the monte carlo simulation...

1. First, we create matrix the "hold" the simulated data
```{r}
n <- 10000 # Let's do 10000 simulations!
sim <- matrix(,n,10) # let's create a matrix with 10000 x 10 dimensions
dif <- matrix(,n,) # this matrix has 10000 x 1 dimensions for the difference
str(dif)
str(sim)
```

2. Now let's randomly sample the colonies

```{r}
for(i in 1:n) {
sim[i,]<- sample(colonies) # draw a random sample of colonies per quadrat
dif[i,] <- abs(mean(sim[i,1:6])-mean(sim[i,7:10]))
}
# sim
hist(dif)
```



3. Let's figure out how many random combination means are lower than 3.75

```{r}
(3.75 > dif)[1:100]
sum(3.75 > dif)
```

4. 
```{r}
#sum(3.75 == dif)
sum(3.75 < dif)/n

# sum(3.75 <= dif)/n
```

## ANOVA

```{r}
ants.aov <- aov(Colonies ~ Habitat, data=df)
summary(ants.aov)

```

## ANOVA

#moisture <- c(.23,.34, .42, .36, .32, .44, .53, .43, .32, .23)
#Analysis of Variance
x<- seq(0,20,.2)
plot(x,dnorm(x, mean(Colonies), sd(Colonies)), ylim=c(0,.25), type="l")
lines(x,dnorm(x, mean(forest), sd(forest))/2,col="green")
lines(x,dnorm(x, mean(field), sd(field))/2,col="red")

model1 <- formula(Colonies~Habitat)
model1.aov <- aov(model1)
summary(model1.aov)
summary(aov(Colonies~Habitat))

1-pf(8.7805,1,8)

x<- seq(0,15,.1)
plot(x,df(x,1,8), ylim=c(0,.95), type="l")
lines(c(8.7805,8.7805),c(0,.2), col="red")
text(9.5,.2, "F-obs", col="red")

qf(.95,1,8)
lines(c(qf(.95,1,8),qf(.95,1,8)),c(0,.2), col="blue")
text(6.5,.2, "F-critical", col="blue")

# AIC
null <- lm(Colonies~1)
model1 <- lm(Colonies~Habitat)
#model2 <- lm(Colonies~Habitat)
AIC(null, model1, model2)

# Bayesian...maybe the end of the semester...

# checking assumptions
plot(model1.aov)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
