---
title: "Bryan R Markdown"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r readdata, include=FALSE}
filepath = "/home/CAMPUS/bawa2018/Climate_Change_Narratives/Data/FA20/Williams_SaltLakeCityUT_data.csv"
climate_data = read.csv(filepath)
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
```


```{r plotdailyTMAX, include=FALSE}
plot(TMAX~NewDate, climate_data, pch = 16, cex=.2, col = "blue")
TMAX.lm = lm(TMAX ~ NewDate, data = climate_data) 
coef(TMAX.lm)
abline(coef(TMAX.lm),col ="orange", lwd = 3)
```

```{r MonthlyPlot, include=FALSE}
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)
plot(MonthlyTMAXMean$TMAX, ty='l')
```

```{r MayPlot, include=FALSE}
#plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month=="05"], ty='l')
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],ty='l', xlim=c(1950, 2020))
May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
summary(May.lm)
abline(coef(May.lm), col="red")

```



```{r plotMinDailyTemp, include=FALSE}
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
# Fixing the Format of Month and Year as numeric
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
# First I create a vector of months
Months = c("January", "February", "March", "April",
"May", "June", "July", "August", "September", "October",
"November", "December")
# Create a panel so I can see all the figures at
# once.
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMAXresult <-NA; TMINresult <- NA
```

```{r TMAX, include=FALSE}
for (i in 1:12) {
    # plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month==i],
    # ty='l')
    plot(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), main = Months[i])
  Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ])
  summary(Month.lm)
  abline(coef(Month.lm), col = "red")
  TMAXresult <- rbind(TMAXresult, cbind(Months[i],
  round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared,3)))
  }
```

```{r TMIN, include=FALSE}
#par(mfrow=c(4,3),mar=c(5,4,1,1))
for (i in 1:12) {
MonthMin_lm <- lm(TMIN ~ YEAR, data=MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ])
TMINresult <- rbind(TMINresult, cbind(Months[i],round(coef(MonthMin_lm)[2], 4), round(summary(MonthMin_lm)$coefficients[2,4], 4), round(summary(MonthMin_lm)$r.squared, 3)))
summary(MonthMin_lm)
plot(MonthlyTMINMean$TMIN, ty='l', ylab='Monthly avg min temp', xlab='Years',main=Months[i]
)
abline(coef(MonthMin_lm),col='blue')
}
```
```{r Results, include=FALSE}
library(htmlTable)
Results <- data.frame(Month = TMINresult[c(2:13),1],
TMINSlope = TMINresult[c(2:13),2],
TMIN_P = as.numeric(TMINresult[c(2:13),3]),
TMINRsq = TMINresult[c(2:13),4],
TMAXSlope = TMAXresult[c(2:13),2],
TMAX_P = as.numeric(TMAXresult[c(2:13),3]),
TMAXRsq = TMAXresult[c(2:13),4])
Results$starTMIN = "NS"
Results$starTMIN[Results$TMIN_P <= .05] = "*"
Results$starTMIN[Results$TMIN_P < 0.01] = "**"
Results$starTMIN[Results$TMIN_P < 0.001] = "***"
Results$starTMAX = "NS"
Results$starTMAX[Results$TMAX_P < 0.05] = "*"
Results$starTMAX[Results$TMAX_P < 0.01] = "**"
Results$starTMAX[Results$TMAX_P < 0.001] = "***"
Results$TMINslope=paste(Results$TMINSlope, Results$starTMIN)
Results$TMAXslope=paste(Results$TMAXSlope, Results$starTMAX)
colnames(Results) <- c("Month", "2", "3", "R^2", "5", "6",
"R^2", "8", "9", "Slope TMIN", "Slope TMAX")
htmlTable(Results[,c(1, 10, 4, 11, 7)])
```
Based on my analysis, my data indicate that there is  a trend of increasing temperature in Salt Lake City for the month of July, rejecting the null hypothesis. (slope = 0.019, r2 = 0, p-value = 0.024).
```{r July TMAXPlot, echo=FALSE}
month_i <- 7
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$MONTH==month_i,],ty='l', xlim=c(1950, 2020), ylab="Maximum Average July Temperature")
July.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$MONTH==month_i,])
summary(July.lm)
abline(coef(July.lm), col="red")
```
Based on my analysis, my data indicate that there is not a trend of increasing temperature in Salt Lake City for the month of January, failing to reject the null hypothesis (slope = 0.0435, r2 = 0, p-value = 0.29).
```{r Jan TMIN Plot, echo=FALSE}
#plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month=="05"], ty='l')
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="05",],ty='l', xlim=c(1950, 2020), ylab="Minimum Average January Temperature")
Jan.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="05",])
summary(Jan.lm)
abline(coef(Jan.lm), col="cyan")
Janresult <- NULL
Janresult <- rbind(Janresult, cbind(Months[i],
  round(coef(Jan.lm)[2], 4), round(summary(Jan.lm)$coefficients[2, 4], 4), round(summary(Jan.lm)$r.squared,3)))
Janresult
```
Based on my analysis, my data indicate that there is a trend of decreasing snowfall in Salt Lake City for the month of January, rejecting the null hypothesis (slope = -0.0488, r2 = 0.2271, p-value = 0.02).
```{r Jan Snowfall, echo=FALSE}
i <- 1
MonthlySNOWMean = aggregate(SNOW ~ Month + Year, climate_data, mean)
MonthlySNOWMean$Month <- as.numeric(MonthlySNOWMean$Month)
MonthlySNOWMean$Year <- as.numeric(MonthlySNOWMean$Year)
plot(SNOW ~ Year, data = MonthlySNOWMean[MonthlySNOWMean$Month == i, ], ty = "l", las = 1, xlim = c(1940, 2020), main = Months[i], ylab="Average January Snowfall")
  SNOWMonth.lm <- lm(SNOW ~ Year, data = MonthlySNOWMean[MonthlySNOWMean$Month == i, ])
  summary(SNOWMonth.lm)
  abline(coef(SNOWMonth.lm), col = "cyan")
  SNOWresult <- NULL
  SNOWresult <- rbind(SNOWresult, cbind(Months[i],
  round(coef(SNOWMonth.lm)[2], 4), round(summary(SNOWMonth.lm)$coefficients[2, 4], 4), round(summary(SNOWMonth.lm)$r.squared,3)))
SNOWresult
```
