---
title: "Blog Template"
name: "Bryan Williams"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This blog will evaluate the effect of climate change on winters in the Salt Lake Valley, the effect on its ski resorts, and the implications for environmental justice efforts.

![View from Alta Ski Resort in Salt Lake City, Utah](https://cdn.vox-cdn.com/thumbor/p-x8T_IqgSdQrOzaVw_eZ5_XDQ0=/0x0:2400x1600/920x613/filters:focal(1008x608:1392x992):format(webp)/cdn.vox-cdn.com/uploads/chorus_image/image/66515430/merlin_2094703.7.jpg)

Variables examined will include:

* Temperature (degrees Celsius)
* Precipitation (mm/day)

```{r readdata}
### 1: Importing data
filepath = "/home/CAMPUS/bawa2018/Climate_Change_Narratives/Data/FA20/Williams_SaltLakeCityUT_data.csv"
climate_data = read.csv(filepath)

### 1.1: Cleaning date-string objects
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
```

#### Temperature exploration 

Below is a plot of all daily `TMAX` values over the full set of observations for Salt Lake City, UT.

```{r plotTMAXvstime,echo=F}
### 2: Plotting Temp vs. Date
plot(TMAX~NewDate, climate_data, pch = 16, cex=.2, col = "blue")
TMAX.lm = lm(TMAX ~ NewDate, data = climate_data) 
coef(TMAX.lm)
abline(coef(TMAX.lm),col ="orange", lwd = 3)
```
Yikes! That plot looks pretty horrible. Let's take a look at monthly averages of daily `TMAX`.

```{r creating_months, echo=FALSE}
### 3: Project guide 3, creating months to aggregate data
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")

### 3.1: Aggregating TMAX into monthly means
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

### 3.2: Visualize new monthly averages
plot(MonthlyTMAXMean$TMAX, ty='l')
```

#### January is cold (but getting less so)!

In Utah, it appears that our winter months have experienced warming. As such, I am going to drill down on the month of January. I've plotted January data below and have created a best-fit line (linear model/ordinary least squares model) for monthly January averages of daily `TMAX` against year to see if there is some type of discernible trend.

```{r JanPlot}
#plot(MonthlyTMAXMean£TMAX[MonthlyTMAXMean£Month=="05"], ty='l')
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",],ty='l', xlim=c(1950, 2020))
Jan.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="01",])
summary(Jan.lm)
abline(coef(Jan.lm), col="red")

```



```
