---
title: "A Look at Climate Change in Oklahoma"
author: "Claudia Chandra"
date: "02/20/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r importingCSV, echo=FALSE, results='hide'}
# Importing Scripts

file = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Data/Los Angeles_878024.csv"

import = read.csv(file)
```


```{r checking, echo=FALSE, results='hide'}
#plot(TMAX~DATE, import)
```


```{r missings, echo=FALSE}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```




```{r, echo=FALSE, results='hide'}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```




```{r, echo=FALSE, results='hide'}
unique(import$STATION_NAME)
```


```{r, echo=FALSE}
LosAngeles <- subset(import, STATION_NAME=="LOS ANGELES DOWNTOWN USC CA US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
```
 

```{r, echo=FALSE, results='hide'}
LosAngeles$Month = months(LosAngeles$NewDate) # Creates problems.
LosAngeles$Month = format(as.Date(LosAngeles$NewDate), format = "%m")
LosAngeles$Year = format(LosAngeles$NewDate, format="%Y")
```

```{r creatingmonthlymeans, echo=FALSE, results='hide'}
MonthlyMean = aggregate(TMAX ~ Month + Year, LosAngeles, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)
```


##Average TMAX - May 
```{r May, echo=FALSE}
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="05",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="05",])

abline(coef(May.lm), col="red")
```

##Average TMAX - June 
```{r June, echo=FALSE}
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="06",], ty='l') 
June.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="06",])

abline(coef(June.lm), col="red")
```

Change (degrees / year) : `r coef(May.lm)[2]` 
Degrees / 100 years :  `r round(coef(May.lm)[2]*100, 3)` 
Probability of `r round(summary(May.lm)$coefficients[2,4],4)`. 

##Average TMAX - November 
```{r Nov, echo=FALSE}
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="11",], ty='l') 
Nov.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="11",])
abline(coef(Nov.lm), col="blue")

```

##Average TMAX - December
```{r Dec, echo=FALSE}
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
abline(coef(Dec.lm), col="blue")
```


```{r, echo=FALSE, results='hide'}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, LosAngeles, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

##Average TMIN - November 
```{r Nov2, echo=FALSE}

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="11",], ty='l') 
Nov.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="11",])

abline(coef(Nov.lm), col="purple")

```

##Average TMIN - December 
```{r Dec2, echo=FALSE}
plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",])
#summary(Dec.lm)

abline(coef(Dec.lm), col="purple")

```


Slope: `r coef(Dec.lm)[2]` 
probability: `r round(summary(Dec.lm)$coefficients[2,4],7)`
r-squared: `r round(summary(Dec.lm)$r.squared, 3)`. 

what do the r-squared and probability values mean? 
Is this formula replicable? If not, how can i find the slope of the trend line for each of my graphs? 

##Average TMIN - May 
```{r May2, echo=FALSE}
plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="05",], ty='l') 
May.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="05",])
#summary(May.lm)

abline(coef(May.lm), col="orange")

```



##Average TMIN - June
```{r, echo=FALSE}
plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="06",], ty='l') 
June.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="06",])
#summary(June.lm)

abline(coef(June.lm), col="orange")

```

##Precipitation


```{r, echo=FALSE}
LosAngeles$PRCP[LosAngeles$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, LosAngeles, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```

each dot is a yearly mean calculated from the sum of all rainfall recorded that year 

blue line = mean 

red line = trend line 

Precipitation declining at -0.0085364 cm yr$^{-1}~$ or -0.09cm decade$^{-1}$ 


```{r, echo=FALSE}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="red")

n <- 5 
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
#lines(YearlySum$YEAR, y_lag, col="red")
```






