---
title: "Importing and Processing CSV Files"
author: "Marc Los Huertos"
date: "12/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing CSV Files

First, we will import the CSV file. In most cases, we don't need to preprocess the csv files. 
```{r importingCSV}
# Importing Scripts

file = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Data/Los Angeles_878024.csv"

import = read.csv(file)
```

## Checking the Data

```{r checking}
plot(TMAX~DATE, import)
```

What's going on?  What is the deal with the -9999? These are used for missing data. We need to remove them!

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```

Okay, now we'll check again, for dates less then Dec. 31, 1913:
```{r}
plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```

Yikes! What's wrong? As it turns it the problem is that with how the dates are specified. In particular, the Dec 31 to Jan 1 transition: 

19131231 -> 19140101, if you use these as numbers it should be 19131232, 19131232, etc. You see we are missing lots numbers!


```{r}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

## Fixing the Dates

First, we convert the date a string or character values. Next, we'll convert the strings to a data format. 

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```
## Subset Sites

Olivia has two sites in here data, so, we need to subset it.

```{r}
unique(import$STATION_NAME)
```

Let's choose the FAIRPLEX NY US because the record is longer than the airport.

```{r}
LosAngeles <- subset(import, STATION_NAME=="LOS ANGELES DOWNTOWN USC CA US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
```


## A new plot
```{r}
plot(TMAX~NewDate, LosAngeles, ty='l')
```

## Evaluating the trend

Creating a linear model and then evaluate the rate (slope). 
```{r}
# Linear Model
LosAngeles.lm <- lm(TMAX~NewDate, data=LosAngeles)
summary(LosAngeles.lm)
plot(TMAX~NewDate, LosAngeles, ty='l')
abline(coef(LosAngeles.lm), col='red')
```

## Creating Month and Year Variables

So, let's figure out how to see how changes happen for individual months.
```{r}
# Get months
LosAngeles$Month = months(LosAngeles$NewDate) # Creates problems.
LosAngeles$Month = format(as.Date(LosAngeles$NewDate), format = "%m")
LosAngeles$Year = format(LosAngeles$NewDate, format="%Y")
```

### Creating Monthly Means
```{r creatingmonthlymeans}
MonthlyMean = aggregate(TMAX ~ Month + Year, LosAngeles, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(TMAX ~ Month + Year, LosAngeles, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```


```{r plottingMonthyMeans}
plot(MonthlyMean$TMAX, ty='l')

```
below is sd
```{r plottingMonthySD}
plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(TMAX~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```

### Selecting for 1 Month -- May

Perhaps, we can get a better handle on this stuff if we analyze for just one month at a time -- certainly easier to visualize!

```{r May}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="05"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="05",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="05",])
summary(May.lm)





abline(coef(May.lm), col="red")



```
```{r June}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="06"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="06",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="06",])
summary(May.lm)

abline(coef(May.lm), col="red")
```

Now, the change is `r coef(May.lm)[2]` degress/year or `r round(coef(May.lm)[2]*100, 3)` degress/100 years (YIKES!) with a probability of `r round(summary(May.lm)$coefficients[2,4],4)`. Although we can't reject the null hypothesis, it is worrying!

## Next Steps

I think you should evaluate every month and see what happens. You might also consider looking at the TMIN as well. Could be important!

## Lets try Dec

```{r}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, LosAngeles, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

2. Now we plot the mins, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="12"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degress/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temerpature, that year is one, is quite problematic.

4. What we have not determined is the cause. So, be careful when you describe the results, cause and effect cannot be analyzed using this method.


# Departure from Mean

```{r}
PRCP_mean = mean(LosAngeles$PRCP, na.rm=T)

plot(PRCP~NewDate, data=LosAngeles)
abline(h=PRCP_mean, col="blue")

```


## Experimental Portion --- Precipitation

Precipiation might depend more on the departure from the mean (often referred as as normal, whatever that means!).  I think it's worth pursuing, but haven't finished the analysis yet.

First, we need a "mean" -- The IPCC uses 1961-1990 as a norm, I don't know what is the standard for California, so we should look that up.

Second, we need to remove the missing values and evalaute which years have complete years. If you are missing rainy months, then the whole year should be thrown out -- but what about partial years in the drought season?  

Third, we will need to decide what level of aggredation -- monthly, yearly, etc. 

Fourth, in CA the water year starts in Oct 1. Should we follow the same convention?


```{r}
LosAngeles$PRCP[LosAngeles$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, LosAngeles, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```

A yearly mean, based on the annual sum for the entire records. Not sure this is appropriate.

Figure has points of the yearly sum of rainfall and the blue line mean. The greenline is the trend and red line is a five year running average, I think!  I am still trying to understand what the code is doing.
```{r}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="green")

n <- 5
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
lines(YearlySum$YEAR, y_lag, col="red")
```

The model suggests that the precipitation is declines at a rate of `r coef(YearlySum.lm)[2]` cm yr$^{-1}~$, or `r round(coef(YearlySum.lm)[2]*10, 2)` cm decade$^{-1}$.

```{r}
summary(YearlySum.lm)
```

Other attempts that didn't work!
```{r}
YearlySum2 = subset(YearlySum, select=c(PRCP))
YearlyMean = mean(YearlySum2$PRCP)

#filter(YearlySum2)



#Doesn't work...
# ma <- function(x,n=5){filter(x,rep(1/n,n), sides=2)}
# ma(YearlySum$PRCP)  


yearlyts =ts(YearlySum2) 
str(yearlyts)




```
yearlyts.lag = lag(yearlyts, 5)[6:112,]; head(yearlyts.lag)

plot(yearlyts, lty = 1, col = 3)

lines(yearlyts.lag)

YearlySum$Departure = YearlyMean - YearlySum$PRCP
YearlySum$YEAR = as.numeric(YearlySum$Year)

plot(PRCP~YEAR, data=YearlySum, las=1, ty="l")
abline(h=YearlyMean, col="blue")

lines(Departure ~ YEAR, data=YearlySum, col="red")
