---
title: "The Other Side of Climate Change: Ojai Valley Cooling"
author: "Olivia Whitener"
date: "2/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Importing CSV Files

```{r importingCSV}
# Importing Scripts

file = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Olivia/Ojai Temperatures - EA30.csv"

import = read.csv(file)
```

## Checking the Data

```{r checking}
plot(TMAX~DATE, import)
```

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```

```{r}
plot(TMAX~DATE, import[import$DATE<19181231,], ty='l')
```


```{r}
plot(TMAX~DATE, import[import$DATE<19180102,], ty='l')
```

## Fixing the Dates


```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19180102,], ty='l')
```
## Subset Sites

```{r}
unique(import$STATION_NAME)
```

Let's choose the BARROW W POST W ROGERS AIRPORT AK US because the record is longer than the airport.

```{r}
LosAngeles <- subset(import, STATION_NAME=="OJAI CA US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
```


## A new plot
```{r}
plot(TMAX~NewDate, LosAngeles, ty='l')
```

## Evaluating the trend

Creating a linear model and then evaluate the rate (slope). 
```{r}
# Linear Model
LosAngeles.lm <- lm(TMAX~NewDate, data=LosAngeles)
summary(LosAngeles.lm)
plot(TMAX~NewDate, LosAngeles, ty='l')
abline(coef(LosAngeles.lm), col='red')
```

## Creating Month and Year Variables

So, let's figure out how to see how changes happen for individual months.
```{r}
# Get months
LosAngeles$Month = months(LosAngeles$NewDate) # Creates problems.
LosAngeles$Month = format(as.Date(LosAngeles$NewDate), format = "%m")
LosAngeles$Year = format(LosAngeles$NewDate, format="%Y")
```

### Creating Monthly Means
```{r creatingmonthlymeans}
MonthlyMean = aggregate(TMAX ~ Month + Year, LosAngeles, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(TMAX ~ Month + Year, LosAngeles, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```


```{r plottingMonthyMeans}
plot(MonthlyMean$TMAX, ty='l')

```
below is sd
```{r plottingMonthySD}
plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(TMAX~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```

### Selecting for 1 Month


```{r April}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="04"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="04",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="04",])
summary(May.lm)

abline(coef(May.lm), col="red")

```


```{r July}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="07"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="07",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="07",])
summary(May.lm)

abline(coef(May.lm), col="red")

```


```{r August}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="08"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="08",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="08",])
summary(May.lm)

abline(coef(May.lm), col="red")


```


```{r November}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="11"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="11",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="11",])
summary(May.lm)

abline(coef(May.lm), col="red")

```

```{r}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, LosAngeles, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

2. Now we plot the mins, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')
plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="01"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="01",], ty='l') 
Jan.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="01",])
summary(Jan.lm)

abline(coef(Jan.lm), col="red")

```



```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="03"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="03",], ty='l') 
March.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="03",])
summary(March.lm)

abline(coef(March.lm), col="red")

```


```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="05"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="05",], ty='l') 
May.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="05",])
summary(May.lm)

abline(coef(May.lm), col="red")

```


```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="06"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="06",], ty='l') 
June.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="06",])
summary(June.lm)

abline(coef(June.lm), col="red")

```


```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="07"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="07",], ty='l') 
July.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="07",])
summary(July.lm)

abline(coef(July.lm), col="red")

```


```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="08"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="08",], ty='l') 
August.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="08",])
summary(August.lm)

abline(coef(August.lm), col="red")

```


```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="09"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="09",], ty='l') 
Sept.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="09",])
summary(Sept.lm)

abline(coef(Sept.lm), col="red")

```


```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="10"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="10",], ty='l') 
Oct.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="10",])
summary(Oct.lm)

abline(coef(Oct.lm), col="red")

```


## Experimental Portion --- Precipitation

Precipiation might depend more on the departure from the mean (often referred as as normal, whatever that means!).  I think it's worth pursuing, but haven't finished the analysis yet.

First, we need a "mean" -- The IPCC uses 1961-1990 as a norm, I don't know what is the standard for California, so we should look that up.

Second, we need to remove the missing values and evalaute which years have complete years. If you are missing rainy months, then the whole year should be thrown out -- but what about partial years in the drought season?  

Third, we will need to decide what level of aggredation -- monthly, yearly, etc. 

Fourth, in CA the water year starts in Oct 1. Should we follow the same convention?


```{r}
LosAngeles$PRCP[LosAngeles$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, LosAngeles, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```

A yearly mean, based on the annual sum for the entire records. Not sure this is appropriate.

Figure has points of the yearly sum of rainfall and the blue line mean. The greenline is the trend and red line is a five year running average, I think!  I am still trying to understand what the code is doing.
```{r}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="green")

n <- 5
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
lines(YearlySum$YEAR, y_lag, col="red")
```

The model suggests that the precipitation is declines at a rate of `r coef(YearlySum.lm)[2]` cm yr$^{-1}~$, or `r round(coef(YearlySum.lm)[2]*10, 2)` cm decade$^{-1}$.

```{r}
summary(YearlySum.lm)
```