---
title: "Temperature Changes in NYC and Less Severe Winters"
author: "Eleanor Tahbaz"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

In this blog, I will discuss the rise in temperatures in and around new york city and its impact on two possible topic the impact on the four seasons and how they are all much warmer. 

Main Focus: less cold/severe winters

## The Four Seasons of NYC

Here I will explain what most people picture as a usual year in NYC and the differences in temperature and percipitation within the seasons

## What We Saw in 2019

Next, I will go into the seasonal temperatures we saw in 2019 in NYC and how the expected temperatures and weather patterns were very different.

For example: There was not a single big snow storm that had snow stay on the ground for more than 24 hours which is uncommon for NYC, especially in January and February

## this is a plot showing the increase in temperature in and around NYC
```{r}
file.path = "/home/CAMPUS/erta2020/EleanorTahbaz_NY,NY region_data.csv"
climate_data = read.csv("/home/CAMPUS/erta2020/EleanorTahbaz_NY,NY region_data.csv")
str(climate_data)
strDates <-as.character(climate_data$DATE)
climate_data$NewDate <-as.Date(strDates, "%m/%d/%Y")
plot(TMAX~DATE, climate_data)
```

## The Relationship Between Temperature Increases and Less Severe Winters


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
