---
title: "Regional Analysis"
author: "Olivia"
date: "9/8/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r readdata, echo=FALSE}
#file.choose()
filepath = "/home/CAMPUS/ocma2018/Climate_Change_Narratives/Data/FA20/olivia_chicagomidway_data.csv"
climate_data = read.csv(filepath) 
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%Y-%m-%d")
```

## Analysis for Chicago Midway

```{r plotdailydata}
plot(TMAX~NewDate, data=climate_data, ty='l')
TMAX.lm = lm(TMAX ~ NewDate, data=climate_data) 
slopeintercept = coef(TMAX.lm)
abline(coef(TMAX.lm), col = 'pink') 
lm(TMAX ~ NewDate, data=climate_data)
lm(TMAX ~ NewDate, data=climate_data)
TMAX.lm = lm(TMAX ~ NewDate, data= climate_data)
slopeintercept = coef(TMAX.lm)
plot(TMAX ~ NewDate, data= climate_data, las=1)
abline(slopeintercept, col="pink", lwd=3)
(TMAX.lm.sum = summary(lm(TMAX ~ NewDate, data=climate_data)))
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
```
## For this graph of daily temperatures across the years, the probability value is <2e-16. This tiny number is so close to zero, that we can reject the null hypothesis (the assumption that the slope is zero, or that there is not an upward trend) and safely assume that the coefficient of maximum temperature by year has a slope that is not zero. If the p value were closer to one, there would be a greater chance of variation from the model I have created.

## Month by Month Analysis
```{r plotmonthbymonth}
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],
ty='l', xlim=c(1950, 2020))
May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
summary(May.lm)
abline(coef(May.lm), col="pink")
Months = c("January", "February", "March", "April",
"May", "June", "July", "August", "September", "October",
"November", "December")
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMAXresult <- NA 
for (i in 1:12) {
plot(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH ==
i, ], ty = "l", las = 1, xlim = c(1940, 2020),
main = Months[i])
Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH ==
i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "pink")
TMAXresult <- rbind(TMAXresult, cbind(Months[i],
round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2,
4], 4), round(summary(Month.lm)$r.squared,
3)))
}
```


## As seen below, the maximum temperatures in November, from 1940 to 2020, display a rising trend. Having lived in Chicago all my life, fall tends to begin early, with temperatures dropping around November and October. However, as the graph below shows, autumn has been getting warmer and warmer. The P value here is 0.188, which, while not allowing the rejection of the null hypothesis, is not altogether improbable. (need some clarification here)

```{r NOVMAX}
i <- 11
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$Month <- as.numeric(MonthlyTMAXMean$Month)
MonthlyTMAXMean$Year <- as.numeric(MonthlyTMAXMean$Year)
plot(TMAX ~ Year, data = MonthlyTMAXMean[MonthlyTMAXMean$Month == i, ], ty = "l", las = 1, xlim = c(1940, 2020), main = Months[i])
  TMAXMonth.lm <- lm(TMAX ~ Year, data = MonthlyTMAXMean[MonthlyTMAXMean$Month == i, ])
  summary(TMAXMonth.lm)
  abline(coef(TMAXMonth.lm), col = "pink")
 TMAXresult <- NULL
  TMAXresult <- rbind(TMAXresult, cbind(Months[i],
  round(coef(TMAXMonth.lm)[2], 4), round(summary(TMAXMonth.lm)$coefficients[2, 4], 4), round(summary(TMAXMonth.lm)$r.squared,3)))
TMAXresult
```


## Minimum Temperature
```{r plottmin}
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
# Fixing the Format of Month and Year as numeric
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)
plot(MonthlyTMINMean$TMIN, ty='l')
library(xtable)
### 3.1: Specifying plotting window parameters
par(mfrow = c(4, 3), mar = c(5, 4, 1, 1) + 0.1) 
  TMINresult <- NULL 
for (i in 1:12) { 
Month_sub_DF <- MonthlyTMINMean[MonthlyTMINMean$MONTH==i, ] 
MonthMin_lm <- lm(TMIN ~ YEAR, data=Month_sub_DF)
TMINresult <- rbind(TMINresult, 
                      cbind(Months[i], 
                            round(coef(MonthMin_lm)[2], 3),
                            round(summary(MonthMin_lm)$coefficients[2,4], 3), 
                            round(summary(MonthMin_lm)$r.squared, 2)))
  plot(TMIN~YEAR,data=Month_sub_DF, type='l', ylab='Min T (*C)', xlab='Years',main=Months[i], bty="L", las=1) 
  abline(coef(MonthMin_lm),col="pink") 
}

#these didn't work - but leaving them in in case I missed something
#Results <- data.frame(Month = TMINresult[c(2:13),1],
#TMINSlope = TMINresult[c(2:13),2],
#TMIN_P = as.numeric(TMINresult[c(2:13),3]),
#TMINRsq = TMINresult[c(2:13),4],
#TMAXSlope = TMAXresult[c(2:13),2],
#TMAX_P = as.numeric(TMAXresult[c(2:13),3]),
#TMAXRsq = TMAXresult[c(2:13),4])
#Results$starTMIN = "NS"
#Results$starTMIN[Results$TMIN_P <= .05] = "*"
#Results$starTMIN[Results$TMIN_P < 0.01] = "**"
#Results$starTMIN[Results$TMIN_P < 0.001] = "***"
#Results$starTMAX = "NS"
#Results$starTMAX[Results$TMAX_P < 0.05] = "*"
#Results$starTMAX[Results$TMAX_P < 0.01] = "**"
#Results$starTMAX[Results$TMAX_P < 0.001] = "***"
#Results$TMINslope=paste(Results$TMINSlope, Results$starTMIN)
#Results$TMAXslope=paste(Results$TMAXSlope, Results$starTMAX)
#colnames(Results) <- c("Month", "2", "3", "R^2", "5", "6",
#"R^2", "8", "9", "Slope TMIN", "Slope TMAX")
#print(xtable(Results[,c(1, 10, 4, 11, 7)]))
```


## Similarly, a look at November's minimum temperatures from 1940 to 2020, these low temperature averages increase (shown by the positive slope of the line of best fit). This graph has a P value of 0.00234 or 0.002, which is still relatively close to zero, allowing us to reject the null hypothesis.

```{r NOVMIN}
i <- 11
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$Month <- as.numeric(MonthlyTMINMean$Month)
MonthlyTMINMean$Year <- as.numeric(MonthlyTMINMean$Year)
plot(TMIN ~ Year, data = MonthlyTMINMean[MonthlyTMINMean$Month == i, ], ty = "l", las = 1, xlim = c(1940, 2020), main = Months[i])
  TMINMonth.lm <- lm(TMIN ~ Year, data = MonthlyTMINMean[MonthlyTMINMean$Month == i, ])
  summary(TMINMonth.lm)
  abline(coef(TMINMonth.lm), col = "pink")
 TMINresult <- NULL
  TMINresult <- rbind(TMINresult, cbind(Months[i],
  round(coef(TMINMonth.lm)[2], 4), round(summary(TMINMonth.lm)$coefficients[2, 4], 4), round(summary(TMINMonth.lm)$r.squared,3)))
TMINresult
```


## Precipitation: 
```{r plotprecipitation}

climate_data$PRCP[climate_data$PRCP==-9999] <- NA
Missing <- aggregate(is.na(climate_data$PRCP),
list(climate_data$Month, climate_data$Year), sum)
Missing$Date = as.numeric(Missing$Group.1) + as.numeric(Missing$Group.2)/12
plot(x ~ Date, data=Missing)
TotalPPT <- aggregate(climate_data$PRCP,
list(climate_data$Month, climate_data$Year), sum, na.rm=T)
names(TotalPPT) = c("Group.1", "Group.2", "ppt")
NonMissing <- Missing[Missing$x < 5, c(1:3)]
library(dplyr)
PPT <- merge(TotalPPT, NonMissing, all.y=TRUE)
PPT$Date <- as.numeric(PPT$Group.1) + as.numeric(PPT$Group.2)/12
head(PPT)
PRCP_mean = mean(PPT$ppt)
plot(ppt~Date, data=PPT)
abline(h=PRCP_mean, col="pink")
MonthlyPRCPMean = aggregate( PRCP~ Month + Year, climate_data, sum)
MonthlyPRCPMean$YEAR = as.numeric(MonthlyPRCPMean$Year)
# Fixing the Format of Month and Year as numeric
MonthlyPRCPMean$YEAR = as.numeric(MonthlyPRCPMean$Year)
MonthlyPRCPMean$MONTH = as.numeric(MonthlyPRCPMean$Month)
head(MonthlyPRCPMean)
plot(MonthlyPRCPMean$PRCP, ty='l')
library(xtable)
# 3.1: Specifying plotting window parameters
par(mfrow = c(4, 3), mar = c(5, 4, 1, 1) + 0.1) 
  PRCPresult <- NULL 
 for (i in 1:12) { 
Month_sub_DF <- MonthlyPRCPMean[MonthlyPRCPMean$MONTH==i, ] 
MonthPrcp_lm <- lm(PRCP ~ YEAR, data=Month_sub_DF)
PRCPresult <- rbind(PRCPresult, 
                     cbind(Months[i], 
                            round(coef(MonthPrcp_lm)[2], 3), 
                           round(summary(MonthPrcp_lm)$coefficients[2,4], 3), 
                            round(summary(MonthPrcp_lm)$r.squared, 2))) 
  plot(PRCP~YEAR,data=Month_sub_DF, type='l', ylab='Min T (*C)', xlab='Years',main=Months[i], bty="L", las=1) 
  abline(coef(MonthMin_lm),col="pink") 
 }
```


## Snow Data will be included here, possibly replacing precipitation:

```{r plotsnow}
MonthlySnowDF = aggregate(SNOW ~ Month + Year, climate_data, sum) 
MonthlySnowDF$YEAR = as.numeric(MonthlySnowDF$Year)
# Fixing the Format of Month and Year as numeric
MonthlySnowDF$YEAR = as.numeric(MonthlySnowDF$Year)
MonthlySnowDF$MONTH = as.numeric(MonthlySnowDF$Month)
head(MonthlySnowDF)
plot(MonthlySnowDF$SNOW, ty='l')
library(xtable)
# 3.1: Specifying plotting window parameters
par(mfrow = c(4, 3), mar = c(5, 4, 1, 1) + 0.1) 
  SNOWresult <- NULL 
 for (i in 1:12) { 
Month_sub_DF <- MonthlySNOWMean[MonthlySnowDF$MONTH==i, ] 
MonthSnow_lm <- lm(SNOW ~ YEAR, data=Month_sub_DF)
SNOWresult <- rbind(SNOWresult, 
                     cbind(Months[i], 
                            round(coef(MonthSnow_lm)[2], 3), 
                           round(summary(MonthSnow_lm)$coefficients[2,4], 3), 
                            round(summary(MonthSnow_lm)$r.squared, 2))) 
  plot(SNOW~YEAR,data=Month_sub_DF, type='l', ylab='Min T (*C)', xlab='Years',main=Months[i], bty="L", las=1) 
  abline(coef(MonthSnow_lm),col="pink") 
 }

```


## Means by Month (not in graph form, as a summary)
```{r index}
tapply(MonthlyTMAXMean$TMAX,MonthlyTMAXMean$Month,function(x) {mean(x,na.rm=T)})
tapply(MonthlyTMINMean$TMIN,MonthlyTMINMean$Month,function(x) {mean(x,na.rm=T)})
tapply(MonthlyPRCPMean$PRCP,MonthlyPRCPMean$Month,function(x) {mean(x,na.rm=T)})
```