---
title: "SB Climate and Fires"
author: "By: Emily"
date: "2/11/2019"
output: html_document
---

# Why I Chose this Topic
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 After witnessing firsthand the incredible destruction that the Thomas Fire brought to my hometown, I became determined to bring awareness to the correlation between climate change, extreme weather, and natural disasters. The Santa Barbara County Multi-Jurisdictional Hazard Mitigation Plan (Hazard Mitigation Plan) categorizes Wildfire as a “high probability/medium impact hazard.” In this blog, I will compare climate data from the National Oceanic and Atmospheric Administration (NOAA) with wildfire data from the Hazard Mitigation Plan to prove my hypothesis that climate change is causing wildfires in Santa Barbara to become more frequent and intense.


# 
```{r,results="hide", echo=FALSE  }
filepath="/home/CAMPUS/ecka2018/Climate_Change_Narratives/EmilyKuhn_Santabarbara_data.csv"

climate_data = read.csv(filepath)
head(climate_data)
str(climate_data)

```

Santa Barbara is a coastal city in California that is vulnerable to a number of natural hazards that could be exacerbated by climate change. Some of the most impactful and likely of these events include drought/water shortage, flooding/sea level rise, and wildfires (Hazard Mitigation Plan).

Although I didn't have enough time to thoroughly show a correlation between all three of these threats and climate change, I will briefly discuss my reasons for believing that there is a relationship between them. 



```{r, results="hide", echo=FALSE }
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates,"%Y-%m-%d")

```

## SB April Climate since 1893
The first step that I took towards proving my hypothesis about wildfires was graphing Santa Barbara climate data that had been recorded from 1820 to the present. 
```{r, results="hide", echo=FALSE }
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)


plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",],
ty='l', xlim=c(1893, 2020), ylab="Max Temp (C)")
April.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="04",])
summary(April.lm)

abline(coef(April.lm), col="red")
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)

MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)
(MonthlyTMINMean)
Months=c("January", "February", "March", "April",
"May", "June", "July", "August", "September", "October",
"November", "December")
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMAXresult <- NA


library(xtable)

```


This graph shows the average temperature change in Santa Barbara during the month of April. The positive slope of 0.012 indicates that average temperature is increasing at a rate of about 1 degree per every century. We know that this is significant data because the P-value is less than 5%

# Wildfire Acres Burned
Next, I wanted to see if over time there would be an indication of increased intensity in Santa Barbara wildfires. So I used the wildfire data (Hazard Mitigation Plan) to graph the number of acres burned in the  major fires recorded since 1920. 

```{r,results="hide", echo=FALSE  }
firefile = "/home/CAMPUS/ecka2018/Climate_Change_Narratives/student_folders/Kuhn/WildfireAcres.csv"
wildfires = read.csv(firefile)

str(wildfires)
plot(Acres.Burned ~ Year, data= wildfires)

wildfires.lm = lm(Acres.Burned ~ Year, data= wildfires)
summary(wildfires.lm)
abline(coef(wildfires.lm))



```

Unfortunately, the data had so much variance in the number of acres burned per year that there was no real relationship found. 

However, I did notice that there seemed to be more fires in recent years, so I decided to graph the instances of major wildfires to see if they would show a significant increase since 1920. 

# Wildfire Instances
This data did indeed indicate a significant relationship and a stark increase in the the number of fires that ocurred each year. The slope of 0.083 indicates that in the 2000s there is estimated to be thirteen times the number of wildfires that ocurred in 1920. The R squared value (0.6421) indicates that about 64% of the variance in the data is explained by the line and the P-value of 0.0094 indicates that the data is significant. 

```{r,results="hide", echo=FALSE}

filerangepath= "/home/CAMPUS/ecka2018/Climate_Change_Narratives/student_folders/Kuhn/WildfireRanges.csv"
ranges=read.csv(filerangepath)

str(ranges)
fires.lm = lm(Fires~Time,data=ranges)

plot(Fires~Time,data=ranges)
abline(fires.lm, col="red")

summary(fires.lm)

```

The Thomas fire burned a total of 281,893 acres (CalFire.calgov), making it the largest fire in Califorinia's recorded history. This came on the heels of the 2012–2014 California drought which was cumulatively the worst unbroken drought of the last millennium. (...) Though the state of California declared the drought over in 2014, Santa Barbara remained in drought conditions until 2017 with its two main water sources at 14.9% capacity 0.2% capacity during 2016.



