---
title: "A Blog Thing"
author: "Toni Anderson"
date: "2/11/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Initial Climate Data

```{r}
##Read CSV Data
filepath= "/home/CAMPUS/ttab2018/Climate_Change_Narratives/student_folders/Anderson/tonianderson_burbank_realdata.csv"
climate_data = read.csv(filepath)
head(climate_data)
str(climate_data)
names(climate_data)

##Fix Dates
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")

##Plot Data
plot(TMAX~NewDate, climate_data, ty='l', main="TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="red" )
c <- coef(lm(TMAX~NewDate, climate_data))
abline(c, col="red")

##TMIN
plot(TMIN~NewDate, climate_data, ty='l', main="TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue" )
c <- coef(lm(TMIN~NewDate, climate_data))
abline(c, col="blue")
```

##Monthly Data Max and Mins

```{r}
## Define Shit
lm(TMAX~NewDate, data=climate_data)
summary(lm(TMAX~NewDate, data=climate_data))

##Monthly Averages
climate_data$Month = format(as.Date(climate_data$NewDate), format="%m")
climate_data$Year = format(as.Date(climate_data$NewDate), format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month +Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX, ty='l')

##Plot May
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",], ty='l', xlim=c(1950, 2020), main="May TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="blue")
May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
summary(May.lm)
abline(coef(May.lm), col="red")

##TMIN Definitions
plot(TMIN~NewDate, climate_data, ty='l', main="TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue" )
c <- coef(lm(TMIN~NewDate, climate_data))
abline(c, col="red")
MonthlyTMINMean = aggregate(TMIN ~ Month +Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)


##Every Month! Come back and do 1 by 1
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(3, 4), mar = c(5, 4, 3, 2) + 0.1 )
TMAXresult <- NA 
for (i in 1:12) {
  plot(TMAX ~ YEAR, data =  MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(15, 35),main = Months[i])
Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}

##Daily Mins (Decide which months [Oct, Nov, Dec; summer too?]
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(3, 4), mar = c(5, 4, 3, 2) + 0.1 )
TMINresult <- NA 
for (i in 1:12) {
  plot(TMIN ~ YEAR, data =  MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(0, 20),main = Months[i])
Month.lm <- lm(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMINresult <- rbind(TMINresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}

```
```{r}
##Precipitation Data (Initial Data)
filepath= "/home/CAMPUS/ttab2018/Climate_Change_Narratives/student_folders/Anderson/tonianderson_burbank_realdata.csv"
climate_data = read.csv(filepath)
head(climate_data)
str(climate_data)
names(climate_data)

##Fix Dates
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")

##Plot Data
plot(PRCP~NewDate, climate_data, ty='l', ylim=c(0, 300), main="Precipitation vs. Time(Years)", xlab="Time(Years)", ylab="Precipitation Celcius)", col.main="black", col.lab="green" )
c <- coef(lm(PRCP~NewDate, climate_data))
abline(c, col="red")

```
```{r}
##Precipitaion Data (Get sum, then create monthly plots)
## Define Shit
lm(PRCP~NewDate, data=climate_data)
summary(lm(PRCP~NewDate, data=climate_data))

##Monthly Sums
climate_data$Month = format(as.Date(climate_data$NewDate), format="%m")
climate_data$Year = format(as.Date(climate_data$NewDate), format="%Y")
MonthlyPRCPSum = aggregate(PRCP ~ Month +Year, climate_data, sum)
MonthlyPRCPSum$YEAR = as.numeric(MonthlyPRCPSum$Year)
MonthlyPRCPSum$MONTH = as.numeric(MonthlyPRCPSum$Month)
str(MonthlyPRCPSum)


##Plot May
plot(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="05",], ty='l', xlim=c(1950, 2020), main="May Precipitation vs. Time(Years)", xlab="Time(Years)", ylab="Precipitation", col.main="black", col.lab="blue")
May.lm <- lm(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="05",])
summary(May.lm)
abline(coef(May.lm), col="red")


##Every Month! Come back and do 1 by 1
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(3, 4), mar = c(5, 4, 3, 2) + 0.1 )
PRCPresult <- NA 
for (i in 1:12) {
plot(PRCP ~ YEAR, data =  MonthlyPRCPSum[MonthlyPRCPSum$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(0, 400), main = Months[i], xlab="Total Precipation", ylab="Year", col.main="black", col.lab="green" )
Month.lm <- lm(PRCP ~ YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
PRCPresult <- rbind(PRCPresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}

```
##TMAX Fall
```{r}
##TMAX Months

#October
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",], ty='l', xlim=c(1950, 2020), main="Oct TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="red")
Oct.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="10",])
summary(Oct.lm)
abline(coef(Oct.lm), col="red")
#November
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",], ty='l', xlim=c(1950, 2020), main="Nov TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="red")
Nov.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="11",])
summary(Nov.lm)
abline(coef(Nov.lm), col="red")
#December
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",], ty='l', xlim=c(1950, 2020), main="Dec TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="red")
Dec.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="12",])
summary(Dec.lm)
abline(coef(Dec.lm), col="red")
#As one!
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(2, 3), mar = c(5, 4, 3, 2) + 0.1 )
TMAXresult <- NA 
for (i in 10:12) {
  plot(TMAX ~ YEAR, data =  MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(15, 35), xlab="Time (Years)", ylab="TMAX (Degrees Celcius)", col.lab="red", main = Months[i])
Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}
```

```{r}
##TMIN Months

#October
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="05",], ty='l', xlim=c(1950, 2020), main="Oct TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue")
Oct.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="10",])
summary(Oct.lm)
abline(coef(Oct.lm), col="blue")
#November
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="11",], ty='l', xlim=c(1950, 2020), main="Nov TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue")
Nov.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="11",])
summary(Nov.lm)
abline(coef(Nov.lm), col="blue")
#December
plot(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="12",], ty='l', xlim=c(1950, 2020), main="Dec TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue")
Dec.lm <- lm(TMIN~YEAR, data=MonthlyTMINMean[MonthlyTMINMean$Month=="12",])
summary(Dec.lm)
abline(coef(Dec.lm), col="blue")
#As one!
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(2, 3), mar = c(5, 4, 3, 2) + 0.1 )
TMINresult <- NA 
for (i in 10:12) {
  plot(TMIN ~ YEAR, data =  MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(0, 20),xlab="Time (Years)", ylab="TMIN (Degrees Celcius)", col.lab="blue", main = Months[i])
Month.lm <- lm(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "blue")
TMINresult <- rbind(TMINresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}
```

```{r}
##Precipitation Months

#October
plot(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="10",], ty='l', xlim=c(1950, 2020), main="Oct Precipitation vs. Time(Years)", xlab="Time(Years)", ylab="Precipitation", col.main="black", col.lab="green")
Oct.lm <- lm(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="10",])
summary(Oct.lm)
abline(coef(Oct.lm), col="green")
#November
plot(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="11",], ty='l', xlim=c(1950, 2020), main="Nov Precipitation vs. Time(Years)", xlab="Time(Years)", ylab="Precipitation", col.main="black", col.lab="green")
Nov.lm <- lm(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="11",])
summary(Nov.lm)
abline(coef(Nov.lm), col="green")
#December
plot(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="12",], ty='l', xlim=c(1950, 2020), main="Dec Precipitation vs. Time(Years)", xlab="Time(Years)", ylab="Precipitation", col.main="black", col.lab="green")
Dec.lm <- lm(PRCP~YEAR, data=MonthlyPRCPSum[MonthlyPRCPSum$Month=="12",])
summary(Dec.lm)
abline(coef(Dec.lm), col="green")

#As one
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
 par(mfrow = c(2, 3), mar = c(5, 4, 3, 2) + 0.1 )
 PRCPresult <- NA 
 for (i in 10:12) {
plot(PRCP ~ YEAR, data =  MonthlyPRCPSum[MonthlyPRCPSum$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(0, 300), main = Months[i], xlab="Year", ylab="Total Precipitation", col.main="black", col.lab="green" )
Month.lm <- lm(PRCP ~ YEAR, data = MonthlyPRCPSum[MonthlyPRCPSum$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "green")
PRCPresult <- rbind(PRCPresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}
```

###Maximum Temperature Monthly Averages 

```{r echo=FALSE}
plot(TMAX~NewDate, climate_data, ty='l', main="TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="red" )
c <- coef(lm(TMAX~NewDate, climate_data))
abline(c, col="red")

```

This graph shows Burbank’s average monthly maximum temperatures from 1940 until today, representing the entirety of the data set. With a p-value of 0.015, it is safe to reject the null hypothesis, establishing a concrete relationship between time and temperature. Furthermore, with a positive slope of 4.77 x 10^-5^, this graph asserts that this relationship is positive, or increasing over time. Though the slope itself may seem small, it is very significant because in a graph with such varied data, even a small net increase has big impacts on future variation patterns.^8^ In climate data especially, where even half a degree can cause major environmental impacts, even relatively small trends have big consequences.^4^


###Minimum Temperature Monthly Averages 

```{r echo=FALSE}
##TMIN
plot(TMIN~NewDate, climate_data, ty='l', main="TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue" )
c <- coef(lm(TMIN~NewDate, climate_data))
abline(c, col="blue")

```

This graph also shows the monthly average temperature, but this time, it is showcasing the minimums. Once again we can reject the null hypothesis with a p-value of 0.001. The slope is 4.62 x 10^-5^; again a relatively small slope, but one with important implications. While TMIN is still increasing, it is doing so far less rapidly than TMAX, leading to an increasing gap between the daily TMAX and TMIN. This is the first indication of climate whiplash; as gaps become more pronounced and varied over time, there will be a large effect on precipitation consistency, therefore kickstarting whiplash climate patterns.^6^

###Maximum and Minimum Temperature Fall Data 

```{r echo=FALSE, fig.height=2.5}
##Fall Season
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(1, 3), mar = c(5, 4, 3, 2) + 0.1 )
TMAXresult <- NA 
for (i in 10:12) {
  plot(TMAX ~ YEAR, data =  MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(0, 35), xlab="Time (Years)", ylab="TMAX (Degrees Celcius)", col.lab="red", main = Months[i])
Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}
```

```{r echo=FALSE, fig.height=2.5}
##Fall Season
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(1, 3), mar = c(5, 4, 3, 2) + 0.1 )
TMINresult <- NA 
for (i in 10:12) {
  plot(TMIN ~ YEAR, data =  MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(0, 35),xlab="Time (Years)", ylab="TMIN (Degrees Celcius)", col.lab="blue", main = Months[i])
Month.lm <- lm(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "blue")
TMINresult <- rbind(TMINresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}
```


Looking at the data more closely, the trends become even more apparent. The fall season, shown through the three months of October, November, and December, all have a relatively consistent increase in daily average temperature over time. The fall season is particularly significant for climate whiplash because it is the annual transitional period between the wet and dry season, and, over time, has amassed a very volatile climate pattern as the fire season extends into the traditionally short and intense rainy season.^8^

However, in comparison, the trends for the minimum temperature have unique distinct and unique trends for each month during this season. October shows a consistent dramatic increase, November remains stagnant, while December actually shows a decrease over time. These trends prove the correlation seen in the yearly tables; the gap between maximum and minimum temperatures is becoming more pronounced, contributing to wild swings between warm, cold, wet, and dry weather.^8^ However, what is important to note here is the distinct difference in the trends month to month. These differences prove that there are more significant increases and decreases in overall temperature.  This variation is indicative of highly varied precipitation rates, therefore supporting the existence of climate whiplash in the LA  County region.^8^

The data becomes even more striking when coupled with the total monthly precipitation data of the fall season. While October is increasing precipitation and temperature, November is significantly decreasing after a series of giant spikes in the mid-1900s. December’s trend line remains stagnant, however, the plotted data points showcase a huge variation in precipitation year to year, with significant valleys and spikes all over the graph. Each month has incredible variation in their precipitation totals; however, what stands out to me is that precipitation spikes are not consistent month to month. For example, October’s precipitation spike in the early 2000s is followed by a distinct drop in November’s precipitation. This shows that the wet and dry seasons in LA County are constantly changing places, both year to year and month to month. This is the trademark sign of climate whiplash; and according to the graphs, this discord in rain schedule is only becoming more pronounced as the maximum and minimum temperatures change around it. Historically, this has been demonstrated through a series of multi-year droughts interrupted by very wet seasons.^8^  If these trends continue, the possibility for dangerous flooding, fires, and other disasters increases exponentially.^6^

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
