---
title: "Draft Blog"
author: "Emily Tomz"
date: "12/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open the file

```{r read file}
climate_data.csv ="/home/CAMPUS/erta2018/Climate_Change_Narratives/student_folders/Tomz/Tomz_Blog_v0_files/1957996_modified.csv"
# climate_data.csv ="/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/student_folders/Tomz/Tomz_Blog_v0_files/1957996_modified.csv"
climate_data <- read.csv(climate_data.csv)
summary(climate_data)
head(climate_data)
str(climate_data)
names(climate_data)
```



## Re-format Dates

```{r dates reformatting}
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")
head(climate_data)
```




## Maximum Temperature over Time

```{r max daily temps, echo=FALSE}
plot(TMAX ~ NewDate, data = climate_data, ty = "l", xlab = "Date", ylab = "Degrees Celcius",
     main = "Maximum Daily Temperature, Zurich, Switzerland")

TMAX.lm <- lm(TMAX ~ NewDate, data = climate_data)
summary(TMAX.lm)

abline(coef(TMAX.lm), col = "red")
legend(1910, 0, col = c(1), coef(TMAX.lm))
```
The p-value is very small, so we reject the null hypothesis. So, it is unlikely that this trend occured only by chance. From the regression, it's evident that the temperature is changing over time. It is decreasing with a slope of -1.416x10^-5 degrees C per year.


## Monthly Means

```{r max: monthly means}
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format = "%Y")

MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)

MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

str(MonthlyTMAXMean)
summary(TMAX.lm)

plot(MonthlyTMAXMean$TMAX, ty = 'l', ylab = "Degrees Celcius", main = "Maximum Mean Montly Temperature, Zurich, Switzerland")
abline(coef(TMAX.lm), col = "blue")
legend(900, 5, coef(TMAX.lm))
```

```{r max: month of may}
plot(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$Month == "05",], 
     ty = 'l', xlim = c(1900, 2020), ylab = "Degrees Celcius", xlab = "Date", 
     main = "May: Maximum Daily Temperature, Zurich, Switzerland")

May.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$Month == "05",])
summary(May.lm)

abline(coef(May.lm), col = "green4")
legend(1980, 23, col = c(1), coef(May.lm))
```

```{r max: all months}
# Create vector of months
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September",
           "October", "November", "December")

# Create panel to see all figures at once
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMAXresult <- NA
for (i in 1:12){
  plot(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i,],
       ty = 'l', las = 1, xlim = c(1940, 2020), main = Months[i])
  Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i,])
  summary(Month.lm)
  
  abline(coef(Month.lm), col = "purple")
  
  TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Month.lm)[2], 4),
                                        round(summary(Month.lm)$coefficients[2, 4], 4),
                                        round(summary(Month.lm)$r.squared, 3)))
}
```




////////////////////////////////////////////////////////////////////////////////////////////////
////////////////////////////////////////////////////////////////////////////////////////////////


## Minimum Temperatures over Time

```{r min: daily temps}
MonthlyTMINMean = aggregate(TMIN ~ Month + Year, climate_data, mean)

MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)

#Fix format of month and year as numeric
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$Month = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)

# Create plot of monthly average of daily min temps
plot(MonthlyTMINMean$TMIN, ty = 'l', ylab = "Degrees Celcius", main = "Monthly Minimum Daily Temperature, Zurich, Switzerland")
TMIN.lm <- lm(TMIN ~ NewDate, data = climate_data)
abline(coef(TMIN.lm), col = "red")
legend(900, -5, coef(TMIN.lm))
```

## Minimum Monthly Temperatures

```{r min: monthly means}
climate_data$Month = format(as.Date(climate_data$NewDate), format = "%m")
climate_data$Year = format(climate_data$NewDate, format = "%Y")

MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)

MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

str(MonthlyTMAXMean)
summary(TMAX.lm)

plot(MonthlyTMAXMean$TMAX, ty = 'l', ylab = "Degrees Celcius", main = "Maximum Mean Montly Temperature, Zurich, Switzerland")
abline(coef(TMAX.lm), col = "blue")
legend(900, 5, coef(TMAX.lm))
```

## * THIS BLOCK IS WHERE I AM GETTING STUCK WITH NEEDING FINITE YLIM VALUES *



```{r min: all months}
# Checking ot make sure dataframe exists and variable names make sense...
str(MonthlyTMINMean)
# Create vector of months
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September",
           "October", "November", "December")

# Create panel to see all figures at once
par(mfrow = c(4, 3), mar = c(5, 4, 3, 2) + 0.1)
TMINresult <- NA
for (i in 1:12){
#  plot(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$MONTH == i,],
#       ty = 'l', las = 1, xlim = c(1940, 2020), main = Months[i])
  
# Line 165 is created a subset with zero rows...MONTH doesn't exists in dataframe.

plot(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month == i,],
       ty = 'l', las = 1, xlim = c(1940, 2020), main = Months[i])
  
  Month.lm <- lm(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$Month == i,])
  #fixed month in line 173
  summary(Month.lm)
  abline(coef(Month.lm), col = "purple")
  
  TMINresult <- rbind(TMINresult, cbind(Months[i], round(coef(Month.lm)[2], 4),
                                        round(summary(Month.lm)$coefficients[2, 4], 4),
                                        round(summary(Month.lm)$r.squared, 3)))
  
  # typo in rbind...TMIX -> TMIN
}
```

```{r min: all months}
library(xtable)
Results <- data.frame(Month = TMINresult[c(2:13), 1],
                      TMINSlope = TMINresult[c(2:13), 2],
                      TMIN_P = as.numeric(TMINresult[c(2:13), 3]),
                      TMINRsq = TMINresult[c(2:13), 4],
                      TMAXSlope = TMAXresult[c(2:13), 2],
                      TMAX_P = as.numeric(TMAXresult[c(2:13), 3]),
                      TMAXRsq = TMAXresult[c(2:13), 4])

```