---
title: "A Blog Thing"
author: "Toni Anderson"
date: "2/11/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Initial Climate Data

```{r}
##Read CSV Data
filepath= "/home/CAMPUS/ttab2018/Climate_Change_Narratives/student_folders/Anderson/tonianderson_burbank_realdata.csv"
climate_data = read.csv(filepath)
head(climate_data)
str(climate_data)
names(climate_data)

##Fix Dates
strDates <- as.character(climate_data$DATE)
climate_data$NewDate <- as.Date(strDates, "%m/%d/%Y")

##Plot Data
plot(TMAX~NewDate, climate_data, ty='l', main="TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="blue" )
c <- coef(lm(TMAX~NewDate, climate_data))
abline(c, col="red")
```

##Monthly Data Max and Mins

```{r}
## Define Shit
lm(TMAX~NewDate, data=climate_data)
summary(lm(TMAX~NewDate, data=climate_data))

##Monthly Averages
climate_data$Month = format(as.Date(climate_data$NewDate), format="%m")
climate_data$Year = format(as.Date(climate_data$NewDate), format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month +Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)
str(MonthlyTMAXMean)
plot(MonthlyTMAXMean$TMAX, ty='l')

##Plot May
plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",], ty='l', xlim=c(1950, 2020), main="May TMAX(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMAX(Degrees Celcius)", col.main="black", col.lab="blue")
May.lm <- lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
summary(May.lm)
abline(coef(May.lm), col="red")

##TMIN Definitions
plot(TMIN~NewDate, climate_data, ty='l', main="TMIN(Degrees Celcius) vs. Time(Years)", xlab="Time(Years)", ylab="TMIN(Degrees Celcius)", col.main="black", col.lab="blue" )
c <- coef(lm(TMIN~NewDate, climate_data))
abline(c, col="red")
MonthlyTMINMean = aggregate(TMIN ~ Month +Year, climate_data, mean)
MonthlyTMINMean$YEAR = as.numeric(MonthlyTMINMean$Year)
MonthlyTMINMean$MONTH = as.numeric(MonthlyTMINMean$Month)
head(MonthlyTMINMean)


##Every Month! Come back and do 1 by 1
Months = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
par(mfrow = c(1, 1), mar = c(5, 4, 3, 2) + 0.1 )
TMAXresult <- NA 
for (i in 1:12) {
  plot(TMAX ~ YEAR, data =  MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(15, 35),main = Months[i])
Month.lm <- lm(TMAX ~ YEAR, data = MonthlyTMAXMean[MonthlyTMAXMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}

##Daily Mins (Get new Data!)
plot(MonthlyTMINMean$TMIN, ty='l')
library(xtable)
Results <- data.frame(Month = TMINresult[c(2:13),1], TMINSlope = TMINresult[c(2:13), 2],)
Results$starTMIN = "NS"
Results$starTMIN[Results$TMIN_P <= .05] = "*"
Results$starTMIN[Results$TMIN_P <= 0.01] = "**"
Results$starTMIN[Results$TMIN_P < 0.001] = "***"
Results$starTMAX = "NS"
Results$starTMAX[Results$TMAX_P < 0.05] = "*"
Results$starTMAX[Results$TMAX_P < 0.01] = "**"
Results$starTMAX[Results$TMAX_P < 0.001] = "***"
Results$TMINslope=paste(Results$TMINSlope, Results$starTMIN)
Results$TMAXslope=paste(Results$TMAXSlope, Results$starTMAX)
colnames(Results) <- c("Month", "2", "3", "R^2", "5", "6",
"R^2", "8", "9", "Slope TMIN", "Slope TMAX")
print(xtable(Results[,c(1, 10, 4, 11, 7)]))

```
```
for (i in 1:12) {
  plot(TMIN ~ YEAR, data =  MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), ylim=c(15, 35),main = Months[i])
Month.lm <- lm(TMIN ~ YEAR, data = MonthlyTMINMean[MonthlyTMINMean$MONTH == i, ])
summary(Month.lm)
abline(coef(Month.lm), col = "red")
TMINresult <- rbind(TMINresult, cbind(Months[i], round(coef(Month.lm)[2], 4), round(summary(Month.lm)$coefficients[2, 4], 4), round(summary(Month.lm)$r.squared, 3)))
}

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
