---
title: "Blog: Climate Change in Alvin, TX"
author: "Mireya Valencia"
date: "2/20/2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r importingCSV, echo=FALSE}
# Importing Scripts

file = "/home/CAMPUS/mvxu2015/Climate_Change_Narratives/Mireya/HoustonWeather.csv"

import = read.csv(file)
```



```{r checking, include=FALSE}
plot(TMAX~DATE, import)
```


```{r missings, include=FALSE}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```


```{r, include=FALSE}
plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```



```{r, include=FALSE}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```





```{r, include=FALSE}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```


```{r, include=FALSE}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```



```{r, include=FALSE}
unique(import$STATION)
```



```{r, include=FALSE}
Alvin <- subset(import, STATION=="GHCND:USC00410204", select=c(STATION, DATE, NewDate, TMIN, TMAX, PRCP))
```


```{r, include=FALSE}
plot(TMAX~NewDate, Alvin, ty='l')
```



```{r, include=FALSE}
# Linear Model
Alvin.lm <- lm(TMAX~NewDate, data=Alvin)
# summary(Alvin.lm)
# plot(TMAX~NewDate, Alvin, ty='l')
# abline(coef(Alvin.lm), col='red')
```


```{r, include=FALSE}
# Get months
Alvin$Month = months(Alvin$NewDate) # Creates problems.
Alvin$Month = format(as.Date(Alvin$NewDate), format = "%m")
Alvin$Year = format(Alvin$NewDate, format="%Y")
```

```{r creatingmonthlymeans, include=FALSE}
MonthlyMean = aggregate(TMAX ~ Month + Year, Alvin, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

# MonthlySD = aggregate(TMAX ~ Month + Year, Alvin, sd)

# MonthlySD$YEAR = as.numeric(MonthlySD$Year)
# MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlyMean$NewDate = MonthlyMean$YEAR + (MonthlyMean$MONTH - 1)/12

# head(MonthlySD)
```


```{r plottingMonthyMeans, include=FALSE, results='hide'}
plot(MonthlyMean$TMAX, ty='l')

```
Alternative Hypothesis: The maximum temperature in Alvin, TX has increaesd from 1998-2013.

Null hypothesis: The maximum temperature in Alvin, TX has decreased from 1998-2013. 

Below are the monthly mean temperatures in Alvin, TX from 1998-2013. There is an overall increasing trend in the data. However, with p=0.3042, we cannot reject nor accept our null hypothesis because the data is not statistically.
```{r plottingMonthlyMean}
# plot(Monthly$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlyMean, ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year")
SD.lm <- lm(TMAX~NewDate, data=MonthlyMean)
# summary(SD.lm)

abline(coef(SD.lm), col="red")

```
As shown in the graph above, there is a huge portion of missing data from about 1920-1960. A gap in data this large poses questions about the validity of the data set. To see if the trend exists independently of this missing data, we can look only at the data collected after 1960. 
### Subsetting the Data
```{r}
MonthlyMean2=subset(MonthlyMean, subset=YEAR>1962)

```

```{r plottingMonthySD}
# plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlyMean2, ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year")
MonthlyMean2.lm <- lm(TMAX~NewDate, data=MonthlyMean2)
# summary(MonthlyMean2.lm)

abline(coef(MonthlyMean2.lm), col="red")

```
As shown above, the mean monthly temperatures from 1970-2013 are still increasing, but with a smaller slope than the complete data set. With p-value=0.6986, we cannot reject or support the hypethesis. However, looking at the graph from 1970 on shows us that the increasng trend of max. temperatures persists regardless of missing data. 
```{r May, include=FALSE}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="05"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="05",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="05",])
summary(May.lm)

abline(coef(May.lm), col="red")

```
In order to analyze statistically significant data, we can look at maximum temperatures for each month rather than the entire year. 

In the month of December, for instance, the maximum monthly temperature increased(p=.015). We reject the null hypothesis, meaning that the maximum temperature of December increased from 1998-2013. 
```{r}
# plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='p')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year", main="December") 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
# summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```
```{r}
# plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='p')
plot(TMAX~YEAR, data=MonthlyMean2[MonthlyMean$Month=="12",], ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year", main="December") 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean2[MonthlyMean$Month=="12",])
# summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```

The months of October (p=.00405) and March (p=.0466) show similar trends of increasing maximum temperature and are statistically significant, therby rejecting the null hypothesis. 
```{r}
# plot(MonthlyMean$TMAX[MonthlyMean$Month=="10"], ty='p')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="10",], ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year", main="October") 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="10",])
# summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```
```{r}
# plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='p')
plot(TMAX~YEAR, data=MonthlyMean2[MonthlyMean$Month=="10",], ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year", main="October") 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean2[MonthlyMean$Month=="10",])
# summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


```{r}
# plot(MonthlyMean$TMAX[MonthlyMean$Month=="03"], ty='p')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="03",], ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year", main="March") 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="03",])
# summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```
```{r}
# plot(MonthlyMean$TMAX[MonthlyMean$Month=="03"], ty='p')
plot(TMAX~YEAR, data=MonthlyMean2[MonthlyMean$Month=="03",], ty='p', pch=20, ylab="Maximum Temperature", xlab= "Year", main="March") 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean2[MonthlyMean$Month=="03",])
# summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```



```{r, include=FALSE}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, Alvin, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```



```{r, include=FALSE}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="12"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```






```{r, include=FALSE}
PRCP_mean = mean(Alvin$PRCP, na.rm=T)

plot(PRCP~NewDate, data=Alvin)
abline(h=PRCP_mean, col="blue")

```


```{r, include=FALSE}
Alvin$PRCP[Alvin$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, Alvin, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```


```{r, include=FALSE}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="green")

n <- 5
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
lines(YearlySum$YEAR, y_lag, col="red")
```


```{r, include=FALSE}
summary(YearlySum.lm)
```


```{r, include=FALSE}
YearlySum2 = subset(YearlySum, select=c(PRCP))
YearlyMean = mean(YearlySum2$PRCP)

#filter(YearlySum2)



#Doesn't work...
# ma <- function(x,n=5){filter(x,rep(1/n,n), sides=2)}
# ma(YearlySum$PRCP)  


yearlyts =ts(YearlySum2) 
str(yearlyts)




```
yearlyts.lag = lag(yearlyts, 5)[6:112,]; head(yearlyts.lag)

plot(yearlyts, lty = 1, col = 3)

lines(yearlyts.lag)

YearlySum$Departure = YearlyMean - YearlySum$PRCP
YearlySum$YEAR = as.numeric(YearlySum$Year)

plot(PRCP~YEAR, data=YearlySum, las=1, ty="l")
abline(h=YearlyMean, col="blue")

lines(Departure ~ YEAR, data=YearlySum, col="red")
