---
title: "1st Plot Attempt"
author: "Siddharth Namachivayam"
date: "2/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plot of Albuquerque Data

Here is my plot of Albuquerque maximum temperatures overlayed with a best fit line:

```{r abq_plot, echo=FALSE}
climate_data=read.csv("/Users/siddharth/R_Projects/siddharth_abq_data.csv")
strDates=as.character(climate_data$DATE)
climate_data$DATES=as.Date(strDates,"%Y-%m-%d")
plot(TMAX~DATES, climate_data,ty='l',xlab="YEAR")
abline(lm(TMAX~DATES,climate_data),col="red")
title(main="TMAX vs YEAR")
```

Lets look at some of the results for this regression:

```{r regression results, echo=FALSE}
summary(lm(TMAX~DATES,climate_data))
```

Note that the p-value for the y-intercept is 2x10^-6 and the p-value for the slope of the line is .02. Both of these values are less than .05. In other words, we can say with 95% confidence that our temperature is increasing at a rate  of 1.356x10^-5 degrees C per year and has an intercept of 21.28 degrees C.  

Let's go ahead and clean up our original graph by looking at monthly averages one month at a time. Here's a graph of monthly averages for the month of May.

```{r monthly, echo=FALSE}
climate_data$Month = format(as.Date(climate_data$DATES), format = "%m")
climate_data$Year = format(climate_data$DATES, format="%Y")
MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, climate_data, mean)
MonthlyTMAXMean$YEAR = as.numeric(MonthlyTMAXMean$Year)
MonthlyTMAXMean$MONTH = as.numeric(MonthlyTMAXMean$Month)

plot(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",],
     ty='l')
May.lm=lm(TMAX~YEAR, data=MonthlyTMAXMean[MonthlyTMAXMean$Month=="05",])
abline(May.lm,col="red")
title(main="Average TMAX for May vs YEAR")
```


