% subsection{Climate Science Report}

\subsubsection{Rational}

Learning about climate science requires students to dig deep into meteorology, atmospheric chemistry, physics, biogeochemistry, etc. Thus, learning about the science requires a bit of patience and lots of hard work. 

This asssignment has been designed to give you the capacity to be familiar and develop some expertise in one topic of climate change and to become a resource to others in the class.

\subsubsection{Assignment}

Investigate the assigned topic so you can write a thoughtful and critical review of the topics. Be sure to include how data might be used to counter common arguments that critique climate change science. Submit a written summary of your research findings and their references. The paper should be less than 8 pages (double spaced). 

Please include the following sections:

\begin{itemize*}
  \item Historical Development.
  \item Data Sources and Methods of Analysis.
  \item Areas of Uncertainty.
  \item Politicization of Evidence and Conclusions;
\end{itemize*}
 
\subsubsection{Submission Format and Naming Convention}

Submit via \texttt{Sakai} using the following naming convention: Climate\_Science\_Report\_F17\_XXXXX.pdf, where the XXXXX refer to the five digit \href{https://github.com/marclos/Climate_Change_Narratives/raw/master/Admin/RandomNumbers.pdf}{assigned random numbers}.

\subsubsection{Climate Science Report Grading}

The Climate Science Reports will be grading using Table \ref{tab:climatesciencereportgrading}. 

\begin{table}[h]
\caption{Climate Science Grading Standards.}
\label{tab:climatesciencereportgrading}
\begin{tabular}{lll}\hline
Criteria                & Standard      & Percent \\
\hline\hline
Historical Development  & Completness   & 20\% \\
Data Sources and Methods  & Completeness  & 20\Logical Fallacies & Identified and Confronted   & 20\% \\
Critical Thinking & Advanced                    & 20\%\\
% Graphics          && 20\%\\
Writing           && 20\% \\
Submission        && 20\% \\
\hline
\end{tabular}
\end{table}

