%\subsection{Op-Ed 2}

\subsubsection{Rational}

Successful editorials requires two things: 1) a compelling and newsworthy opinion and 2) 

\subsubsection{Assignment}

Select a regional newspaper where you can submit your Op Ed. Learn the format and length allowed for a submission. 


Using the Op-Ed guidelines, write an Op-Ed to summarize 2-3 salient points from your Blog where you should:

\begin{itemize}
  \item Describe regional climate changes and predictions that include ecological impacts; 
  \item Cite instances of how various scientific arguments have been distorted and politicized;
  \item Identify how conventional scientific standards have been compromised and how arguments that might be based on distortions can be countered.
\end{itemize}

Write and submit the editorial that highlights the key aspects of climate change that you discovered and link that to a newsworthy item. Be sure to cite and provide the URL to your blog. Submit the editorial and include the 'proof of receipt' with you submssion. 


\subsubsection{Submission Format and Naming Convention}

Uses the Op-Ed guidelines, submit a draft Op-Ed via \texttt{Sakai}. Include a description of the local or regional papers that this Op-Ed might be submitted and several examples of Op-Eds that have discussed environmental issues.

Write an Op-Ed to propose what makes a good public product with respect to criticisms of climate science debates and criticisms. In other words, describe (2-3) ways that climate change skeptisism might misuse the data analysis and how one might prevent the misuse, be sure to cite your blog as an attepmpt to accomlish these goals. 

Submit Op-Ed to the appropriate regional or local paper.



\subsubsection{Grading of Published Op Ed 2}

The 2nd Op Ed will be grading using the criteria in Table \ref{tab:oped2grading}.

\begin{table}[h]
\centering
\caption{Published Op Ed Grading Criteria}
\label{tab:oped2grading}
\begin{tabular}{ll}\hline
Standard            & Percent \\ \hline\hline    
Accuracy            & 20\%  \\
Completeness        & 20\% \\
Clarity             & 20\% \\
Timeliness          & 20\% \\
Use of Technology   & 20\% \\ \hline
\end{tabular}
\end{table}




