%\subsection{Analysis of Regional Data}

\subsubsection{Rational}

EA students...need to...

Learning to analyze data requires a range of skills that include collecting, analyzing, and interpreting data. For our purposes, this portion of the class is what might traditionally understood as ``doing science.'' We will learn how to test a hypothesis and what it means if we reject the null hypothesis. We will create figures that can be used to communicate our results and finally, we interpret the results. 

Ultimately, this analysis will be used a template for our blogs and inform our second Opinion Editorials. 

\subsubsection{Assignment}

Using the resources supplied, it will be up to you to download, pre-process, and analyze a trend analysis using R -- where the slope, r$^2$, and probability are calculated\footnote{We will have to learn what these are to be able to explain our results! Be sure to ask lots of questions about the statistics so you appreciate this important topic that nearly every scientific field relies!} and explained. 

Using R studio, analyze a long-term climate record, create 3-4 figures that will be used to communicate these climate records, e.g. 100-year temperature \textbf{and} precipitation record for a specific region. Be sure to include language about the ``null'' hypothesis for your trend analysis.

\begin{enumerate*}
  \item Download and analyze data (i.e. make inferences) to create an public product; I have uploaded all the climate data on a network drive, \url{//fargo/classes/EA30-LosHuertos}{//fargo/classes/EA30-LosHuertos}.\footnote{I haven't been able to get the directory working consistently, so stay tuned on this.}
  
\end{enumerate*}
 that describes the methods (data sources), data quality, and trends. 


\subsubsection{Submission Format and Naming Convention}

As specified by the milestones (Table \ref{tab:milestones}), submit the draft analysis and results using Rstudio. 

The Rmd file (and the compiled html) should be saved the the 'student\_submissions' directory using the following naming convention:

\begin{centering}
Region\_XXXXX.Rmd 

Region\_XXXXX.html
\end{centering}

NOTE: Be sure the file still compiles. For example, you may need to change the path to the Data directory. 

Since the regional analysis has been down within Rstudio, you will use the version control procedures to commit and push your analysis onto the Github repository. Thus, be sure to commit and push your files so I have access to the files. 

\subsubsection{Data Analysis Grading}

The Data Analysis html files will be grading using the criteria in Table \ref{tab:datagrading}.

\begin{table}[h]
\caption{Summary of Data Analysis grading standards.}
\label{tab:datagrading}
\begin{tabular}{llc}\hline
Criteria            &   Standard    & Percent \\ \hline\hline
Records  & Compelling, e.g. Over 60 years & 10\% \\
Knowledge of Data & Limitations and Methods of Collection & 10\% \\
Analysis & p-values and $R^2$ reported  & 20\% \\
Analysis          & Validated Model     & 20\% \\
Interpretation    & Accurate, e.g. rejected null   & 10\% \\
Graphics          & Publishable Quality & 20\% \\
Accessible        & Pushed and named correctly  10\% \\
\hline
\end{tabular}
\end{table}