---
title: "Using R to Process Weather Data"
author: "Marc Los Huertos"
date: "7/20/2016"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyr)
library(dplyr)
library(stringr)
library(reshape2)
source("summarySE.R")
```
# R Resources for Climate Data

## Tutorials and Libraries

Numerous resources are available that might help us (beside this document!) access and process weather data. 

Here is a preliminary list: 

* [R Blogger to Access, Clean, and Plot NOAA tempeature data](http://www.r-bloggers.com/accessing-cleaning-and-plotting-noaa-temperature-data/)
* Next...suggestions?

# Accessing the Data

## What sites are available for downloading the data?

* [NCDC--NOAA Datasets](https://www.ncdc.noaa.gov/cdo-web/datasets)

For this project, the Global Historical Climate Network may be the most useful data, which can be downloaded from the following an [ftp site](ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/daily/).

## How to download data

### Identify URL for data source

Once finding data sources, you might find URLs with ftp or http protocols -- it shouldn't matter which type you find. 

For example, using the NOAA's website, I found the following URL, [ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/daily/](ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/daily/), to download data from the Global Historical Climatalogical Network (GHCDN), where an entire dataset can be downloaded.  

## Regional Datasets

Since we are interested in a more regional approach, you may want to find a less daunting file.

Copy the URL into a clipboard and then create a object with the link:

```{r sourcefile}
sourcefile ="TBD??"
```

### Create a Folder for Data
Once a site URL has been identified, we can use R to download the data into a specific directory. I suggest you create a subdirectory within Data that you can track and store your data. 

### Download the Data

Note: The site can stall, so be careful when trying to download large files via web browers.

Alternatively, you could choose to download the entire dataset, as I have already done. First, I downloaded the file to my computer, then uploaded it into the R project Data directory.

```{r tarfile, echo=T}
tarfile = "ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/daily/ghcnd_all.tar.gz"
```

But this file needs to have some pre-processing before we can do anything with it.

# Pre-Processing Data

## How to Uncompress Data?

Files storing centuries of weather data from thousands of sites can become quite large. Thus, the files are compressed. There are a number of compression methods, available, e.g. zip, gz, and tar. Fortunately, we can use R to uncompress a range of compressed file types.

Examples of functions to uncompress data:

```{r uncompress, eval=FALSE}
# Uncompress the files.
untar(tarfile, files="Data")
```

It takes a while (2+ hours) to uncompress, so I have done that already and loaded some of the data into the github site.

# Reading Data into R

## What is the data format?

Before reading data, we need to determine the file structure. For example, XX data has the following format where each column is separated by tabs, which are lost in text rendions, so it looks a bit chaotic.


AG000060680  22.8000    5.4331 1362.0    TAMANRASSET                    GSN     60680

## Identifying the file path
```{r findfilepath}
stationfile = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Data/ghcnd-stations.txt"
```

## Reading Data into R

ID            1-11   Character
YEAR         12-15   Integer
MONTH        16-17   Integer
ELEMENT      18-21   Character
VALUE1       22-26   Integer
MFLAG1       27-27   Character
QFLAG1       28-28   Character
SFLAG1       29-29   Character
VALUE2       30-34   Integer
MFLAG2       35-35   Character
QFLAG2       36-36   Character
SFLAG2       37-37   Character
  .           .          .
  .           .          .
  .           .          .
VALUE31    262-266   Integer
MFLAG31    267-267   Character
QFLAG31    268-268   Character
SFLAG31    269-269   Character

```{r readinginputdata}
# read.table(stationfile, header=F, fill=T, row.names=NULL); head(stations)
stations = (read.fwf(stationfile, fill=T, widths= c(11, 9, 10, 7, 3, 32, 3, 4, 9), ))
```

# Processing Imported Data

## Creating Variable Names

```{r creatingvariablenames}
names(stations)= c("ID", "LAT", "LONG", "ELEV", "STATE", "NAME", "GSN", "HCN_CRN", "WHOID")

head(stations)
```

# Selecting and Example Location

## What is the structure of imported data?

Here's what the data look like:

```{r dataframestructure}
str(stations)
```

## Records from One Location
Here's an example of data from Arizona...

```{r Arizona}
stations[stations$ID=="US1AZMR0019",] 
# head(stations[stations$HCN_CRN==" CRN",])
```

Let's get the a different site into R

# Processing Data

## Creating Loops
I often forget how to make loops, so I often use simple examples that help me remember, for example, 

```{r practiceloops}
# practicing loops
for (year in c(2010,2011,2012,2013,2014,2015)){
  print(paste("The year is", year))
}
```

Since the data have a re-occuring set of variable names, I decided to create a vector of variable names, many of which are nearly the same. So, as you'll see, I had to create a loop to avoid having to type a ton (or 31 :-)) of different variables.

```{r newvariables}
# Create New Varible Names
MFLAG=NA; QFLAG=NA; SFLAG=NA; VALUE=NA
for (i in 1:31){
VALUE[i] = paste("DATE", i, sep="")
MFLAG[i] = paste("MFLAG", i, sep="")
QFLAG[i] = paste("QFLAG", i, sep="")
SFLAG[i] = paste("SFLAG", i, sep="")
}

# Vector of variable names converted from a transposed matrix
tmp = as.vector(t(matrix(data=c(VALUE, MFLAG, QFLAG, SFLAG), ncol=4)))
Names = c("ID", "YEAR", "MONTH", "ELEMENT", tmp); length(Names)
```

# Process Selected Data Files

## Reading .dly file into R

Now that we have a good idea of what these files look like, I have created a loop to read the files and select the rows where the maximum temperature was recorded ("TMAX"). 
```{r processing}
setwd("/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Data")

dly_list = list.files(pattern="*.dly"); head(dly_list)
#for (i in 1:length(dly_list)) 
for (i in 1:1){
tmp <- read.fwf(dly_list[i], widths = c(11, 4, 2, 4, rep(c(5, 1, 1, 1),31)))
names(tmp) <- Names
assign(dly_list[i], subset(tmp, ELEMENT=="TMAX", select=c(1:4, seq(5, by = 4, length.out=31))))
}
```

Once the file has been read, then I modiefied the file to make it useful, where I needed to re-arrange the file so we don't have a months across the top, but as rows. 

```{r}
tmp1 = melt(AGM00060515.dly, id=c("ID", "YEAR", "MONTH", "ELEMENT"))
head(tmp1)
tmp1$Day = as.numeric(str_sub(tmp1$variable,6,7)); head(tmp1)
tmp1$value[tmp1$value==-9999] = NA; head(tmp1)
tmp1$Temperature = tmp1$value/10

drops <- c("variable","value")
tmp1 <-tmp1[ , !(names(tmp1) %in% drops)]
tmp1$DECADE = round(tmp1$YEAR, -1)
# names(tmp1)
```

Now the data is ready to be visualized!

# Presenting the Results

## Summarizing the Results

```{r summarySE}
# called summarySE function
summarydf <- summarySE(tmp1, "Temperature", "DECADE", na.rm=T)
```

## Using ggplot2

The ggplot2 library is relatively new to me, so I am sure we can make some icremental improvements as the semester progresses. First, let's load the library.

```{r ggplot2library}
library(ggplot2)
```

## Creating a Figure 

```{r presenting}
# Think the color=DECADE thing can be deleted, but I haven't tried it yet. In any case, the legend is lame and I need to get rid of it!

ggplot(summarydf, aes(y=Temperature, x=DECADE, color= DECADE)) + geom_errorbar(aes(ymin=Temperature-se, ymax=Temperature+se), width=.2) + geom_line() + geom_point()

```

# Creating a Shiny App

Now we need to create a shiny app!


## NOAA dataset

New NOAA Directory -- ftp://ftp.ncdc.noaa.gov/pub/data/noaa/

```{r otherstuff}
library(raster)
library(XML)

coords.fwt <- read.fwf("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.txt",widths=c(6,1,5,1,38,7,1,8,9,8,1,8),sep=";",skip=22,fill=T)
Names = c("USAF", "X1", "WBAN", "X2", "STATION_NAME", "X3", "CTRY", "X4", "ST", "X5", "CALL", "X6", "LAT", "X7", "LON", "X8", "ELEV", "X9", "BEGIN", "X10", "END")
Widths = c(6,       1,    5,      1,        29,         1,    2,      3,    2,    1,    4,      1,    8,     1,     8,    1,    7,     1,     8,      1,    8)

coords.fwt <- read.fwf("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.txt",widths=Widths,sep=";",skip=22,fill=T); names(coords.fwt)=Names; coords.fwt[c(30,4000,20000),]

coords <- data.frame(ID=paste(as.factor(coords.fwt[,1])),WBAN=paste(as.factor(coords.fwt[,3])),Lat=as.numeric(paste(coords.fwt$LAT)),Lon=as.numeric(paste(coords.fwt$LON)));  coords[c(30,4000,20000),]
```


# NOAA Locations

```{r NOAApoints}
plot(Lat ~ Lon, data=coords, xlim=c(-180, 180) )
```

