---
title: "Climate Trends in Barrow, Alaska"
author: "Katie Graham"
date: "2/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r importingCSV, include=FALSE}
# Importing Scripts

file = "/home/CAMPUS/kygj2015/Climate_Change_Narratives/Data/KatieGraham - Barrow, Alaska.csv"

import = read.csv(file)
```

```{r checking, include=FALSE} 
#plot(TMAX~DATE, import)
```

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA

import$SNOW[import$SNOW==-9999] = NA
import$SNWD[import$SNWD==-9999] = NA
```

```{r, include=FALSE}
#plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```

```{r, include=FALSE}
#plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

```{r, include=FALSE}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```
```{r, include=FALSE}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```

```{r, include=FALSE}
#unique(import$STATION_NAME)
```

```{r, include=FALSE}
BarrowAlaska <- subset(import, STATION_NAME=="BARROW W POST W ROGERS AIRPORT AK US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP, SNOW, SNWD))
```

```{r, include=FALSE}
plot(TMAX~NewDate, BarrowAlaska, ty='l')
```

```{r, include=FALSE}
# Linear Model
BarrowAlaska.lm <- lm(TMAX~NewDate, data=BarrowAlaska)
summary(BarrowAlaska.lm)
plot(TMAX~NewDate, BarrowAlaska, ty='l')
abline(coef(BarrowAlaska.lm), col='red')
```

```{r, include=FALSE}
# Get months
BarrowAlaska$Month = months(BarrowAlaska$NewDate) # Creates problems.
BarrowAlaska$Month = format(as.Date(BarrowAlaska$NewDate), format = "%m")
BarrowAlaska$Year = format(BarrowAlaska$NewDate, format="%Y")
```

```{r creatingmonthlymeans, include=FALSE}
MonthlyMean = aggregate(TMAX ~ Month + Year, BarrowAlaska, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(TMAX ~ Month + Year, BarrowAlaska, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```


```{r plottingMonthyMeans, include=FALSE}
plot(MonthlyMean$TMAX, ty='l')

```
```{r plottingMonthySD, include=FALSE}
plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(TMAX~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```


### Selecting for 3 Months -- December, January, February

```{r December}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
December.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
summary(December.lm)

abline(coef(December.lm), col="red")

```

Now, the change is `r coef(December.lm)[2]` degrees/year or `r round(coef(December.lm)[2]*100, 3)` degrees/100 years with a probability of `r round(summary(December.lm)$coefficients[2,4],4)`. We can reject the null hypothesis that there is no relationship, since the p value is less than 0.05.

```{r January}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="01"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="01",], ty='l') 
January.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="01",])

summary(January.lm)

abline(coef(January.lm), col="red")

```

Now, the change is `r coef(January.lm)[2]` degrees/year or `r round(coef(January.lm)[2]*100, 3)` degrees/100 years with a probability of `r round(summary(January.lm)$coefficients[2,4],4)`. We can reject the null hypothesis that there is no relationship, since the p value is less than 0.05.

```{r February}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="02"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="02",], ty='l') 
February.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="02",])
summary(February.lm)

abline(coef(February.lm), col="red")

```

Now, the change is `r coef(February.lm)[2]` degrees/year or `r round(coef(February.lm)[2]*100, 3)` degrees/100 years with a probability of `r round(summary(February.lm)$coefficients[2,4],4)`. We can reject the null hypothesis that there is no relationship, since the p value is less than 0.05.

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, BarrowAlaska, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

2. Now we plot the mins, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="12"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degrees/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-squared of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temperature, that year is one, is quite problematic.

4. What we have not determined is the cause. So, be careful when you describe the results, cause and effect cannot be analyzed using this method.

## TMAX

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMAX = aggregate(TMAX ~ Month + Year, BarrowAlaska, mean)

MonthlyMeanTMAX$YEAR = as.numeric(MonthlyMeanTMAX$Year)
head(MonthlyMeanTMAX)
```

2. Now we plot the maximums, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMAX$TMAX, ty='l')

plot(MonthlyMeanTMAX$TMAX[MonthlyMeanTMAX$Month=="12"], ty='l')

plot(TMAX~YEAR, data=MonthlyMeanTMAX[MonthlyMeanTMAX$Month=="12",], ty='l') 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMeanTMAX[MonthlyMeanTMAX$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degrees/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-squared of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temperature, that year is one, is quite problematic.

4. What we have not determined is the cause. So, be careful when you describe the results, cause and effect cannot be analyzed using this method.


# Departure from Mean

```{r}
PRCP_mean = mean(BarrowAlaska$PRCP, na.rm=T)

plot(PRCP~NewDate, data=BarrowAlaska)
abline(h=PRCP_mean, col="blue")

```


## Experimental Portion --- Precipitation

Precipiation might depend more on the departure from the mean (often referred as as normal, whatever that means!).  I think it's worth pursuing, but haven't finished the analysis yet.

First, we need a "mean" -- The IPCC uses 1961-1990 as a norm, I don't know what is the standard for California, so we should look that up.

Second, we need to remove the missing values and evalaute which years have complete years. If you are missing rainy months, then the whole year should be thrown out -- but what about partial years in the drought season?  

Third, we will need to decide what level of aggredation -- monthly, yearly, etc. 

Fourth, in CA the water year starts in Oct 1. Should we follow the same convention?


```{r}
BarrowAlaska$PRCP[BarrowAlaska$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, BarrowAlaska, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```

A yearly mean, based on the annual sum for the entire records. Not sure this is appropriate.

Figure has points of the yearly sum of rainfall and the blue line mean. The greenline is the trend and red line is a five year running average, I think!  I am still trying to understand what the code is doing.
```{r}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="green")

n <- 5
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
lines(YearlySum$YEAR, y_lag, col="red")
```

The model suggests that the precipitation is declines at a rate of `r coef(YearlySum.lm)[2]` cm yr$^{-1}~$, or `r round(coef(YearlySum.lm)[2]*10, 2)` cm decade$^{-1}$.

```{r}
summary(YearlySum.lm)
```

Other attempts that didn't work!
```{r}
YearlySum2 = subset(YearlySum, select=c(PRCP))
YearlyMean = mean(YearlySum2$PRCP)

#filter(YearlySum2)



#Doesn't work...
# ma <- function(x,n=5){filter(x,rep(1/n,n), sides=2)}
# ma(YearlySum$PRCP)  


yearlyts =ts(YearlySum2) 
str(yearlyts)




```
yearlyts.lag = lag(yearlyts, 5)[6:112,]; head(yearlyts.lag)

plot(yearlyts, lty = 1, col = 3)

lines(yearlyts.lag)

YearlySum$Departure = YearlyMean - YearlySum$PRCP
YearlySum$YEAR = as.numeric(YearlySum$Year)

plot(PRCP~YEAR, data=YearlySum, las=1, ty="l")
abline(h=YearlyMean, col="blue")

lines(Departure ~ YEAR, data=YearlySum, col="red")


## SNWD - Snow Depth (mm)

1. We create a monthly mean for each month.

```{r} 
plot(SNWD~DATE, import)
```
```{r} 
plot(SNOW~DATE, import)
```



```{r}
plot(SNWD~DATE, import[import$DATE<19131231,], ty='l')
```

```{r}
plot(SNOW~DATE, import[import$DATE<19130102,], ty='l')
```

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```
```{r}
plot(SNWD~NewDate, import[import$DATE<19130102,], ty='l')
```

```{r}
# Linear Model
BarrowAlaska.lm <- lm(SNOW~NewDate, data=BarrowAlaska)
summary(BarrowAlaska.lm)
plot(SNOW~NewDate, BarrowAlaska, ty='l')
abline(coef(BarrowAlaska.lm), col='red')
```

```{r}
MonthlyMeanSNWD = aggregate(SNWD ~ Month + Year, BarrowAlaska, mean)

MonthlyMeanSNWD$YEAR = as.numeric(MonthlyMeanSNWD$Year)
head(MonthlyMeanSNWD)
```

```{r}
plot(SNOW~NewDate, BarrowAlaska, ty='l')
```

```{r}
# Linear Model
BarrowAlaska.lm <- lm(SNOW~NewDate, data=BarrowAlaska)
summary(BarrowAlaska.lm)
plot(SNOW~NewDate, BarrowAlaska, ty='l')
abline(coef(BarrowAlaska.lm), col='red')
```

```{r}
# Get months
BarrowAlaska$Month = months(BarrowAlaska$NewDate) # Creates problems.
BarrowAlaska$Month = format(as.Date(BarrowAlaska$NewDate), format = "%m")
BarrowAlaska$Year = format(BarrowAlaska$NewDate, format="%Y")
```

```{r}
MonthlyMean = aggregate(SNOW ~ Month + Year, BarrowAlaska, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Frank -- change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(SNOW ~ Month + Year, BarrowAlaska, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```


```{r}
plot(MonthlyMean$SNOW, ty='l')

```

```{r}
plot(MonthlySD$SNOW, ty='l')

plot(SNOW~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(SNOW~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```

### Selecting for 3 Months -- December, January, February

```{r December2}
plot(MonthlyMean$SNOW[MonthlyMean$Month=="12"], ty='l')
plot(SNOW~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
December.lm <- lm(SNOW~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
summary(December.lm)

abline(coef(December.lm), col="red")

```

```{r January2}
plot(MonthlyMean$SNOW[MonthlyMean$Month=="01"], ty='l')
plot(SNOW~YEAR, data=MonthlyMean[MonthlyMean$Month=="01",], ty='l') 
January.lm <- lm(SNOW~YEAR, data=MonthlyMean[MonthlyMean$Month=="01",])
summary(January.lm)

abline(coef(January.lm), col="red")

```

```{r February2}
plot(MonthlyMean$SNOW[MonthlyMean$Month=="02"], ty='l')
plot(SNOW~YEAR, data=MonthlyMean[MonthlyMean$Month=="02",], ty='l') 
February.lm <- lm(SNOW~YEAR, data=MonthlyMean[MonthlyMean$Month=="02",])
summary(February.lm)

abline(coef(February.lm), col="red")

```