---
title: "Temperature Changes at Ewa Plantation on O`ahu"
author: "Kelli Kokame"
date: "2/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


After downloading a dataset off of NOAA's climate database for the Ewa Plantation station on the island of O`ahu, we inputted it into R for processing. This dataset includes variables like maxmimum temperature, minimum temperature, and precipitation. This blog will explore trends in temperature and precipitation at this specific station.  

```{r importingCSV}
# Importing Scripts

file = "/home/CAMPUS/kak22013/Climate_Change_Narratives/Kelli/Kokame.csv"

import = read.csv(file)
```

## Let's take a look at our data. 

```{r checking}
plot(TMAX~DATE, import)
```

When we first import the data, it is not yet in a form that we can use and visualize well. As you can see, there are several values that don't make sense. How could the maximum temperatures be almost -10000 degrees? It's actually not--the value -9999 is used to represent having no data. We can change these values to "NA" to ensure that the big negative values do not skew our data. 

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```

After changing the -9999 values to NA, we can check to see if our data is now usable:

```{r}
plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```

Even though we removed the missing values, the data is still unusable. The reason why has to do with how data used to be recorded. Instead of chronologically entering data in a way that Excel could read and turn into a continous timeline, dates used to be inputted in the format of YYYYMMDD. This means that, for example, 19131231 (December 31, 1913) would then turn into 19140101 (January 1, 1914). Chronologically, there are several missing values in between which is why there are so many gaps. The data still needs to be chaned into a form in which dates are continuous and numerically consecutive. In this way, it is true scientists have to manipulate data--but it is not so much the content of the data as the format.  


```{r}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

## Fixing the Dates

To get the data to organize like we want, we first need to convert the date to string or character values. After that, we can convert the strings to a data format. 

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```
## Subset Sites

I have two sites in this data, so, we need to subset it.

```{r}
unique(import$STATION_NAME)
```

Let's choose EWA PLANTATION 741 HI US as the station name...

```{r}
Ewa <- subset(import, STATION_NAME=="EWA PLANTATION 741 HI US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
```


## A new plot
```{r}
plot(TMAX~NewDate, Ewa, ty='l')
Ewa.lm <- lm(TMAX~NewDate, data=Ewa)
summary(Ewa.lm)
plot(TMAX~NewDate, Ewa, ty='l')
abline(coef(Ewa.lm), col='red')
summary(Ewa.lm)
```
As shown, the maximum temperature annually has had a slight upward trend from 1905 when the records began until 2006. The slope is about 3.1e-05 (a very tiny slope). However, we can see that there is a lot of missing data beginning at roughly 1980. While I tried to look into why there were no available measurements for about a 20 year span, no explanations were offered online. 

But since there's a lot of missing data beginning at about 1980, let us focus on the trend from roughly 1900 up until that point to see if the missing data has an effect on the trend slope. 

Here is the graph of maximum temperatures from 1905 up to 1978: 
```{r}
plot(TMAX~NewDate, Ewa[Ewa$DATE< 19780101,], ty='l')
```


## Evaluating the trend

Now, let us analyze the trend of the maxmimum temperature data from 1905 to 1978.  
```{r}
# Linear Model
Ewa.lm <- lm(TMAX~NewDate, data=Ewa[Ewa$DATE< 19780101,])
summary(Ewa.lm)
plot(TMAX~NewDate, Ewa[Ewa$DATE< 19780101,], ty='l')
abline(coef(Ewa.lm), col='red')
```

The trend of maxmimum temperatures has a steeper slope of 6.565e-05 when excluding all of the missing data.  

Let's do the same with minimum temperature measurements. 

```{r}
# Linear Model
Ewa.lm <- lm(TMIN~NewDate, data=Ewa[Ewa$DATE< 19780101,])
summary(Ewa.lm)
plot(TMIN~NewDate, Ewa[Ewa$DATE< 19780101,], ty='l')
abline(coef(Ewa.lm), col='red')
```


As you can see, minimum temperatures also have an upward trend, with a slope of about 5.8e-05. 


## Creating Month and Year Variables

While Hawaii's temperatures are relatively constant year-round, let's still take a look at individual months and their data across different years.

```{r}
# Get months
Ewa$Month = months(Ewa$NewDate) # Creates problems.
Ewa$Month = format(as.Date(Ewa$NewDate), format = "%m")
Ewa$Year = format(Ewa$NewDate, format="%Y")
```

### Creating Monthly Means
```{r creatingmonthlymeans}
MonthlyMean = aggregate(TMAX ~ Month + Year, Ewa, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
MonthlyMean$MONTH = as.numeric(MonthlyMean$Month)


# Change mean to sd and you'll get standard deviation for each month/year.

MonthlySD = aggregate(TMAX ~ Month + Year, Ewa, sd)

MonthlySD$YEAR = as.numeric(MonthlySD$Year)
MonthlySD$MONTH = as.numeric(MonthlySD$Month)
MonthlySD$NewDate = MonthlySD$YEAR + (MonthlySD$MONTH - 1)/12

head(MonthlySD)
```


```{r plottingMonthyMeans}
plot(MonthlyMean$TMAX, ty='l')
```

below is the standard deviation. 


```{r plottingMonthySD}
plot(MonthlySD$TMAX, ty='l')

plot(TMAX~ NewDate, data=MonthlySD, ty='l')
SD.lm <- lm(TMAX~NewDate, data=MonthlySD)
summary(SD.lm)

abline(coef(SD.lm), col="red")

```

### Selecting for 1 Month -- July

Let's now look at the trend of temperatures in July, a summer month, throughout the years. 

```{r May}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="07"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="07",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="07",])
summary(May.lm)

abline(coef(May.lm), col="red")

```

Now, the change is `r coef(May.lm)[2]` degress/year or `r round(coef(May.lm)[2]*100, 3)` degress/100 years (YIKES!) with a probability of `r round(summary(May.lm)$coefficients[2,4],4)`. Although we can't reject the null hypothesis, it is worrying!


### Selecting for 1 Month -- December

Now let's try the same thing with December, a winter month. 

```{r}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="12"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",], ty='l') 
Dec.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, Ewa, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

Here's a graphic showing the trend of minimum temperatures in December over the years: 

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="12"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="12",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degress/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temerpature, that year is one, is quite problematic.

Let's look at the trends of minimum temperature throughout the years for the month of July. 

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="07"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="07",], ty='l') 
July.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="07",])
summary(July.lm)

abline(coef(July.lm), col="red")

```

The minimum temperatures in the month of July are increasing as well, with a slope of .025. This means that at the current trend, temperatures are increasing at a rate of .025 degrees fahrenheit per year, or 2.5 degrees fahrenheit per century. The understood goal is to keep climate change within 2 degrees celcius or 3.6 degrees fahrenheit above preindustrial levels, and while we can by no means generalize the trends of Ewa Station to the entire globe, this is still a concerning rate. 

## Discussion

While the data at the Ewa Plantation station is incomplete, it does undoubtedly point to a trend that we would expect to come with climate change. Both the minimum and maximum temperature trends increase over the years. Also, both months in the summer as well as in the winter see this upward climbing trend. 

Though the trend line is not very steep, any change has to be considered over long periods of time. If the temperature follow along this same trend for a much longer period of time, the temperature could then have an overall significant increase. Also, this data unfortunately does not include data from the most recent years, so it could be missing trends in temperature caused by increases in carbon dioxide levels since 2006. If the station had included readings from 2015 and onward, we could possibly see the effects of carbon dioxide in the atmosphere reaching over 400 parts per million for the first time in recorded history. 





# Departure from Mean

```{r}
Ewa$PRCP[Ewa$PRCP==-9999] <- NA
PRCP_mean = mean(Ewa$PRCP, na.rm=T)

plot(PRCP~NewDate, data=Ewa)
abline(h=PRCP_mean, col="green")

```
```{r}
plot(MonthlyMeanPRCP$PRCP, ty='l')

plot(MonthlyMeanPRCP$PRCP[MonthlyMeanPRCP$Month=="07"], ty='l')

JulyPRCP.lm <- lm(PRCP~YEAR, data=MonthlyMeanPRCP[MonthlyMeanPRCP$Month=="07",])
summary(JulyPRCP.lm)
abline(coef(JulyPRCP.lm), col="blue")
```

```{r}
MonthlyMeanPRCP = aggregate(PRCP ~ Month + Year, Ewa, mean)

MonthlyMeanPRCP$YEAR = as.numeric(MonthlyMeanPRCP$Year)
head(MonthlyMeanPRCP)
plot(MonthlyMeanPRCP$PRCP, ty='l')
```


## Experimental Portion --- Precipitation

Precipiation might depend more on the departure from the mean (often referred as as normal, whatever that means!).  I think it's worth pursuing, but haven't finished the analysis yet.

First, we need a "mean" -- The IPCC uses 1961-1990 as a norm, I don't know what is the standard for California, so we should look that up.

Second, we need to remove the missing values and evalaute which years have complete years. If you are missing rainy months, then the whole year should be thrown out -- but what about partial years in the drought season?  

Third, we will need to decide what level of aggredation -- monthly, yearly, etc. 

Fourth, in CA the water year starts in Oct 1. Should we follow the same convention?


```{r}
Ewa$PRCP[Ewa$PRCP==-9999] <- NA
YearlySum = aggregate(PRCP ~ Year, Ewa, sum)
YearlySum$YEAR = as.numeric(YearlySum$Year) 
YearlyMean = mean(YearlySum$PRCP)
```

A yearly mean, based on the annual sum for the entire records. Not sure this is appropriate.

Figure has points of the yearly sum of rainfall and the blue line mean. The greenline is the trend and red line is a five year running average, I think!  I am still trying to understand what the code is doing.
```{r}
plot(PRCP~YEAR, data=YearlySum, las=1, ty="p")
abline(h=YearlyMean, col="blue")
YearlySum.lm = lm(PRCP~YEAR, data=YearlySum)
abline(coef(YearlySum.lm), col="green")

n <- 5
k <- rep(1/n, n)
k

y_lag <- stats::filter(YearlySum$PRCP, k, sides=1)
lines(YearlySum$YEAR, y_lag, col="red")
```

The model suggests that the precipitation is declines at a rate of `r coef(YearlySum.lm)[2]` cm yr$^{-1}~$, or `r round(coef(YearlySum.lm)[2]*10, 2)` cm decade$^{-1}$.

```{r}
summary(YearlySum.lm)
```

Other attempts that didn't work!
```{r}
YearlySum2 = subset(YearlySum, select=c(PRCP))
YearlyMean = mean(YearlySum2$PRCP)

#filter(YearlySum2)



#Doesn't work...
# ma <- function(x,n=5){filter(x,rep(1/n,n), sides=2)}
# ma(YearlySum$PRCP)  


yearlyts =ts(YearlySum2) 
str(yearlyts)




```
yearlyts.lag = lag(yearlyts, 5)[6:112,]; head(yearlyts.lag)

plot(yearlyts, lty = 1, col = 3)

lines(yearlyts.lag)

YearlySum$Departure = YearlyMean - YearlySum$PRCP
YearlySum$YEAR = as.numeric(YearlySum$Year)

plot(PRCP~YEAR, data=YearlySum, las=1, ty="l")
abline(h=YearlyMean, col="blue")

lines(Departure ~ YEAR, data=YearlySum, col="red")
