---
title: "20190309_monthly_TMAX_data"
author: "bljj2015"
date: "3/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r D1 monthly PRCP, }
D1_daily_precip$Month = format(as.Date(D1_daily_precip$date), format = "%m") 
D1_daily_precip$Year = format(D1_daily_precip$date, format="%Y")
D1_MonthlyPRCPsum = aggregate(pppt_tot ~ Month + Year, D1_daily_precip, sum)
D1_MonthlyPRCPsum$YEAR = as.numeric(D1_MonthlyPRCPsum$Year)
D1_MonthlyPRCPsum$MONTH = as.numeric(D1_MonthlyPRCPsum$Month)
str(D1_MonthlyPRCPsum)
##plot
plot(pppt_tot~YEAR, data=D1_MonthlyPRCPsum[D1_MonthlyPRCPsum$Month=="09",], ty='l', xlim=c(1890, 2020)) 
D1_PRCP.lm <- lm(pppt_tot~YEAR, data=D1_MonthlyPRCPsum[D1_MonthlyPRCPsum$Month=="09",]) 
abline(coef(D1_PRCP.lm), col="red")
summary(D1_PRCP.lm)
#YEAR = 0.07644 , p-value = 0.782 , adj r2 = -0.01806
```



```{r Saddle monthly PRCP, }
Saddle_daily_precip$Month = format(as.Date(Saddle_daily_precip$date), format = "%m") 
Saddle_daily_precip$Year = format(Saddle_daily_precip$date, format="%Y")
Saddle_MonthlyPRCPsum = aggregate(ppt_tot ~ Month + Year, Saddle_daily_precip, sum)
Saddle_MonthlyPRCPsum$YEAR = as.numeric(Saddle_MonthlyPRCPsum$Year)
Saddle_MonthlyPRCPsum$MONTH = as.numeric(Saddle_MonthlyPRCPsum$Month)
str(Saddle_MonthlyPRCPsum)
##plot
plot(ppt_tot~YEAR, data=Saddle_MonthlyPRCPsum[Saddle_MonthlyPRCPsum$Month=="09",], ty='l', xlim=c(1890, 2020)) 
Saddle_PRCP.lm <- lm(ppt_tot~YEAR, data=Saddle_MonthlyPRCPsum[Saddle_MonthlyPRCPsum$Month=="09",]) 
abline(coef(Saddle_PRCP.lm), col="red")
summary(Saddle_PRCP.lm)
#YEAR = 0.5643 , p-value = 0.476 , adj r2 = -0.01317
```



```{r C1 monthly PRCP, }
C1_daily_precip$Month = format(as.Date(C1_daily_precip$date), format = "%m") 
C1_daily_precip$Year = format(C1_daily_precip$date, format="%Y")
C1_MonthlyPRCPsum = aggregate(ppt_tot ~ Month + Year, C1_daily_precip, sum)
C1_MonthlyPRCPsum$YEAR = as.numeric(C1_MonthlyPRCPsum$Year)
C1_MonthlyPRCPsum$MONTH = as.numeric(C1_MonthlyPRCPsum$Month)
str(C1_MonthlyPRCPsum)
##plot
plot(ppt_tot~YEAR, data=C1_MonthlyPRCPsum[C1_MonthlyPRCPsum$Month=="09",], ty='l', xlim=c(1890, 2020)) 
C1_PRCP.lm <- lm(ppt_tot~YEAR, data=C1_MonthlyPRCPsum[C1_MonthlyPRCPsum$Month=="09",]) 
abline(coef(C1_PRCP.lm), col="red")
summary(C1_PRCP.lm)
#YEAR = -0.07338 , p-value = 0.758 , adj r2 = -0.01433
```



```{r B14W monthly PRCP}
Boulder_14_W$Month = format(as.Date(Boulder_14_W$NewDate), format = "%m") 
Boulder_14_W$Year = format(Boulder_14_W$NewDate, format="%Y")
B14W_MonthlyPRCPsum = aggregate(PRCP ~ Month + Year, Boulder_14_W, sum)
B14W_MonthlyPRCPsum$YEAR = as.numeric(B14W_MonthlyPRCPsum$Year)
B14W_MonthlyPRCPsum$MONTH = as.numeric(B14W_MonthlyPRCPsum$Month)
str(B14W_MonthlyPRCPsum)
##plot
plot(PRCP~YEAR, data=B14W_MonthlyPRCPsum[B14W_MonthlyPRCPsum$Month=="09",], ty='l', xlim=c(1990, 2020)) 
B14W_PRCP.lm <- lm(PRCP~YEAR, data=B14W_MonthlyPRCPsum[B14W_MonthlyPRCPsum$Month=="09",]) 
abline(coef(B14W_PRCP.lm), col="red")
summary(B14W_PRCP.lm)
# YEAR | ADJ R2 | P-VALUE | STAT SIG?
# 1.155 | -0.06123 | 0.719 | no
```



```{r CCC monthly PRCP}
Coal_Creek_Canyon$Month = format(as.Date(Coal_Creek_Canyon$NewDate), format = "%m") 
Coal_Creek_Canyon$Year = format(Coal_Creek_Canyon$NewDate, format="%Y")
CCC_MonthlyPRCPsum = aggregate(PRCP ~ Month + Year, Coal_Creek_Canyon, sum)
CCC_MonthlyPRCPsum$YEAR = as.numeric(CCC_MonthlyPRCPsum$Year)
CCC_MonthlyPRCPsum$MONTH = as.numeric(CCC_MonthlyPRCPsum$Month)
str(CCC_MonthlyPRCPsum)
##plot
plot(PRCP~YEAR, data=CCC_MonthlyPRCPsum[CCC_MonthlyPRCPsum$Month=="09",], ty='l', xlim=c(1890, 2020)) 
CCC_PRCP.lm <- lm(PRCP~YEAR, data=CCC_MonthlyPRCPsum[CCC_MonthlyPRCPsum$Month=="09",]) 
abline(coef(CCC_PRCP.lm), col="red")
summary(CCC_PRCP.lm)
##YEAR = 0.3374, p-value = 0.637, adj r2 = -0.02327
```




```{r Boulder monthly PRCP}
Boulder$Month = format(as.Date(Boulder$NewDate), format = "%m") 
Boulder$Year = format(Boulder$NewDate, format="%Y")
Boulder_MonthlyPRCPsum = aggregate(PRCP ~ Month + Year, Boulder, sum)
Boulder_MonthlyPRCPsum$YEAR = as.numeric(Boulder_MonthlyPRCPsum$Year)
Boulder_MonthlyPRCPsum$MONTH = as.numeric(Boulder_MonthlyPRCPsum$Month)
str(Boulder_MonthlyPRCPsum)
##plot
plot(PRCP~YEAR, data=Boulder_MonthlyPRCPsum[Boulder_MonthlyPRCPsum$Month=="09",], ty='l', xlim=c(1890, 2020)) 
Boulder_PRCP.lm <- lm(PRCP~YEAR, data=Boulder_MonthlyPRCPsum[Boulder_MonthlyPRCPsum$Month=="09",]) 
abline(coef(Boulder_PRCP.lm), col="red")
summary(Boulder_PRCP.lm)
##YEAR = 0.2008, p-value = 0.123, adjusted R2=0.01176
```






##old table 3 of daily prcp, replaced w/ monthly prcp sum
_**Table 3:** Linear regression model of daily precipitation_

Climate Station | change in daily PRCP | Adjusted R-squared | p-value | Statistical significance?
--------------- | ----------- | ------------- | ------------ | ---------------
1. D-1               | +2.24e-05 | 0.000262  | 0.0164 | yes
2. Saddle            | +2.57e-04 | 0.00725  | <2e-16 | yes
3. C-1               | -8.74e-05 | 0.00954   | <2e-16 | yes
4. Boulder 14 W      | +8.30e-05 | 0.000492 | 0.0531 | no
5. Coal Creek Canyon | +5.15e-06 | -6.66e-05 | 0.688 | no
6. Boulder           | +1.62e-06 | -5.28e-06 | 0.379 | no


## OLD D1 DAILY TMAX GRAPH I USED BEFORE MARC SAID IT WASN'T INFORMATIVE SINCE THE GRAPH DOESN'T TELL MUCH

```{r D1 daily TMAX & PRCP graphs, echo=FALSE}
plot(airtemp_max~date, D1_daily_temp, main="Figure 5 DELETEME. Daily Max Temp at D-1 Station", xlab = "Date (Year)", ylab = "Daily Max Temperature (°C)", ty='l', las=1)
abline(lm(airtemp_max ~ date, data=D1_daily_temp), col="red")

```

##**OLD**##

```{r Boulder monthly TMAX all months}
par(mfrow = c(4, 3), mar = c(1, 2, 2, 1) + 0.1) 
TMAXresult <- NA 
for (i in 8:10) { 
  # plot(Boulder_MonthlyTMAXMean£TMAX[Boulder_MonthlyTMAXMean£Month==i], 
  # ty='l') 
  plot(TMAX ~ YEAR, data = Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), main = Months[i])
  Boulder_Month.lm <- lm(TMAX ~ YEAR, data = Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$MONTH == i, ]) 
  summary(Boulder_Month.lm)
  abline(coef(Boulder_Month.lm), col = "red")
  Boulder_TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Boulder_Month.lm)[2], 4), round(summary(Boulder_Month.lm)$coefficients[2, 4], 4), round(summary(Boulder_Month.lm)$r.squared, 3)))
}
Boulder_Aug_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="08",])
summary(Boulder_Aug_TMAX.lm)
Boulder_Sep_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="09",])
summary(Boulder_Sep_TMAX.lm)
Boulder_Oct_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="10",])
summary(Boulder_Oct_TMAX.lm)
```

##Boulder_14_W

```{r}
Boulder$Month = format(as.Date(Boulder$NewDate), format = "%m")
Boulder$Year = format(as.Date(Boulder$NewDate), format="%Y")
```

```{r}
Boulder_MonthlyTMAXMean = aggregate(TMAX ~ Month + Year, Boulder, mean)
Boulder_MonthlyTMAXMean$YEAR = as.numeric(Boulder_MonthlyTMAXMean$Year)
Boulder_MonthlyTMAXMean$MONTH = as.numeric(Boulder_MonthlyTMAXMean$Month)
str(Boulder_MonthlyTMAXMean)
plot(Boulder_MonthlyTMAXMean$TMAX, ty='l')
```

```{r Boulder monthly TMAX all months}
par(mfrow = c(4, 3), mar = c(1, 2, 2, 1) + 0.1) 
TMAXresult <- NA 
for (i in 8:10) { 
  # plot(Boulder_MonthlyTMAXMean£TMAX[Boulder_MonthlyTMAXMean£Month==i], 
  # ty='l') 
  plot(TMAX ~ YEAR, data = Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$MONTH == i, ], ty = "l", las = 1, xlim = c(1940, 2020), main = Months[i])
  Boulder_Month.lm <- lm(TMAX ~ YEAR, data = Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$MONTH == i, ]) 
  summary(Boulder_Month.lm)
  abline(coef(Boulder_Month.lm), col = "red")
  Boulder_TMAXresult <- rbind(TMAXresult, cbind(Months[i], round(coef(Boulder_Month.lm)[2], 4), round(summary(Boulder_Month.lm)$coefficients[2, 4], 4), round(summary(Boulder_Month.lm)$r.squared, 3)))
}
Boulder_Aug_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="08",])
summary(Boulder_Aug_TMAX.lm)
Boulder_Sep_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="09",])
summary(Boulder_Sep_TMAX.lm)
Boulder_Oct_TMAX.lm <- lm(TMAX~YEAR, data=Boulder_MonthlyTMAXMean[Boulder_MonthlyTMAXMean$Month=="10",])
summary(Boulder_Oct_TMAX.lm)
```

```{r}
Boulder_MonthlyPRCPMean = aggregate(PRCP ~ Month + Year, Boulder, sum)
Boulder_MonthlyTMAXMean$YEAR = as.numeric(Boulder_MonthlyTMAXMean$Year)
Boulder_MonthlyTMAXMean$MONTH = as.numeric(Boulder_MonthlyTMAXMean$Month)
str(Boulder_MonthlyTMAXMean)
plot(Boulder_MonthlyTMAXMean$TMAX, ty='l')
```