---
title: "Where is California focusing treatments to prevent wildfire?"
author: "Sabrina Wilk"
date: "12/15/2018"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 

Since the 1990s, fire management agencies have increasingly used hazardous fuel treatments, such as prescribed burns and mechanical vegetation removal, to combat wildfire. Research in the past twenty years has sought to improve effectiveness of these treatments, answering questions such as: How is hazardous fuel best treated? Where should efforts be concentrated? This study examines how the presence of neighborhoods and previous fire sites affects where California’s firefighting organizations focus hazardous fuel treatments.

## What are hazardous fuels?

Hazardous fuels are defined as “live and dead vegetation that has accumulated and increases the likelihood of unusually large wildland fires” (CalFire). This includes dead trees, or ‘snags,’ as well as dense underbrush and clumps of trees that are particularly close together. Fire management agencies classify sites of dense vegetation as problem areas because the vegetation acts as tinder when a fire starts, with flames racing through the underbrush and climbing up the small, crowded trees. Firefighters refer to abundance of vegetation as ‘fuel load,’ with areas of low fuel load requiring little management and areas of high fuel load often requiring extensive removal.

As costs of wildfire suppression quadrupled from 1985 to 2014, pressure increased on fuel treatment programs as a way to prevent large, dangerous, and costly wildfires from starting—or at least, from spreading so rapidly (Vaillant and Reinhardt 2017). In 2001, fuel treatment became one of four independent categories in the national fire budget, placing it alongside categories targeting suppression, prevention, and miscellaneous (Hoover 2017). This budgetary shift marked the rise in importance of fuel reduction programs within the national wildfire discussion.

## How did hazardous fuels become an issue?

For most of the twentieth century, firefighting organizations suppressed fires whenever possible. Deadly fires in Michigan in the 1880s and 1890s, as well as fires in Maine, New York, and southwestern Washington state, burned millions of acres, killed over 500 people, and sparked an anti-wildfire response across the United States. The Forest Service, founded in 1905, was besieged by public demands for protection from these types of blazes. In response, the FS established standardized wildfire suppression practices (Williams 2005).

However, lack of fires had unexpected ramifications, particularly in ecosystems that had historically experienced frequent, low-intensity fires such as those in California (Ryan et al 2013). Decades of fire suppression led to higher tree density, reduced tree size, and increased amounts of understory vegetation. The FS, along with other fire management agencies that had followed its lead, had unintentionally created a hotbed for larger, more intense wildfires. In California in the late 1900s and early 2000s, this resulted in an increased number of ‘crown fires’—fires that consume entire patches of forest rather than burning at low intensity throughout the understory (Ryan et al 2013; Agee and Skinner 2005). Not only did crown fires pose a larger threat to nearby towns and cities; they also affected the health of ecosystems by burning them to the extent that they could not regenerate (Keane et al 2008; Keane and Karau 2010).

By 2001, management agencies realized that fuel buildup was driving larger fires and had shifted from their previous ‘absolute suppression’ mindset to a heightened emphasis on fuel reduction (Williams 2005). Firefighting organizations began to allow natural, low-intensity fires to burn and even initiated prescribed burns in fire-prone or high fuel-load areas (Ryan et al 2013). In some areas, agencies also undertook laborious mechanical treatments, thinning crowded trees and removing underbrush. Reducing fuel promoted regular flora generation while also keeping fires at low-intensity, preserving ecosystem health and citizen safety by preventing larger, more destructive fires (Ryan et al 2013). Many of the same practices continue today.

##My question: Where do California’s fire management agencies focus their fuel treatments?

Fuel reduction programs are limited by funding: despite their rise in importance, they require time and ample amounts of labor, which makes them costly. As a result, firefighting organizations, in California preliminarily CalFire, must prioritize certain areas over others. I hypothesized that areas where residential neighborhoods meet undeveloped land, or ‘Wildland Urban Interfaces’ (WUIs), are areas of high priority because of human proximity. Therefore, I expected acreage of treatment near WUIs to be higher than acreage farther from WUIs. I also hypothesized that sites of previous fires affect where CalFire carries out treatment, and that acreage of treatment near previous fire sites is greater than acreage farther from previous fire sites.

## Results

Controlled burns accounted for a significantly greater amount of total treatment area than did mechanical removals (98.5% and 1.5%, respectively), while areas that were both burned and treated mechanically took up .1% of total treatment area. None of the treatments (controlled burn, mechanical, or both) accounted for significantly greater area on or near previous fire sites than not (Table 1). Similarly, neither controlled burns nor mechanical treatments accounted for significantly greater area near WUIs than not (Table 2). The only significant relationship was between overlap of treatments and WUIs; a significantly greater total area of treatment overlap occurred near WUIs than not (Figure 2).

View this link for images of the mapping process:
https://docs.google.com/document/d/1uRrmEc1T8CJPTqHSPjwRxxq7opztaVexhrCqaru-Th4/edit?usp=sharing

Table 1. Area of land (in hundreds of m^2) treated in relation to previous fire sites.

```{r cars, echo=FALSE, eval=FALSE}
library(readr)
table_prevfires <- 
  read.csv("table_prevfires.csv", header = TRUE)
library(knitr)
kable(table_prevfires[0:4,], col.names = c("Treatment type", "Intersects", " ", "Does not intersect", " ", "P-value"))

```

 Table 2. Area of land (in hundreds of m^2) treated in relation to WUIs.
 
```{r, echo=FALSE, eval=FALSE}
library(readr)
table_WUIs <- read.csv("table_WUIs.csv")
library(knitr)
kable(table_WUIs[0:4,], col.names = c("Treatment type", "Intersects", " ", "Does not intersect", " ", "P-value"))

```
```{r, echo=TRUE, eval=FALSE}
df <- read.csv("mechburn_WUIs.csv", header=TRUE)
df$Not = factor(df$Not,labels = c("Intersects", "Does not intersect"))
boxplot(overlaparea ~ Not, data = df, ylab = "Area (in hundreds of m^2)")
```

Figure 1. Mean total area of mechanical and burn treatment overlap (in hundreds of m^2) that intersects and does not intersect with WUIs.

## Discussion

Controlled burns accounted for far more land than did mechanical treatments. This may be because mechanical treatments are more expensive and require more labor to carry out in widespread areas. In addition, controlled burns became common practice before mechanical treatments and so were standardized prior to most mechanical methods (Williams 2005).

Contrary to my hypothesis, mean total area of treatment did not correlate with location of previous fire sites. I had predicted that treatment area would be greater near fire sites on the grounds that previous fires likely marked fire-prone areas, and that CalFire and other state agencies would therefore prioritize those areas. While mechanical treatments occurred in nearly equal amounts irregardless of fire sites, an insignificantly greater area of controlled burns occurred near previous fire sites than not (covering 1.7 times as much area on or near previous fire sites). This may reflect a slight prioritization of previous fire sites over other areas for controlled burns. With the tremendous fires occurring in northern California in recent fire seasons—including the Carr Fire and Mendocino Complex Fire, among others—CalFire may increase its attention to these sites and other areas where fires have occurred. It is difficult to map out the effects that wind, landscape, and vegetation will have on fires, but paying attention to historical blazes can inform fire agencies of potential fire prone areas.

In fact, locations of fire-prone areas have largely been disregarded through California’s history, and that reflects in the number of homes built in fire-prone areas (Thompson 2018). Many of these are in WUIs (CalFire). I hypothesized that presence of WUIs influenced how state fire agencies selected treatment locations and predicted that area of treatments near WUIs would be significantly greater than area of treatments farther from WUIs. Areas of treatment overlap—in other words, where mechanical and burn treatments both occurred—confirmed my hypothesis. Over 400 times as much area where treatments overlapped occurred close to WUIs; however, treatment overlap comprised a very small amount of total treated state land (.1%).

The lack of significant difference in overall mechanical treatment area nearer to vs farther from WUIs was surprising, given that CalFire has programs set in place specifically to remove fuel near WUIs. For instance, the CalFire website describes fuel management zones as “areas, usually surrounding communities, where the natural vegetative cover is reduced in density…fuel ladders are greatly reduced, and overstory and understory vegetation is spatially separated so that a ground fire will not, under normal fire conditions, climb into the canopy and turn into a crown fire” (CalFire). However, agencies must also undertake mechanical treatments in non-developed areas with high risk of fire ignition, and this is likely part of the reason that mechanically-treated area was not significantly greater near WUIs. This division in priorities likely also plays a role in the lack of significant difference in area of controlled burns near and far from WUIs. While people’s safety is a priority, other variables such as an area’s risk of ignition, fire regime, vegetation, and topography play a role in location of treatments as well.

Ultimately, multiple factors need to be addressed in order to mitigate fire danger. These include better planning for housing developments, with fire-prone areas in mind, as well as using pro-active suppression strategies such as pre-positioning by fire agencies (Thompson 2018, Gossner 2018). Fuel treatments are an important part of this discussion, both in California and across the U.S. They may be able to combat rising fire sizes and consequently lower high suppression costs. Studies should continue to address how to strategize fuel treatment locations as well as develop the most effective methodologies to remove fuel in different vegetation types.

## Methods

I accessed GIS layers of controlled burns and mechanical treatments in California from ca.gov, and obtained geographical data on state, federal, and local Direct Protection Areas from the Forest Service website. I intersected the layers of controlled burns and mechanical treatments, respectively, with state DPAs to create layers of each type of treatment on state lands. I named the new layers ‘State Burns’ and ‘State Mechanical.’ To find areas where controlled burns and mechanical treatments had both been performed, I intersected ‘State Burns’ and ‘State Mechanical,’ and named the new layer ‘State Burns Mechanical Overlap.’ These layers allowed me to find the total area (in m^2) of controlled burns, mechanical treatments, and treatment overlap on state lands.

I also accessed GIS data on WUIs and locations of fires prior to 2018 on ca.gov and the FS website. I created a 400m buffer around the WUI layer and selected for burns and mechanical treatments with any part of their boundaries intersecting the buffer. This allowed me to create layers of controlled burns and mechanical treatments on state lands that were near WUIs. I used a 400m buffer to more accurately include controlled burns: while it is unlikely that agencies would permit controlled burns to reach the residential areas of WUIs, part of a burn may extend within 400m of a WUI. Intersecting layers of WUIs and treatments may, however, result in a generous estimate of fuel treatment area near WUIs because only a small section of the treatment needed to intersect the WUI buffer to be included in the calculations, and some of the treatments were larger than 40,000 km^2.

I repeated this process with the previous fires data to create a new layer of controlled burns and mechanical treatments intersecting sites of previous fires. This again may result in a generous estimate of treatment area close to previous fire sites, because only a small section of the treatment needed to intersect the previous fires layer to be included in the calculations.

I then decided to sort total treatment area for each of the three treatments by county. I accessed a GIS layer of counties in California from ca.gov. To calculate total area of state treatments (controlled burn, mechanical, and burn/mechanical overlap) in each county, I used the ArcMap ‘split’ function to divide treatment layers into counties. I then needed to find how much area per treatment was near WUIs or near previous fire sites. The split function did not work with the layers of treatment intersection with WUIs and previous fire sites. Instead, I created a new layer from each county, then used the ‘select by location’ function to select areas of WUI overlap and previous fire overlap within each county. I did this for all three treatment types. This allowed me to calculate total area of each treatment, within each county, that intersected with WUIs or with previous fire sites. I recorded areas in an Excel spreadsheet.

I calculated mean area of treatment overlap with WUIs in counties then did the same for previous fires. I also calculated standard deviation. I used t-tests to analyze for statistical difference between treatment area that intersected vs did not intersect WUIs and previous fire sites, respectively, for each of the three treatments.

## References

Agee, J.K., and C.N. Skinner. 2005. Basic principles of forest fuel reduction treatments. Forest Ecology and Management 211 (1-2): 83-96.

“CA County Boundaries.” GIS Layer.  Sourced from “California Open Data Portal: CA Geographic Boundaries.” https://data.ca.gov/dataset/ca-geographic-boundaries. Accessed December 20, 2018.

“Direct Protection Areas for Wildland Fire Protection.” GIS Layer. Sourced from “State Level Datasets.” USDA Forest Service, Pacific Southwest Region. https://www.fs.usda.gov/detail/r5/landmanagement/gis/?cid=STELPRDB5327836. Accessed October 29, 2018.

“Fire Perimeters Version 17_1.” GIS Layer. Sourced from “FRAP Mapping.” CalFire. http://frap.fire.ca.gov/data/frapgisdata-sw-fireperimeters_download. Accessed October 29, 2018.

“Fuels Treatment.” 2012. Cal Fire. calfire.ca.gov.

Gossner, A. 2018. “Fund California fire departments to deploy engines before fires break out.” San Francisco Chronicle.

Hoover, K. 2017. Wildfire Management Funding: Background, Issues, and FY2018 Appropriations. Congressional Research Service.

Keane, R.E., J.K. Agee, P. Fulé, J.E. Keeley, C. Key, S.G. Kitchen, R. Miller, and L.A. Schulte. 2008. Ecological effects of large fires on US landscapes: benefit or catastrophe? International Journal of Wildland Fire 17(6): 696-712.

Keane, R.E., and E. Karau. 2010. Evaluating the ecological benefits of wildfire by integrating fire and ecosystem simulation models. Ecological Modelling 221: 1162-1172.

Ryan, K.C., E.E. Knapp, and J.M. Varner. Prescribed fire in North American forests and woodlands: history, current practice, and challenges. Frontiers in Ecology and the Environment 11(1): e15-e24.

Thompson, D. 2018. “Cal Fire chief: State must consider ban on homes in fire-prone areas.” The Mercury News.

Vaillant, N.M, and E.D. Reinhardt. 2017. An Evaluation of the Forest Service Hazardous Fuels Treatment Program—Are We Treating Enough to Promote Resiliency or Reduce Hazard?” Journal of Forestry 115(4): 300-308.

“Wildland Urban Intermix.” GIS Layer. Sourced from “State Level Datasets.” USDA Forest Service, Pacific Southwest Region. https://www.fs.usda.gov/detail/r5/landmanagement/gis/?cid=STELPRDB5327836. Accessed November 2, 2018.

Williams, G.W. 2005. The USDA Forest Service: The First Century. United States Department of Agriculture.






