---
title: "Attribution Science -- Do Extreme Events have a Climate Change Signature"
subtitle: "Is there a Signal in the Noise?"
author: "Environmental Analysis 30 Students, Pomona College"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```
# What is Attribution Science

Attribution Science is a subset of climate science that attempts to evaluate the effect of climate change on weather patterns, in particular extreme weather events. Determining the climate change signature on these extreme events relies a wide range of observations and data sources, such as weather instruments, records, and modeling, as well as, historic climate events and proxy data of weather.  

Unfortunately, climate science has become so politicized reports of extreme events, which can even lead to misunderstandings of these climate events. For example, [Vance 2019](S2019_Blogs/Vance/Vance2.html) documents how social media has been used to make claims about extreme events that obscure important long-term trends. In particular, in the (US) South, climate change skepticism runs, as well as, the use of coal for electrical generation. However, this region may be may be hardest hit by the effects of climate change unless we curb CO<sub>2</sub> global emissions. 

## Operationalizing Questions in Attribution Science

Asking how climate change influenced a certain extreme event is one of the first stages of attribution science. For example, after extensive flooding in Colorad in 2013, researchers formulated a testable hypothesis to evaluate the role of climate change on the flooding, but the results were contested. [Lai (2019)](S2019_Blogs/Lai/Lai2.html) used the 2013 Colorado flood as a case study to discuss how weather records, weather modeling, and scientific research to understand the causes of the flood.



Science progresses in incremental ways and often relies on proposing hypotheses and testing them. Formulating a reasonable hypothesis that can answer a valid or important question in one of the first stages. However, if the hypothesis doesn't capture the appropriate process than me might  


## Scientific 'Progress': False Positives and Negatives

Howevever, we want to distinquish process from the ideology of progress -- that the human condition has been improving in a linear fashion since recorded history. There is a lot of evidence that not everyone is better off and even if some are better off, it has not been a linear process. 

Climate change is perhaps a good example, where those that enjoy the rewards of burning fossil fuels are not subject to the proportional burden of the impacts. 

And we see some of these differential impacts during and after extreme weather events. But how do we frame these events within climate science?  Are these indicators of climate change (true positives), indicators that the climate is not changing (true negatives)?  Or even more insidious, suggesting that climate change is happening but is not (false postives) or that pointing that climate change is not happening, but it really is (false negatives). 

## Epistomologies and Climate Science

Scientists are very conservative; in other words, they are loath to make a mistake that there may be signal when there isn't one -- false positives, thus they have set up statistical criteria to reduce the changes of this error (called a Type I Error). The criteria is called alpha and is usually set at 0.05 or 1:20 change. Thus, at worse, scientists could make a mistake 5% of the time. 

# Climate Science and Local Knowledges

## Climate, Radiative Forcing, and Carbon Dioxide

Although the Earth's climate has been changing for hundreds of thousands, even millions and billions of years, even small changes to the Earth's average temperatures has significant implications for human activities. 

Climate change science can be extremely complex, but according to the Intergovernmental Panel on Climate Change (IPCC), it is ''extremely likely'' that human influence was the dominant cause of global warming between 1951 and 2010.

### Signal and Noise

Some of the main human activities that contribute to global warming are:

-- Increasing atmospheric concentrations of greenhouse gases 

```{r maunaloa, echo=FALSE}
address <- "ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt"
#download.file(address, "maunaloa", quiet = F, mode = "w", cacheOK = T)

maunaloa <- read.table("maunaloa", skip=70)
names(maunaloa) <- c("year", "month", "decimal.date", "average", "interpolated", "trend", "days")
maunaloa$average[maunaloa$average==-99.99] <- NA
maunaloa <- data.frame(year=maunaloa$year, month=maunaloa$month, decimal.date=maunaloa$decimal.date, average=maunaloa$average)
#head(maunaloa)
#str(maunaloa)
#write.csv(maunaloa, file = "H:\\My Documents\\My Webs\\Mauna_Loa.csv", eol = "\n", row.names = F)
```

```{r fig-margin, fig.margin = TRUE, fig.cap = "Observed CO~2~ concentrations (black) have been steadily increasing since we have begun measuring them in the 1950s, with a best fit line (red). Note that the slope or the rate of change is also increasing.", fig.width=3.5, fig.height=3.5, cache=TRUE, echo=FALSE}
par(las=1)
plot(maunaloa$decimal.date, maunaloa$average, type="l", ylab=expression(CO[2]*~ "(ppm)"), xlab="Year", 
main="Carbon Dioxide Concentration \n Mauna Loa, HI" )
abline(coef(lm(average~year, data=maunaloa)), col="red", lwd=2)
```

-- Global changes to land surface, such as deforestation

FIGURE NEEDED

-- Increasing atmospheric concentrations of aerosols.

FIGURE NEEDED

-- In addition to human activities, some natural mechanisms can also cause climate change, including for example, climate oscillations, changes in solar activity, and volcanic activity.

FIGURE NEEDED

Multiple lines of evidence support attribution of recent climate change to human activities:

* A basic physical understanding of the climate system: greenhouse gas concentrations have increased and their warming properties are well-established.

* Historical estimates of past climate changes suggest that the recent changes in global surface temperature are unusual.

* Computer-based climate models are unable to replicate the observed warming unless human greenhouse gas emissions are included.


Natural forces alone (such as solar and volcanic activity) cannot explain the observed warming.
The IPCC's attribution of recent global warming to human activities is a view shared by the scientific community, and is also supported by 196 other scientific organizations worldwide.

Here are two well undertood proceses: 



2) Human activity such as the burning of fossil fuels (coal, oil, gas) is rapidly increasing the amount of carbon dioxide in Earth’s atmosphere. Note: Not only are carbon dioxide emissions increasing, the rate is accelerating. 

Based on these two processes, we can draw the following conclusion:

We should expect the rising carbon dioxide concentration to warm our planet, with the warming becoming more severe as we add more carbon dioxide. But does this mean that we should also expect wildfires to increase as well?

## Local Knowledges as Embodied Evidence

Despite overwhelming support and proof of climate change,  

# What are Extreme Weather Events?



## Attribution and Environmental Justice

[Lane (2019)](S2019_Blogs/Lane/Lane.html)

# What are the Regional Impacts of Climate Change?

Our team reviewed a diverse collection of scientific literature to evaluate the implications of climate change and 

### Tourism and Jobs

[Budd (2019)](S2019_Blogs/Budd/Budd.html) describes the subtle phenology changes that has been changing with climate change...

### The Pacific Northwest

The impact of cliamte change in the Pacific Northwest and in the state of Washington has become increasingly poignant -- in part because of the issues of fire and smoke. [Generous (2019)](S2019_Blogs/Generous/Generous4.html) evaluated weather data in Spokane while [Meyer 2019](S2019_Blogs/Meyer/Meyer2.html) summarized what scientists know about climate change in Seattle. 


## Public Safety and Health

For the city of New York, [Kaufman (2019)](S2019_Blogs/Kaufman/Kaufman.html) highlighted the...

## Heat Waves and Snow Storms and Mortality

[Clark (2019)](S2019_Blogs/Clark/Clark4.html) explored extreme events in Washington D.C. and their relationship to climate change. Similarly, [Randle (2019)](S2019_Blogs/Randle/Randle.html)
evaluated the impact of extreme cold in Chicago it's relationship the to the polar vortex. 

## Floods, Droughts -- Weather and Hydrology



[Namachivayam 2019](S2019_Blogs/Namachivayam/Namachivayam.html)evaluates rising temperatures in Albuquerque and their potential impact on drought in New Mexico.  

[Wong (2019)](S2019_Blogs/Wong/Wong.html) describes how flooding in the Pearl River and the complexities of weather patterns in East Asia. 



[Yi (2019)](S2019_Blogs/Yi/Yi2.html)

[Bullock Floyd (2019)](S2019_Blogs/Bullock_Floyd/Bullock_Floyd2.html)




## Wildfire Risks and WUIs

In Southern California, the climate is not simply warming or cooling. Thanks to ongoing changes in climate, the rainy season has begun a pendulum effect, swinging from storms and flooding to drought conditions almost yearly in a phenomenon known as 'climate whiplash.' The environmental impacts of this phenomenon are severe, including increased numbers of fires, floods, and droughts plaguing the region. Evidence for climate whiplash, as well as its effects and implications [(Anderson 2019)](S2019_Blogs/Anderson/Anderson3.html).


[Kuhn (2019)](S2019_Blogs/Kuhn/Kuhn.html)

# Previous EA30 Projects

[Climate Change and Fires -- Is this a red herring or the alarm alarmists have been warning about?](Index_2018F.html)

[Why Climate Change Will Affect You Personally?](Index_2017F.html)

[Climate Change and Logical Fallacies](Index_2017S.html)
