---
title: "Attribution Science -- Do Extreme Events have a Climate Change Signature"
subtitle: "Is there a Signal in the Noise?"
author: "Environmental Analysis 30 Students, Pomona College"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```
# What is Attribution Science

## Signal and Noise

## Scientific 'Progress': False Positives and Negatives

## Epistomologies and Climate Science

# Climate Science and Local Knowledges

## Climate, Radiative Forcing, and Carbon Dioxide

Although the Earth's climate has been changing for hundreds of thousands, even millions and billions of years, even small changes to the Earth's average temperatures has significant implications for human activities. Climate change science can be extremely complex, but here are two well undertood proceses: 

1) Carbon dioxide (and other ``greenhouse gases'') trap heat and make Earth warmer than it would be otherwise.

```{r maunaloa, echo=FALSE}
address <- "ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt"
#download.file(address, "maunaloa", quiet = F, mode = "w", cacheOK = T)

maunaloa <- read.table("maunaloa", skip=70)
names(maunaloa) <- c("year", "month", "decimal.date", "average", "interpolated", "trend", "days")
maunaloa$average[maunaloa$average==-99.99] <- NA
maunaloa <- data.frame(year=maunaloa$year, month=maunaloa$month, decimal.date=maunaloa$decimal.date, average=maunaloa$average)
#head(maunaloa)
#str(maunaloa)
#write.csv(maunaloa, file = "H:\\My Documents\\My Webs\\Mauna_Loa.csv", eol = "\n", row.names = F)
```

2) Human activity such as the burning of fossil fuels (coal, oil, gas) is rapidly increasing the amount of carbon dioxide in Earth’s atmosphere. Note: Not only are carbon dioxide emissions increasing, the rate is accelerating. 

```{r fig-margin, fig.margin = TRUE, fig.cap = "Observed CO~2~ concentrations (black) have been steadily increasing since we have begun measuring them in the 1950s, with a best fit line (red). Note that the slope or the rate of change is also increasing.", fig.width=3.5, fig.height=3.5, cache=TRUE, echo=FALSE}
par(las=1)
plot(maunaloa$decimal.date, maunaloa$average, type="l", ylab=expression(CO[2]*~ "(ppm)"), xlab="Year", 
main="Carbon Dioxide Concentration \n Mauna Loa, HI" )
abline(coef(lm(average~year, data=maunaloa)), col="red", lwd=2)
```

Based on these two processes, we can draw the following conclusion:

We should expect the rising carbon dioxide concentration to warm our planet, with the warming becoming more severe as we add more carbon dioxide. But does this mean that we should also expect wildfires to increase as well?

## Local Knowledges as Embodied Evidence



# What are Extreme Weather Events?



## Attribution and Environmental Justice



# What are the Regional Impacts of Climate Change?

Our team reviewed a diverse collection of scientific literature to evaluate the implications of climate change and 

[Anderson](S2019_Blogs/Anderson.html)

## Public Safety and Health



## Heat Waves and Mortality

## Floods, Droughts -- Weather and Hydrology



## Wildfire Risks and WUIs



## Previous Projects

[Climate Change and Fires -- Is this a red herring or the alarm alarmists have been warning about?](Index_2018F.html)

[Why Climate Change Will Affect You Personally?](Index_2017F.html)

[Climate Change and Logical Fallacies](Index_2017S.html)
