---
title: "Importing and Processing CSV Files"
author: "Marc Los Huertos"
date: "12/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing CSV Files

First, we will import the CSV file. In most cases, we don't need to preprocess the csv files. 
```{r importingCSV}
#Importing Scripts

file = "/home/CAMPUS/mwl04747/github/Climate_Change_Narratives/Nicole/Marshfield climate.csv"
import = read.csv(file)
```

## Checking the Data

```{r checking}
plot(TMAX~DATE, import)
```

What's going on?  What is the deal with the -9999? These are used for missing data. We need to remove them!

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
```

Okay, now we'll check again, for dates less then Dec. 31, 1913:
```{r}
plot(TMAX~DATE, import[import$DATE<19132131,], ty='l')
```

Yikes! What's wrong? As it turns it the problem is that with how the dates are specified. In particular, the Dec 31 to Jan 1 transition: 

191231 -> 190102, if you use these as numbers it should be 121232, 12121233, etc. You see we are missing lots numbers!


```{r}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

## Fixing the Dates

First, we convert the date a string or character values. Next, we'll convert the strings to a data format. 

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```

## A new plot
```{r}
plot(TMAX~NewDate, import, ty='l')
```

## Evaluating the trend

Creating a linear model and then evaluate the rate (slope). 
```{r}
# Linear Model
import.lm <- lm(TMAX~NewDate, data=import)
summary(import.lm)
plot(TMAX~NewDate, import, ty='l')
abline(coef(import.lm), col='red')
```

## Time Series

Time series data require specific statistical methods. I started thiking about how to do this in R, but didn't finish... so this is superfolous for now.
```{r}

#2.906*10-5 degrees/year or 2.9*10-2 per year.

# Moving Average!
# average of current sample and 19 previous samples (red)
f20 <- rep(1/600, 600)
y_lag <- filter(import$TMAX, f20, sides=1)
plot(y_lag)

# Create a time series
Tmax = ts(import$TMAX)
```

## Creating Months and Year

So, let's figure out how to see how changes happen for individual months.
```{r}
# Get months
import$Month = months(import$NewDate)
import$Year = format(import$NewDate, format="%Y")

MonthlyMean = aggregate(TMAX ~ Month + Year, import, mean)
MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
head(MonthlyMean)
plot(MonthlyMean$TMAX, ty='l')

plot(MonthlyMean$TMAX[MonthlyMean$Month=="May"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="May",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="May",])
summary(May.lm)

abline(coef(May.lm), col="red")

```

Now, the change is 0.0078 (0.008/year), or 0.8/100 years.

## Next Steps

I think you should evaluate every month and see what happens. You might also consider looking at the TMIN as well. Could be important!


