---
title: "Importing and Processing CSV Files"
author: "Marc Los Huertos"
date: "12/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing CSV Files

First, we will import the CSV file. In most cases, we don't need to preprocess the csv files. 
```{r importingCSV}
<<<<<<< HEAD:David/CopyOfImporting Franks CSV.Rmd
# Importing Scripts

file = "/home/CAMPUS/fcl02013/Climate_Change_Narratives/David Wagner/Hagerstown Maryland NOAA data.csv"
=======
#Importing Scripts
>>>>>>> 87366cfb51d599c2a8821456a32daf84199e6282:Nicole/Importing CSV marshfield.Rmd

file = "/home/CAMPUS/nllk2015/Climate_Change_Narratives/Nicole/Marshfield climate.csv"
import = read.csv(file)
```

## Checking the Data

```{r checking}
plot(TMAX~DATE, import)
```

What's going on?  What is the deal with the -9999? These are used for missing data. We need to remove them!

```{r missings}
import$TMAX[import$TMAX==-9999] = NA
import$TMIN[import$TMIN==-9999] = NA
```

Okay, now we'll check again, for dates less then Dec. 31, 1913:
```{r}
plot(TMAX~DATE, import[import$DATE<19131231,], ty='l')
```

Yikes! What's wrong? As it turns it the problem is that with how the dates are specified. In particular, the Dec 31 to Jan 1 transition: 

19132131 -> 19140101, if you use these as numbers it should be 19131232, 19131232, etc. You see we are missing lots numbers!


```{r}
plot(TMAX~DATE, import[import$DATE<19130102,], ty='l')
```

## Fixing the Dates

First, we convert the date a string or character values. Next, we'll convert the strings to a data format. 

```{r}
strDates <- as.character(import$DATE)
head(strDates)
import$NewDate <- as.Date(strDates, "%Y%m%d")
```

## Checking the New Dates
```{r}
plot(TMAX~NewDate, import[import$DATE<19130102,], ty='l')
```

## A new plot
```{r}
plot(TMAX~NewDate, import, ty='l')
```

## Evaluating the trend

Creating a linear model and then evaluate the rate (slope). 
```{r}
<<<<<<< HEAD:David/CopyOfImporting Franks CSV.Rmd
FAIRPLEX <- subset(import, STATION_NAME=="HAGERSTOWN MD US", select=c(STATION, STATION_NAME, DATE, NewDate, TMIN, TMAX, PRCP))
=======
# Linear Model
import.lm <- lm(TMAX~NewDate, data=import)
summary(import.lm)
plot(TMAX~NewDate, import, ty='l')
abline(coef(import.lm), col='red')
>>>>>>> 87366cfb51d599c2a8821456a32daf84199e6282:Nicole/Importing CSV marshfield.Rmd
```

## Time Series

Time series data require specific statistical methods. I started thiking about how to do this in R, but didn't finish... so this is superfolous for now.
```{r}

#2.906*10-5 degrees/year or 2.9*10-2 per year.

# Moving Average!
# average of current sample and 19 previous samples (red)
f20 <- rep(1/600, 600)
y_lag <- filter(import$TMAX, f20, sides=1)
plot(y_lag)

# Create a time series
Tmax = ts(import$TMAX)
```

## Creating Month and Year Variables

So, let's figure out how to see how changes happen for individual months.
```{r}
# Get months
import$Month = months(import$NewDate)
import$Year = format(import$NewDate, format="%Y")
```

### Creating Monthly Means
```{r creatingmonthlymeans}
MonthlyMean = aggregate(TMAX ~ Month + Year, import, mean)

MonthlyMean$YEAR = as.numeric(MonthlyMean$Year)
head(MonthlyMean)
```


```{r plottingMonthyMeans}
plot(MonthlyMean$TMAX, ty='l')
```

### Selecting for 1 Month -- May

Perhaps, we can get a better handle on this stuff if we analyze for just one month at a time -- certainly easier to visualize!

```{r May}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="May"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="May",], ty='l') 
May.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="May",])
summary(May.lm)

abline(coef(May.lm), col="red")

```

Now, the change is `r coef(May.lm)[2]` degress/year or `r round(coef(May.lm)[2]*100, 3)` degress/100 years (YIKES!) with a probability of `r round(summary(May.lm)$coefficients[2,4],4)`. Although we can't reject the null hypothesis, it is worrying!

## Next Steps

I think you should evaluate every month and see what happens. You might also consider looking at the TMIN as well. Could be important!

## Lets try Dec

```{r}
plot(MonthlyMean$TMAX[MonthlyMean$Month=="April"], ty='l')
plot(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="April",], ty='l') 
April.lm <- lm(TMAX~YEAR, data=MonthlyMean[MonthlyMean$Month=="April",])
summary(April.lm)

abline(coef(April.lm), col="red")

```

## TMIN

1. We create a monthly mean for each month.

```{r}
MonthlyMeanTMIN = aggregate(TMIN ~ Month + Year, import, mean)

MonthlyMeanTMIN$YEAR = as.numeric(MonthlyMeanTMIN$Year)
head(MonthlyMeanTMIN)
```

2. Now we plot the mins, and again, find tons of scatter.

```{r}
plot(MonthlyMeanTMIN$TMIN, ty='l')

plot(MonthlyMeanTMIN$TMIN[MonthlyMeanTMIN$Month=="December"], ty='l')

plot(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="December",], ty='l') 
Dec.lm <- lm(TMIN~YEAR, data=MonthlyMeanTMIN[MonthlyMeanTMIN$Month=="December",])
summary(Dec.lm)

abline(coef(Dec.lm), col="red")

```


3. In this case, we get a slope, `r coef(Dec.lm)[2]` degress/year and a probability of `r round(summary(Dec.lm)$coefficients[2,4],7)` and an r-sqaured of `r round(summary(Dec.lm)$r.squared, 3)`. Cool! As we might expect, the a small amount of the variance is explained by the "Month". Many things predict temerpature, that year is one, is quite problematic.

4. What we have not determined is the cause. So, be careful when you describe the results, cause and effect cannot be analyzed using this method.



